### 2023-01-17 analysis.Rmd
manu

---
title: "PHQ-4-R: Adaptation of the 4-item screening for Anxiety and Depression for a subclinical sensitivity"
subtitle: "Analysis"
author: "Dominique Makowski et al."
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: paper
    code_folding: hide
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(
  digits = 3,
  mc.cores = 4,
  brms.algorithm = "sampling",
  brms.backend = "cmdstanr",
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)

fig.width <- see::golden_ratio(7)
fig.height <- 7
set.seed(333)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

# Study 1

## Participants

```{r, message=FALSE, warning=FALSE, results='asis'}
library(tidyverse)
library(easystats)
library(patchwork)
library(lavaan)
library(ggraph)
library(tidySEM)

df <- haven::read_sav("data/Study 1.sav") |> 
  mutate_all(as.numeric) |> 
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))

paste0(
  "Data from the [study 1](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769) (Murphy et al., 2020), downloaded from OSF, included ",
  report::report_participants(df, age = "Age", sex = NA, gender = "Gender"),
  "."
)
```

## Descriptive

```{r, message=FALSE, warning=FALSE}
data <- select(df, -Age, -Gender)

dens1 <- estimate_density(data, method="kernSmooth")

plot(dens1)

data <- standardize(data)
```

## EFA

```{r, message=FALSE, warning=FALSE}
n <- parameters::n_factors(data, n_max = 15)

n

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ])
plot(n)
```

```{r, message=FALSE, warning=FALSE}
efa1 <- parameters::factor_analysis(data, n=1, sort = TRUE)
efa1

efa4 <- parameters::factor_analysis(data, n=4, rotation = "oblimin", sort = TRUE)
efa4

plot(efa4)
```

## CFA

```{r, message=FALSE, warning=FALSE}
cfa1 <-  parameters::efa_to_cfa(efa1, names="IAS") |> 
  lavaan::cfa(data=data)

cfa4 <- parameters::efa_to_cfa(efa4, names=c("Skin", "Expulsion", "Interoception", "Elimination"), threshold = "max") |> 
  lavaan::cfa(data=data)

anova(cfa1, cfa4)
```

```{r, message=FALSE, warning=FALSE}
cfa4mod <- "
Skin =~ Affective_touch + Tickle + Itch
Expulsion =~ Sneeze + Cough + Wind + Burp
Interoception =~ Heart + Hungry + Breathing + Thirsty + Temp + Sex_arousal + Muscles + Bruise + Pain
Elimination =~ Urinate + Defecate + Vomit
" |> 
lavaan::cfa(data=data)

cfa5 <- "
Skin =~ Affective_touch + Tickle + Itch
Expulsion =~ Sneeze + Cough + Wind + Burp
Interoception =~ Heart + Hungry + Breathing + Thirsty + Temp + Sex_arousal
Nociception =~ Muscles + Bruise + Pain
Elimination =~ Urinate + Defecate + Vomit
" |> 
lavaan::cfa(data=data)

anova(cfa4mod, cfa5)
```

```{r, message=FALSE, warning=FALSE}
cfa5mod <- "
Skin =~ Affective_touch + Tickle + Itch
Expulsion =~ Sneeze + Wind + Burp
Interoception =~ Heart + Hungry + Breathing + Thirsty + Temp + Sex_arousal
Nociception =~ Muscles + Bruise + Pain
Elimination =~ Urinate + Defecate + Vomit
" |> 
lavaan::cfa(data=data)

anova(cfa5, cfa5mod)
```

```{r, message=FALSE, warning=FALSE}
cfa5h <- "
Skin =~ Affective_touch + Tickle + Itch
Expulsion =~ Sneeze + Wind + Burp
Nociception =~ Muscles + Bruise + Pain
Elimination =~ Urinate + Defecate + Vomit
Interoception =~ Heart + Hungry + Breathing + Thirsty + Temp + Sex_arousal + Skin + Expulsion + Nociception + Elimination
" |> 
lavaan::cfa(data=data)

anova(cfa5mod, cfa5h)
```

```{r, message=FALSE, warning=FALSE}
check_perf <- function(mod) {
  perf <- model_performance(mod, metrics = c("RMSEA", "CFI", "SRMR", "AGFI", "NNFI"))
  interpret(perf) |> 
    data_rename(pattern = c("Name", "Value"), replacement = c("Index", "Score")) |> 
    arrange(Index)
}

knitr::kable(check_perf(cfa5mod))
```

## New sample

```{r, message=FALSE, warning=FALSE, results='asis'}
df2 <- haven::read_sav("data/Study 6 IAS.sav") |> 
  mutate_all(as.numeric) |> 
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other")))) |> 
  select(-ESL)

paste0(
  "Data from the [study 6](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769) (Murphy et al., 2020), downloaded from OSF, included ",
  report::report_participants(df2, age = "Age", sex = NA, gender = "Gender"),
  "."
)
```

### Descriptive

```{r, message=FALSE, warning=FALSE}
data <- select(df2, -Age, -Gender)

dens2 <- estimate_density(data, method="kernSmooth")

plot(dens2)

data <- standardize(data)
```

```{r, message=FALSE, warning=FALSE}
efa4b <- parameters::factor_analysis(data, n=4, rotation = "oblimin", sort = TRUE)
```

```{r, message=FALSE, warning=FALSE}
cfa1_refit <- "IAS =~ Pain + Sneeze + Temp + Wind + Affective_touch + Muscles + Tickle + Itch + Defecate + Vomit + Taste + Cough + Bruise + Burp + Urinate + Breathing + Thirsty + Sex_arousal + Hungry + Heart + Blood_Sugar" |> lavaan::cfa(data=data)
cfa4mod_refit <- update(cfa4mod, data=data)
cfa5_refit <- update(cfa5, data=data)
cfa5mod_refit <- update(cfa5mod, data=data)
cfa5h_refit <- update(cfa5h, data=data)

table <- anova(cfa5mod_refit, cfa5h_refit, cfa5_refit, cfa4mod_refit, cfa1_refit)
table
```

```{r, message=FALSE, warning=FALSE}
knitr::kable(check_perf(cfa5mod_refit))
```

# Study 2

## Participants

```{r, message=FALSE, warning=FALSE, results='asis'}
# df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |>
df3 <- read.csv("../IllusionGameReliability/data/preprocessed_questionnaires.csv") |>
  select(Participant, Sex, Age, starts_with("Item_IAS")) |> 
  filter(!is.na(Item_IAS_Elimination_2))

names(df3) <- str_remove_all(names(df3), "Item_IAS_")
names(df3) <- case_when(
  names(df3) == "Interoception_1" ~ "Heart",
  names(df3) == "Interoception_2" ~ "Hungry",
  names(df3) == "Interoception_3" ~ "Breathing",
  names(df3) == "Interoception_4" ~ "Thirsty",
  names(df3) == "Interoception_5" ~ "Temp",
  names(df3) == "Interoception_6" ~ "Sex_arousal",
  names(df3) == "Elimination_1" ~ "Urinate",
  names(df3) == "Elimination_2" ~ "Defecate",
  names(df3) == "Elimination_3" ~ "Vomit",
  names(df3) == "Expulsion_1" ~ "Wind",
  names(df3) == "Expulsion_2" ~ "Burp",
  names(df3) == "Expulsion_3" ~ "Sneeze",
  names(df3) == "Nociception_1" ~ "Muscles",
  names(df3) == "Nociception_2" ~ "Bruise",
  names(df3) == "Nociception_3" ~ "Pain",
  names(df3) == "Skin_1" ~ "Affective_touch",
  names(df3) == "Skin_2" ~ "Tickle",
  names(df3) == "Skin_3" ~ "Itch",
  TRUE ~ names(df3)
  )

report::report_participants(df3, age = "Age", sex = "Sex")
```

## Descriptive

```{r, message=FALSE, warning=FALSE}
data <- select(df3, -Age, -Sex, -Participant)

dens3 <- estimate_density(data, method="kernSmooth")

plot(dens3) +
  theme_modern()

data[] <- standardize(data) |> 
  sapply(as.numeric)
```

## EFA

```{r, message=FALSE, warning=FALSE}
n <- parameters::n_factors(data, n_max = 15)

n
plot(n)
```

```{r, message=FALSE, warning=FALSE}
efa4c <- parameters::factor_analysis(data, n=4, rotation = "oblimin", sort = TRUE, fm="ml")
efa4c

plot(efa4c)
```

##CFA

```{r, message=FALSE, warning=FALSE}
cfa1 <- paste("Interoception =~ ", paste(names(data), collapse = " + ")) |> 
  lavaan::cfa(data=data)

cfa5 <- "
Skin =~ Affective_touch + Tickle + Itch
Expulsion =~ Wind + Burp + Sneeze
Interoception =~ Heart + Hungry + Breathing + Thirsty + Temp + Sex_arousal
Nociception =~ Muscles + Bruise + Pain
Elimination =~ Urinate + Defecate + Vomit
" |> 
  lavaan::cfa(data=data)

cfa4 <- "
Homeostasis =~ Hungry + Thirsty + Sex_arousal + Pain + Temp + Affective_touch + Urinate
Skin =~ Affective_touch + Tickle + Itch + Bruise + Wind
Elimination =~ Defecate + Sneeze + Vomit + Urinate + Burp
Muscle =~ Breathing + Heart + Muscles
" |> 
  lavaan::cfa(data=data)

anova(cfa1, cfa5, cfa4)

cfa4mod <- "
Homeostasis =~ Hungry + Thirsty + Sex_arousal 
Skin =~ Tickle + Itch + Bruise 
Elimination =~ Defecate + Sneeze + Vomit 
Muscle =~ Breathing + Heart + Muscles
" |> 
  lavaan::cfa(data=data)

cfa5short <- "
Skin =~ Tickle + Itch
Expulsion =~ Sneeze
Interoception =~ Heart + Hungry + Breathing + Thirsty + Sex_arousal
Nociception =~ Muscles + Bruise
Elimination =~ Defecate + Vomit
"

anova(cfa4mod, lavaan::cfa(cfa5short, data=data))

cfa4h <- "
Homeostasis =~ Hungry + Thirsty + Sex_arousal 
Skin =~ Tickle + Itch + Bruise 
Elimination =~ Defecate + Sneeze + Vomit 
Muscle =~ Breathing + Heart + Muscles
Interoception =~ Homeostasis + Skin + Elimination + Muscle
" |> 
  lavaan::cfa(data=data)

anova(cfa4mod, cfa4h)

check_perf(cfa4mod)

check_perf(cfa4h)
check_perf(update(cfa4h, data=standardize(df)))
check_perf(update(cfa4h, data=standardize(df2)))
```

```{r, message=FALSE, warning=FALSE, fig.width=16, fig.height=10}
graph_data <- tidySEM::prepare_graph(cfa4h, 
                                     variance_diameter=NA)

nodes <- graph_data$nodes |> 
  mutate(fill = ifelse(shape == "rect", "grey", "#F06292"),
         textsize = ifelse(shape == "rect", rel(3.5), rel(4)),
         fontface = ifelse(shape == "oval", "bold", "plain"),
         hjust = ifelse(shape == "oval", 0.5, -0.3),
         textcolor = ifelse(shape == "oval", "white", "black"),
         label = str_replace(label, "Sex_arousal", "Arousal")) |> 
  select(-x, -y)

edges <- graph_data$edges |>
  mutate(est = as.numeric(est))

p_graph <- tidygraph::tbl_graph(nodes = nodes, edges = edges) |> 
  ggraph(layout = "kk") +
  geom_edge_arc(aes(filter=lhs!="Interoception",
                    edge_width = est,
                    label=est_std),
                strength = 0.15,
                color="#2196F3",
                angle_calc="along",
                label_dodge=unit(-0.02, "npc"),
                label_size = rel(4)) +
  geom_edge_link(aes(filter=lhs=="Interoception",
                     edge_width = est, 
                     # color = color,
                     label=est_std), 
                 color="#3F51B5",
                 angle_calc="along", 
                 label_dodge=unit(-0.02, "npc"),
                 label_size = rel(5)) +
  geom_node_point(aes(shape = shape, color=name, size=shape)) +
  geom_node_text(aes(label = label, fontface=shape),
                 size=nodes$textsize,
                 fontface = nodes$fontface,
                 hjust = nodes$hjust,
                 color = nodes$textcolor) +
  # scale_y_continuous(expand = expansion(add=c(0.1, 0.1))) +
  scale_x_continuous(expand = expansion(add=c(0.3, 0.4))) +
  scale_edge_width_continuous(range=c(0.5, 2), guide="none") +
  scale_color_manual(values=c("Interoception" = "red",
                              "Muscles" = "#E91E63",
                              "Heart" = "#E91E63",
                              "Breathing" = "#E91E63",
                              "Muscle" = "#E91E63",
                              "Skin" = "#FF5722",
                              "Tickle" = "#FF5722",
                              "Itch" = "#FF5722",
                              "Bruise" = "#FF5722",
                              "Elimination" = "#FFC107",
                              "Defecate" = "#FFC107",
                              "Sneeze" = "#FFC107",
                              "Vomit" = "#FFC107",
                              "Homeostasis" = "#9C27B0",
                              "Hungry" = "#9C27B0",
                              "Thirsty" = "#9C27B0",
                              "Sex_arousal" = "#9C27B0"), guide="none") +
  scale_shape_manual(values=c("oval"="circle", "rect"="square"), guide="none") +
  scale_size_manual(values = c(35, 5), guide="none") +
  labs(title = "C. Structural Equation Model")  +
  theme_graph() +
  theme(plot.title = element_text(hjust=0.5))
p_graph
```

## Comparison with MAIA

```{r, message=FALSE, warning=FALSE}
scores <- as.data.frame(predict(cfa4h))
scores$Participant <- df3$Participant
write.csv(scores, "data/IASR_scores.csv", row.names = FALSE)
```

# Study 3

```{r, message=FALSE, warning=FALSE}
df <- read.csv("../IllusionGameReliability/data/preprocessed_questionnaires.csv") |> 
  merge(read.csv("data/IASR_scores.csv"), by="Participant") |> 
  select(-ends_with("_RT"), -ends_with("_SD"))

df$Average <- rowMeans(select(df, starts_with("Item_IAS")))

vars <- c("Homeostasis", "Skin", "Elimination", "Muscle", "Interoception")

# correlation(df[vars], select(df, starts_with("Item_IAS"))) |> 
#   summary() |> 
#   plot()
```

## Personality

```{r, message=FALSE, warning=FALSE}
make_correlation <- function(x, y) {
  set.seed(3)
  cor <- correlation::correlation(
    x, y,
    bayesian = TRUE,
    bayesian_prior = "medium.narrow"
  ) |>
    datawizard::data_remove(c("ROPE_Percentage"))

  cor$BF_Spearman <- format_bf(
    correlation::correlation(
      x, y,
      bayesian = TRUE,
      ranktransform = TRUE,
      bayesian_prior = "medium.narrow"
    )$BF,
    name = NULL, stars = FALSE
  )

  r <- filter(cor, BF >= 10)

  if(nrow(r) == 0) return(list(data = cor, print = "No significant (BF >= 10) correlations."))

  to_print <- r |>
    arrange(Parameter1, desc(BF)) |>
    format_table() |>
    mutate(Coefficient = paste(rho, `95% CI`)) |>
    select(Parameter1, Parameter2, Coefficient, BF, `BF (Spearman)` = BF_Spearman)

  list(data = cor, print = to_print)
}

ipip <- make_correlation(df[vars], select(df, starts_with("IPIP6_"))) 
export_table(ipip$print)

pid <- make_correlation(df[vars], select(df, starts_with("PID5_"))) 
export_table(pid$print)
```

## Psychopathology

```{r, message=FALSE, warning=FALSE}
asq <- make_correlation(df[vars], select(df, starts_with("ASQ_")))
export_table(asq$print)
spq <- make_correlation(df[vars], select(df, starts_with("SPQ_"))) 
export_table(spq$print)
bpd <- make_correlation(df[vars], select(df, starts_with("BPD"))) 
export_table(bpd$print)
phq <- make_correlation(df[vars], select(df, starts_with("PHQ4_"))) 
export_table(phq$print)
```

## Other

```{r, message=FALSE, warning=FALSE}
pi <- make_correlation(df[vars], select(df, starts_with("PI_")))
export_table(pi$print)
gcbs <- make_correlation(df[vars], select(df, starts_with("GCBS_"))) 
export_table(gcbs$print)
age <- make_correlation(df[vars], data.frame(Age = df$Age)) 
export_table(age$print)
sex <- make_correlation(df[vars], data.frame(Sex = ifelse(df$Sex == "Male", 0, 1)) )
export_table(sex$print)
```

```{r, message=FALSE, warning=FALSE}
p_cor <- rbind(
  mutate(ipip$data, Domain = "Normal Personality"),
  mutate(pid$data, Domain = "Maladaptive Personality"),
  mutate(asq$data, Domain = "Autistic Traits"),
  mutate(spq$data, Domain = "Schizotypic Traits"),
  mutate(phq$data, Domain = "Mood Disorders"),
  mutate(gcbs$data, Domain = "Conspiracy Beliefs")
  # mutate(age$data, Domain = "Demographic"),
  # mutate(sex$data, Domain = "Demographic")
) |> 
  mutate(alpha = ifelse(BF >= 10, "BF >= 10", "BF < 10"),
         Parameter1 = fct_relevel(Parameter1, "Interoception", "Muscle", "Homeostasis", "Elimination", "Skin"),
         Domain = fct_relevel(Domain, "Normal Personality", "Maladaptive Personality", "Autistic Traits", "Schizotypic Traits"),
         Parameter2 = str_remove_all(Parameter2, "IPIP6_|PID5_|ASQ_|SPQ_|GCBS_"),
         Parameter2 = str_replace(Parameter2, "Sex", "Sex (Female vs. Male)"),
         Parameter2 = str_replace(Parameter2, "HonestyHumility", "Honesty-Humility"),
         color = as.character(sign(rho))) |>
  ggplot(aes(x = rho, y = Parameter2)) +
  geom_bar(aes(alpha = alpha, fill = rho), stat = "identity") +
  geom_linerange(aes(alpha = alpha, xmin = CI_low, xmax = CI_high, color = color)) +
  geom_vline(xintercept = 0) +
  facet_grid(Domain~Parameter1, scales="free_y", switch ="y") +
  scale_alpha_manual(values = c(0.2, 1)) +
  scale_fill_gradient2(low = "#2196F3", high = "#FF5722", limits = c(-0.25, 0.25)) +
  scale_colour_manual(values = c("#2196F3", "#FF5722")) +
  scale_x_continuous(expand = c(0, 0)) +
  coord_cartesian(xlim = c(-0.3, 0.3)) +
  labs(title = "Correlates of IAS-R Dimensions") +
  guides(alpha = "none", fill = "none", color="none") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        strip.placement.y = "outside",
        strip.background.x = element_rect(fill = "#E0E0E0", color = "white"),
        strip.text = element_text(face = "bold", size = rel(1)))
p_cor
```

# Figures

```{r, message=FALSE, warning=FALSE}
p_dens <- rbind(
  mutate(dens1, Study = "Study 1a"),
  filter(mutate(dens2, Study = "Study 1b"), Parameter %in% dens1$Parameter),
  filter(mutate(dens3, Study = "Study 2"), Parameter %in% dens1$Parameter)
) |> 
  ggplot(aes(x=x, y=y, color=Parameter)) +
  geom_line() +
  facet_wrap(~Study, scales = "free", ncol = 1, strip.position = "right") +
  scale_color_metro_d() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(y = "Density", x = "Answer", title = "A. Distribution of Answers") +
  guides(color = guide_legend(title="Item", ncol=1, byrow=TRUE, keyheight=0)) +
  theme_modern(axis.title.space = 5) +
  theme(axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0),
        legend.text = element_text(size = rel(0.8)),
        legend.spacing.y = unit(0.1, 'cm'),
        strip.background = element_rect(fill = "#E0E0E0", color = "white"))

p_efa <- rbind(
  attributes(efa4)$dataset[names(data)] |> 
    parameters::factor_analysis(n=4, rotation = "oblimin", sort = TRUE, fm="ml") |> 
    data_plot() |> 
    mutate(Study = "Study 1a"),
  attributes(efa4b)$dataset[names(data)] |> 
    parameters::factor_analysis(n=4, rotation = "oblimin", sort = TRUE, fm="ml") |> 
    data_plot() |> 
    mutate(Study = "Study 1b"),
  efa4c |> 
    data_plot() |> 
    mutate(Study = "Study 2") 
) |> 
  mutate(Variable = fct_relevel(Variable, sort(levels(Variable), decreasing = TRUE)),
         fill = as.factor(sign(y)),
         x = abs(y),
         Component = str_replace_all(Component, "ML", "Factor ")) |> 
  ggplot(aes(x = x, y = Variable, fill = fill)) +
  geom_bar(stat = "identity") + 
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75), labels = c("0", ".25", ".50", ".75")) +
  scale_fill_manual(values = c("#E91E63", "#4CAF50")) +
  labs(x = "Loading", title = "B. Exploratory Factor Analysis") +
  guides(fill = "none") +
  theme_modern(axis.title.space = 5) +
  theme(panel.grid.major.y = element_line(),
        axis.title.y = element_blank(),
        axis.line.x = element_blank(),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(size = rel(0.8)),
        plot.title = element_text(face = "bold", hjust = 0),
        strip.placement = "outside",
        strip.background.y = element_rect(fill = "#E0E0E0", color = "white"),
        strip.background.x = element_rect(fill = "#F8BBD0", color = "white")) +
  facet_grid(Study ~ Component, switch = "y")

fig <- ((wrap_elements(p_dens) + wrap_elements(p_efa)) / p_graph)
ggsave("figures/Figure1.png", fig, width=fig.width, height=fig.width*1.3, dpi=150)
ggsave("figures/Figure2.png", p_cor, width=21*0.35, height=29.7*0.35, dpi=150, bg="white")
```