### 2024-02-13 analysis/analysis.qmd
update analysis

---
title: "Revised Interoceptive Accuracy Scale (IAS-R)"
subtitle: "Analysis"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    fig-dpi: 300
    fig-width: 11.33
execute: 
  cache: true
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(lavaan)
library(ggraph)
library(tidySEM)
library(EGAnet)
```

## Study 1 - Reanalysis of Murphy et al. (2020)

Reanalyze their data to confirm the factor structure of the IAS.

### Data Preparation

```{r}
# https://osf.io/3m5nh
df1 <- haven::read_sav("../data/Murphy2020/Study 1.sav") |> 
  mutate_all(as.numeric) |> 
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))

# https://osf.io/3m5nh
df2 <- haven::read_sav("../data/Murphy2020/Study 6 IAS.sav") |> 
  mutate_all(as.numeric) |> 
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))

# Campos (2022) - Study 1 (https://osf.io/j6ef3)
df3 <- haven::read_sav("../data/Campos2022/Dataset_Test.sav") |> 
  mutate_all(as.numeric) |> 
  mutate(Gender = as.character(ifelse(Sex == 1, "Male", ifelse(Sex == 0, "Female", "Other")))) |> 
  rename(Heart=IAS1, Hungry=IAS2, Breathing=IAS3, Thirsty=IAS4,
         Urinate=IAS5, Defecate=IAS6, Taste=IAS7, Vomit=IAS8, Sneeze=IAS9,
         Cough=IAS10, Temp=IAS11, Sex_arousal=IAS12, Wind=IAS13, Burp=IAS14,
         Muscles=IAS15, Bruise=IAS16, Pain=IAS17, Blood_Sugar=IAS18,
         Affective_touch=IAS19, Tickle=IAS20, Itch=IAS21)

# https://osf.io/3eztd
df4 <- haven::read_sav("../data/Lin2023/Study 1 & 3.sav") |> 
  mutate(Gender = as.character(ifelse(sex_dummy == 1, "Male", ifelse(sex_dummy == 0, "Female", "Other")))) |> 
  rename(Age=age, 
         Heart=Heart, Hungry=HUNGR, Breathing=BREAT, Thirsty=Thirs,
         Urinate=URINA, Defecate=Defec, Taste=TASTE, Vomit=VOMIT, Sneeze=Sneez,
         Cough=COUGH, Temp=TEMPE, Sex_arousal=SEXAR, Wind=WIND, Burp=Burp,
         Muscles=MUSCL, Bruise=Bruis, Pain=PAIN, Blood_Sugar=BloSu,
         Affective_touch=Touch, Itch=ITCH)  # No tickle because same Chinese character

# https://osf.io/3eztd
df5 <- haven::read_sav("../data/Lin2023/Study 2.sav") |> 
  mutate(Gender = as.character(ifelse(Sex == "男", "Male", ifelse(Sex == "女", "Female", "Other")))) |> 
  rename(Heart=Heart, Hungry=HUNGR, Breathing=BREAT, Thirsty=Thirs,
         Urinate=URINA, Defecate=Defec, Taste=TASTE, Vomit=VOMIT, Sneeze=Sneez,
         Cough=COUGH, Temp=TEMPE, Sex_arousal=SEXAR, Wind=WIND, Burp=Burp,
         Muscles=MUSCL, Bruise=Bruis, Pain=PAIN, Blood_Sugar=BloSu,
         Affective_touch=Touch, Itch=ITCH)  # No tickle because same Chinese character

df6 <- read.csv("https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data.csv") |>
  rename(Heart=IAS_1, Hungry=IAS_2, Breathing=IAS_3, Thirsty=IAS_4, 
         Urinate=IAS_5, Defecate=IAS_6, Taste=IAS_7, Vomit=IAS_8, Sneeze=IAS_9, 
         Cough=IAS_10, Temp=IAS_11, Sex_arousal=IAS_12, Wind=IAS_13, Burp=IAS_14,
         Muscles=IAS_15, Bruise=IAS_16, Pain=IAS_17, Blood_Sugar=IAS_18,
         Affective_touch=IAS_19, Tickle=IAS_20, Itch=IAS_21)

df7 <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |>
  rename(Gender=Sex, Heart=Item_IAS_Interoception_1, Hungry=Item_IAS_Interoception_2, 
         Breathing=Item_IAS_Interoception_3, Thirsty=Item_IAS_Interoception_4, 
         Temp=Item_IAS_Interoception_5, Sex_arousal=Item_IAS_Interoception_6,
         Urinate=Item_IAS_Elimination_1, Defecate=Item_IAS_Elimination_2,
         Vomit=Item_IAS_Elimination_3, Wind=Item_IAS_Expulsion_1, 
         Burp=Item_IAS_Expulsion_2, Sneeze=Item_IAS_Expulsion_3,
         Muscles=Item_IAS_Nociception_1, Bruise=Item_IAS_Nociception_2,
         Pain=Item_IAS_Nociception_3, Affective_touch=Item_IAS_Skin_1, 
         Tickle=Item_IAS_Skin_2, Itch=Item_IAS_Skin_3) |>
  filter(!is.na(Urinate))
```

### Participants

- **Sample 1**: Data from Murphy's (2020) [study 1](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769), downloaded from OSF, included `r report::report_participants(df1, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 2**: Data from Murphy's (2020) [study 6](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769), downloaded from OSF, included `r report::report_participants(df2, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 3**: Data from Campos' (2022) [study 1](https://osf.io/j6ef3), downloaded from OSF, included `r report::report_participants(df3, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 4**: Data from Lin's (2023) [study 1 and 3](#ttps://osf.io/3eztd), downloaded from OSF, included `r report::report_participants(df4, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 5**: Data from Lin's (2023) [study 2](#ttps://osf.io/3eztd), downloaded from OSF, included `r report::report_participants(df5, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 6**: New dataset from [Makowski (under preparation)](https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data.csv), included `r report::report_participants(df6, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 7**: New dataset from [Makowski (under preparation)](https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv), included `r report::report_participants(df7, age = "Age", sex = NA, gender = "Gender")`.

Total N = `r nrow(df1) + nrow(df2) + nrow(df3) + nrow(df4) + nrow(df5) + nrow(df6) + nrow(df7)`.

### Descriptive

#### Distribution

```{r}
vars <- c("Heart", "Hungry", "Breathing", "Thirsty", "Urinate", "Defecate", "Taste", "Vomit", "Sneeze", "Cough", "Temp",
          "Sex_arousal", "Wind", "Burp", "Muscles", "Bruise", "Pain", "Blood_Sugar", "Affective_touch", "Tickle", "Itch")

dens1 <- estimate_density(select(df1, all_of(vars)), method="kernSmooth") |> 
  mutate(Sample = "Sample 1")
dens2 <- estimate_density(select(df2, all_of(vars)), method="kernSmooth") |> 
  mutate(Sample = "Sample 2")
dens3 <- estimate_density(select(df3, all_of(vars)), method="kernSmooth") |> 
  mutate(Sample = "Sample 3")
dens4 <- estimate_density(select(df4, all_of(setdiff(vars, "Tickle"))), method="kernSmooth") |> 
  mutate(Sample = "Sample 4")
dens5 <- estimate_density(select(df5, all_of(setdiff(vars, "Tickle"))), method="kernSmooth") |> 
  mutate(Sample = "Sample 5")
dens6 <- estimate_density(select(df6, all_of(vars)), method="kernSmooth") |> 
  mutate(Sample = "Sample 6")
dens7 <- estimate_density(select(df7, all_of(setdiff(vars, c("Taste", "Cough", "Blood_Sugar")))), method="kernSmooth") |> 
  mutate(Sample = "Sample 7")

rbind(dens1, dens2, dens3, dens4, dens5, dens6, dens7) |> 
  ggplot(aes(x = x, y = y, color = Parameter)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Distribution of Responses", x = "Response", y = "Density", color = "Item") +
  facet_wrap(~Sample, scales = "free", nrow=4)
```

```{r}
data1 <- normalize(select(df1, all_of(dens1$Parameter)), range=c(1, 5))
data2 <- normalize(select(df2, all_of(dens2$Parameter)), range=c(1, 5))
data3 <- normalize(select(df3, all_of(dens3$Parameter)), range=c(1, 5))
data4 <- normalize(select(df4, all_of(dens4$Parameter)), range=c(1, 5))
data5 <- normalize(select(df5, all_of(dens5$Parameter)), range=c(1, 5))
data6 <- select(df6, all_of(dens6$Parameter))
data7 <- select(df7, all_of(dens7$Parameter))

data_all <- rbind(data1, data2, data3, 
                  mutate(data4, Tickle=NA), mutate(data5, Tickle=NA), 
                  data6, mutate(data7, Taste=NA, Cough=NA, Blood_Sugar=NA)) 
```

#### Correlations

An overall positive intercorrelation pattern, with no clear structure emerging.

::: {.panel-tabset}

##### All data

```{r}
make_correlation <- function(df) {
  correlation::correlation(df, redundant=TRUE) |> 
    correlation::cor_sort() |> 
    correlation::cor_lower() |> 
    mutate(val = paste0(insight::format_value(r), format_p(p, stars_only=TRUE))) |>
    mutate(Parameter2 = fct_rev(Parameter2)) |>
    ggplot(aes(x=Parameter1, y=Parameter2)) +
    geom_tile(aes(fill = r), color = "white") +
    geom_text(aes(label = val), size = 3) +
    labs(title = "Correlation Matrix") +
    scale_fill_metro_c(limit = c(0, 0.75), guide = guide_colourbar(ticks=FALSE)) +
    theme_minimal() +
    theme(legend.title = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1))
}

make_correlation(data_all)
```

##### Sample 1

```{r}
make_correlation(data1)
```

##### Sample 2

```{r}
make_correlation(data2)
```

##### Sample 3

```{r}
make_correlation(data3)
```

##### Sample 4

```{r}
make_correlation(data4)
```

##### Sample 5

```{r}
make_correlation(data5)
```

##### Sample 6

```{r}
make_correlation(data6)
```

##### Sample 7

```{r}
make_correlation(data7)
```

:::

### EGA

See https://r-ega.net/articles/ega.html for details.

#### Unique Variable Analysis (UVA)

::: {.panel-tabset}

##### All data

```{r}
uva <- EGAnet::UVA(data = data_all, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 1

```{r}
uva <- EGAnet::UVA(data = data1, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 2

```{r}
uva <- EGAnet::UVA(data = data2, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 3

```{r}
uva <- EGAnet::UVA(data = data3, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 4

```{r}
uva <- EGAnet::UVA(data = data4, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 5

```{r}
uva <- EGAnet::UVA(data = data5, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 6

```{r}
uva <- EGAnet::UVA(data = data6, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 7

```{r}
uva <- EGAnet::UVA(data = data7, cut.off = 0.3)
uva
uva$keep_remove
```

:::

#### Item Stability

```{r ega1}
#| fig-height: 20
#| fig-width: 12

plots <- list()
for(model in c("glasso", "TMFG")) {
  for(algo in c("walktrap", "louvain")) {
    for(type in c("ega", "ega.fit")) {  # "hierega"
      print(type)
      ega <- EGAnet::bootEGA(
        data = data_all,
        seed=123,
        model=model,
        algorithm=algo,
        EGA.type=type,
        type="resampling",
        plot.itemStability=FALSE,
        verbose=FALSE)
      plots[[length(plots) + 1]] <- plot(ega) + 
        labs(title = paste0(model, " (", algo, ") - ", type))
      }
    }
}

length(plots)
patchwork::wrap_plots(plots, nrow = 4)
```

<!-- ::: {.panel-tabset} -->

<!-- ##### glasso model (Walktrap) -->

<!-- ```{r ega1} -->
<!-- #| fig-height: 12 -->

<!-- plots <- list() -->
<!-- for(d in list(data1, data2, data3, data4, data5, data6)) { -->
<!--   ega <- EGAnet::bootEGA( -->
<!--     data = d, -->
<!--     seed=123, -->
<!--     model="glasso", -->
<!--     algorithm="walktrap", -->
<!--     type="resampling", -->
<!--     EGA.type="ega", -->
<!--     plot.itemStability=FALSE,  -->
<!--     verbose=FALSE) -->
<!--   plots[[length(plots) + 1]] <- plot(ega) -->
<!-- } -->

<!-- EGAnet::dimensionStability(ega) -->
<!-- (plots[[1]] + labs(title = "Sample 1") + -->
<!--   plots[[2]] + labs(title = "Sample 2")) / -->
<!--   (plots[[3]] + labs(title = "Sample 3") + -->
<!--   plots[[4]] + labs(title = "Sample 4")) / -->
<!--   (plots[[5]] + labs(title = "Sample 5") + -->
<!--   plots[[6]] + labs(title = "Sample 6")) -->
<!-- ``` -->

<!-- ##### glasso model (Louvain) -->

<!-- ```{r ega2} -->
<!-- #| fig-height: 12 -->

<!-- plots <- list() -->
<!-- for(d in list(data1, data2, data3, data4, data5, data6)) { -->
<!--   ega <- EGAnet::bootEGA( -->
<!--     data = d, -->
<!--     seed=123, -->
<!--     model="glasso", -->
<!--     algorithm="louvain", -->
<!--     plot.itemStability=FALSE) -->
<!--   plots[[length(plots) + 1]] <- plot(ega) -->
<!-- } -->

<!-- (plots[[1]] + labs(title = "Sample 1") + -->
<!--   plots[[2]] + labs(title = "Sample 2")) / -->
<!--   (plots[[3]] + labs(title = "Sample 3") + -->
<!--   plots[[4]] + labs(title = "Sample 4")) / -->
<!--   (plots[[5]] + labs(title = "Sample 5") + -->
<!--   plots[[6]] + labs(title = "Sample 6")) -->
<!-- ``` -->

<!-- ##### TMFG model (Walktrap) -->

<!-- ```{r ega3} -->
<!-- #| fig-height: 12 -->

<!-- plots <- list() -->
<!-- for(d in list(data1, data2, data3, data4, data5, data6)) { -->
<!--   ega <- EGAnet::bootEGA( -->
<!--     data = d, -->
<!--     seed=123, -->
<!--     model="TMFG", -->
<!--     algorithm="walktrap", -->
<!--     plot.itemStability=FALSE) -->
<!--   plots[[length(plots) + 1]] <- plot(ega) -->
<!-- } -->

<!-- (plots[[1]] + labs(title = "Sample 1") + -->
<!--   plots[[2]] + labs(title = "Sample 2")) / -->
<!--   (plots[[3]] + labs(title = "Sample 3") + -->
<!--   plots[[4]] + labs(title = "Sample 4")) / -->
<!--   (plots[[5]] + labs(title = "Sample 5") + -->
<!--   plots[[6]] + labs(title = "Sample 6")) -->
<!-- ``` -->

<!-- ##### TMFG model (Louvain) -->

<!-- ```{r ega4} -->
<!-- #| fig-height: 12 -->

<!-- plots <- list() -->
<!-- for(d in list(data1, data2, data3, data4, data5, data6)) { -->
<!--   ega <- EGAnet::bootEGA( -->
<!--     data = d, -->
<!--     seed=123, -->
<!--     model="TMFG", -->
<!--     algorithm="louvain", -->
<!--     plot.itemStability=FALSE) -->
<!--   plots[[length(plots) + 1]] <- plot(ega) -->
<!-- } -->

<!-- (plots[[1]] + labs(title = "Sample 1") + -->
<!--   plots[[2]] + labs(title = "Sample 2")) / -->
<!--   (plots[[3]] + labs(title = "Sample 3") + -->
<!--   plots[[4]] + labs(title = "Sample 4")) / -->
<!--   (plots[[5]] + labs(title = "Sample 5") + -->
<!--   plots[[6]] + labs(title = "Sample 6")) -->
<!-- ``` -->

<!-- ##### All data -->

<!-- ```{r ega1} -->
<!-- #| fig-height: 12 -->

<!-- plots <- list() -->
<!-- for(model in c("glasso", "TMFG")) { -->
<!--   for(algo in c("walktrap", "louvain")) { -->
<!--     ega <- EGAnet::bootEGA( -->
<!--       data = data_all, -->
<!--       seed=123, -->
<!--       model=model, -->
<!--       algorithm=algo, -->
<!--       plot.itemStability=FALSE) -->
<!--     plots[[length(plots) + 1]] <- plot(ega) +  -->
<!--       labs(title = paste0(model, " (", algo, ")")) -->
<!--   } -->
<!-- } -->

<!-- patchwork::wrap_plots(plots) -->
<!-- ``` -->

<!-- ::: -->

### Exploratory Factor Analysis (EFA)

#### How Many Factors

::: {.panel-tabset}

##### All data

```{r}
n <- parameters::n_factors(data_all, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4, 5 ), ], row.names = FALSE)
```

##### Sample 1

```{r}
n <- parameters::n_factors(data1, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 2

```{r}
n <- parameters::n_factors(data2, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 3

```{r}
n <- parameters::n_factors(data3, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 4

```{r}
n <- parameters::n_factors(data4, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 5

```{r}
n <- parameters::n_factors(data5, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 6

```{r}
n <- parameters::n_factors(data6, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 7

```{r}
n <- parameters::n_factors(data7, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

:::

#### EFA Models

::: {.panel-tabset}

##### All data

```{r}
#| fig-height: 10

efa5 <- parameters::factor_analysis(data_all, n=5, rotation = "oblimin", sort = TRUE)
plot(efa5)

display(efa5)
```

##### Sample 1

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data2, n=4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

##### Sample 2

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data2, n=4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

##### Sample 3

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data3, n=4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

##### Sample 4

```{r}
#| fig-height: 10

efa5 <- parameters::factor_analysis(data4, n=5, rotation = "oblimin", sort = TRUE)
plot(efa5)

display(efa5)
```

##### Sample 5

```{r}
#| fig-height: 10

efa5 <- parameters::factor_analysis(data5, n=5, rotation = "oblimin", sort = TRUE)
plot(efa5)

display(efa5)
```

##### Sample 6

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data5, n=4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

:::

### Model Construction

#### Structure

```{r}
m1 <- "
Interoception =~ Hungry + Thirsty + Urinate + Defecate + Itch + Bruise + Muscles + Pain + Breathing + Heart + Cough + Sneeze + Wind + Burp
"

m7 <- "
# ----
Sustenance =~ Hungry + Thirsty
Elimination =~ Urinate + Defecate
Skin =~ Itch + Bruise
Nociception =~ Muscles + Pain
Affect =~ Breathing + Heart
# ----
Expulsion =~ Cough + Sneeze + Wind + Burp

# --- 
# Ambiguous: Temp + Vomit + Affective_touch + Sex_arousal + Taste
# Discard: Tickle + Blood_Sugar
"

m8 <- "
# ----
Sustenance =~ Hungry + Thirsty
Elimination =~ Urinate + Defecate
Skin =~ Itch + Bruise
Nociception =~ Muscles + Pain
Affect =~ Breathing + Heart
# ----
Expulsion_Sudden =~ Cough + Sneeze 
Expulsion_Gastric =~ Wind + Burp
" 
# Ambiguous: Temp + Vomit + Affective_touch + Sex_arousal + Taste
# Discard: Tickle + Blood_Sugar

m1_all <- lavaan::cfa(m1, data = data_all)
m7_all <- lavaan::cfa(m7, data = data_all)
m8_all <- lavaan::cfa(m8, data = data_all)

anova(m1_all, m7_all, m8_all) |> 
  parameters::parameters() |> 
  display()
```

#### Higher Order Factors

```{r}
m8h1 <- paste0(m8, "
              Interoception =~ Sustenance + Elimination + Skin + Nociception + Affect + Expulsion_Sudden + Expulsion_Gastric
              ")

m8h3 <- paste0(m8, "
              Valenced =~ Nociception + Affect + Skin
              Expulsion =~ Expulsion_Sudden + Expulsion_Gastric 
              Homeostasis =~ Sustenance + Elimination
              ")

m8h1_all <- lavaan::cfa(m8h1, data = data_all)
m8h3_all <- lavaan::cfa(m8h3, data = data_all)
anova(m8_all, m8h1_all, m8h3_all) |> 
  parameters::parameters() |> 
  display()
```

#### Model Performance

```{r}
#| warning: false

rbind(
  performance::performance(m8_all, metrics=c("Chi2", "RMSEA", "CFI", "SRMR")) |> 
    mutate(Sample = "All"),
  performance::performance(update(m8_all, data=data1), metrics=c("Chi2", "RMSEA", "CFI", "SRMR")) |> 
    mutate(Sample = "Sample 1"),
  performance::performance(update(m8_all, data=data2), metrics=c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 2"),
  performance::performance(update(m8_all, data=data3), metrics=c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 3"),
  performance::performance(update(m8_all, data=data4), metrics=c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 4"),
  performance::performance(update(m8_all, data=data5), metrics=c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 5"),
  performance::performance(update(m8_all, data=data6), metrics=c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 6"),
  performance::performance(lavaan::cfa(str_remove(m8, fixed("\nExpulsion_Sudden =~ Cough + Sneeze")), data=data7), metrics=c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 7")
) |> 
  display()
```

No evidence for higher order factors. The unidimensionality might be a byproduct of the similarities in phrasing of the items.

## Study 2 - Correlates

```{r}
#| warning: false

df1 <- cbind(df1, as.data.frame(predict(lavaan::cfa(m8, data = data1)))) |> 
  mutate(Sample = "Sample 1")
df2 <- cbind(df2, as.data.frame(predict(lavaan::cfa(m8, data = data2)))) |> 
  mutate(Sample = "Sample 2")
df3 <- cbind(df3, as.data.frame(predict(lavaan::cfa(m8, data = data3)))) |> 
  mutate(Sample = "Sample 3")
df4 <- cbind(df4, as.data.frame(predict(lavaan::cfa(m8, data = data4)))) |> 
  mutate(Sample = "Sample 4")
df5 <- cbind(df5, as.data.frame(predict(lavaan::cfa(m8, data = data5)))) |> 
  mutate(Sample = "Sample 5")
df6 <- cbind(df6, as.data.frame(predict(lavaan::cfa(m8, data = data6)))) |> 
  mutate(Sample = "Sample 6")
df7 <- cbind(df7, as.data.frame(predict(lavaan::cfa(str_remove(m8, fixed("\nExpulsion_Sudden =~ Cough + Sneeze")), data = data7)))) |> 
  mutate(Sample = "Sample 7", Expulsion_Sudden = NA)

vars <- names(as.data.frame(predict(m8_all)))
```

### Demographics

```{r}
df_demo <- rbind(
  select(df1, Sample, Age, Gender, all_of(vars)),
  select(df2, Sample, Age, Gender, all_of(vars)),
  select(df3, Sample, Age, Gender, all_of(vars)),
  select(df4, Sample, Age, Gender, all_of(vars)),
  select(df5, Sample, Age, Gender, all_of(vars)),
  select(df6, Sample, Age, Gender, all_of(vars)),
  select(df7, Sample, Age, Gender, all_of(vars))
)

make_lm_data <- function(df, vars, predictor="Age") {
  dat <- data.frame()
  for(resp in vars) {
      if(length(unique(df$Sample)) == 1) {
        m <- lm(as.formula(paste0(resp, " ~ ", predictor)), data = df)
        param <- parameters::parameters(m)[2, ] |> 
          mutate(Sample = unique(df$Sample)) |> 
          select(-Parameter)
      } else {
        m <- lm(as.formula(paste0(resp, " ~ Sample / ", predictor)), data = df)
        param <- parameters::parameters(m) |> 
          filter(str_detect(Parameter, ":")) |> 
          separate(Parameter, c("Predictor", "Sample"), sep = "Sample ") |> 
          mutate(Sample = paste0("Sample ", str_remove(Sample, fixed(paste0(":", predictor)))))
      }

      dat <- param |> 
        mutate(Predictor = predictor,
               Repsonse = resp,
               t_low = CI_low / SE,
               t_high = CI_high / SE) |> 
        rbind(dat)
  }
  dat
}

data <- rbind(
  make_lm_data(df_demo, vars, predictor="Age"),
  make_lm_data(filter(df_demo, Gender %in% c("Male", "Female")), vars, predictor="Gender")
)
```

### BPQ

```{r}
x <- intersect(names(df3)[str_detect(names(df3), "BPQ")], 
               names(df4)[str_starts(names(df4), "BPQ")])

df_bpq <- rbind(
  select(df3, Sample, all_of(x), all_of(vars)),
  select(df4, Sample, all_of(x), all_of(vars))
)

# estimate_density(select(df_bpq, Sample, starts_with("BPQ")), at="Sample")  |> 
#   plot() +
#   facet_grid(~Sample)

plot(parameters::n_components(select(df_bpq, starts_with("BPQ"))))
df_bpq$BPQ_PCA1 <- predict(parameters::principal_components(select(df_bpq, starts_with("BPQ")), n=1))$PC1

data <- rbind(data,
              make_lm_data(df_bpq, vars, predictor="BPQ_PCA1"))

# Why opposite effects?
```

### TAS

```{r}
x <- intersect(names(df3)[str_detect(names(df3), "TAS")], 
               names(df5)[str_starts(names(df5), "TAS")])
```

### PHQ

```{r}
names(df5)[str_detect(names(df5), "PHQ")]

data <- rbind(data,
              make_lm_data(df5, vars, predictor="PHQ9_sum"),
              make_lm_data(df5, vars, predictor="PHQ15_sum"))
```

### BDI

```{r}
# names(df1)
# names(df2)
# names(df3)  # TAS, BPQ
# names(df4)  # BPQ
# names(df5)  # PHQ, TAS
# names(df6)  # BDI, STAI
# names(df7)  # MAIA

data <- rbind(data,
              make_lm_data(df6, vars, predictor="BDI2_Total"))
```

### Summary

```{r}
#| fig-height: 20

data |> 
  mutate(sig = ifelse(p < .01, "p < .01", ifelse(p < .001, "p < .001", "N.S."))) |> 
  ggplot(aes(x=Predictor)) +
  geom_hline(yintercept=0, linetype="dashed") +
  geom_pointrange(aes(y=t, ymin=t_low, ymax=t_high, color=sig), position=position_dodge2(width=0.3)) +
  facet_grid(~Repsonse, scales = "free_y") +
  coord_flip()
```

## Discussion

Benefits of the IAS:
- Straightforward and sensation-centered items

Recommendations:
- Remove Itch (redundant + issue in Chinese)
- Use analog scales

Limitations:
- Not much clear theorethical or empirical structure (small grouping of items)
- Limited variability (clear mode at 4/5)
- Ambiguous items which grouping depends on the context (and its awareness)
  - E.g., heart beating fast, vomit when nauseaous
- Few items for some modalities (e.g., 1 for heart)
- Positive phrasing: benefits but also might exacerbate positivity bias (and thus unidimensionality)

Need for context-specific items (cross-modal when possible, i.E., cardioception, respiroception, etc.).

New Scale: Multimodal Interoceptive Sensitivity Scale (MISS)