### 2024-02-20 analysis/analysis.qmd
update stats

---
title: "Revised Interoceptive Accuracy Scale (IAS-R)"
subtitle: "Analysis"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    fig-dpi: 300
    fig-width: 11.33
execute: 
  cache: true
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(lavaan)
library(ggraph)
library(tidySEM)
library(EGAnet)
```

## Study 1 - Reanalysis of Murphy et al. (2020)

Reanalyze their data to confirm the factor structure of the IAS.

### Data Preparation

```{r}
# https://osf.io/3m5nh
df1 <- haven::read_sav("../data/Murphy2020/Study 1.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))

# https://osf.io/3m5nh
df2 <- haven::read_sav("../data/Murphy2020/Study 6 IAS.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))

# Campos (2022) - Study 1 (https://osf.io/j6ef3)
df3 <- haven::read_sav("../data/Campos2022/Dataset_Test.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Sex == 1, "Male", ifelse(Sex == 0, "Female", "Other")))) |>
  rename(
    Heart = IAS1, Hungry = IAS2, Breathing = IAS3, Thirsty = IAS4,
    Urinate = IAS5, Defecate = IAS6, Taste = IAS7, Vomit = IAS8, Sneeze = IAS9,
    Cough = IAS10, Temp = IAS11, Sex_arousal = IAS12, Wind = IAS13, Burp = IAS14,
    Muscles = IAS15, Bruise = IAS16, Pain = IAS17, Blood_Sugar = IAS18,
    Affective_touch = IAS19, Tickle = IAS20, Itch = IAS21
  )

# https://osf.io/3eztd
df4 <- haven::read_sav("../data/Lin2023/Study 1 & 3.sav") |>
  select(-sex) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(sex_dummy == 1, "Male", ifelse(sex_dummy == 0, "Female", "Other")))) |>
  rename(
    Age = age,
    Heart = Heart, Hungry = HUNGR, Breathing = BREAT, Thirsty = Thirs,
    Urinate = URINA, Defecate = Defec, Taste = TASTE, Vomit = VOMIT, Sneeze = Sneez,
    Cough = COUGH, Temp = TEMPE, Sex_arousal = SEXAR, Wind = WIND, Burp = Burp,
    Muscles = MUSCL, Bruise = Bruis, Pain = PAIN, Blood_Sugar = BloSu,
    Affective_touch = Touch, Itch = ITCH
  ) # No tickle because same Chinese character

# https://osf.io/3eztd
df5 <- haven::read_sav("../data/Lin2023/Study 2.sav") |>
  mutate(Gender = as.character(ifelse(Sex == "男", "Male", ifelse(Sex == "女", "Female", "Other")))) |>
  rename(
    Heart = Heart, Hungry = HUNGR, Breathing = BREAT, Thirsty = Thirs,
    Urinate = URINA, Defecate = Defec, Taste = TASTE, Vomit = VOMIT, Sneeze = Sneez,
    Cough = COUGH, Temp = TEMPE, Sex_arousal = SEXAR, Wind = WIND, Burp = Burp,
    Muscles = MUSCL, Bruise = Bruis, Pain = PAIN, Blood_Sugar = BloSu,
    Affective_touch = Touch, Itch = ITCH
  ) # No tickle because same Chinese character

df6 <- read.csv("https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data.csv") |>
  rename(
    Heart = IAS_1, Hungry = IAS_2, Breathing = IAS_3, Thirsty = IAS_4,
    Urinate = IAS_5, Defecate = IAS_6, Taste = IAS_7, Vomit = IAS_8, Sneeze = IAS_9,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )

df7 <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |>
  rename(
    Gender = Sex, Heart = Item_IAS_Interoception_1, Hungry = Item_IAS_Interoception_2,
    Breathing = Item_IAS_Interoception_3, Thirsty = Item_IAS_Interoception_4,
    Temp = Item_IAS_Interoception_5, Sex_arousal = Item_IAS_Interoception_6,
    Urinate = Item_IAS_Elimination_1, Defecate = Item_IAS_Elimination_2,
    Vomit = Item_IAS_Elimination_3, Wind = Item_IAS_Expulsion_1,
    Burp = Item_IAS_Expulsion_2, Sneeze = Item_IAS_Expulsion_3,
    Muscles = Item_IAS_Nociception_1, Bruise = Item_IAS_Nociception_2,
    Pain = Item_IAS_Nociception_3, Affective_touch = Item_IAS_Skin_1,
    Tickle = Item_IAS_Skin_2, Itch = Item_IAS_Skin_3
  ) |>
  filter(!is.na(Urinate))
```

### Participants

- **Sample 1**: Data from Murphy's (2020) [study 1](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769), downloaded from OSF, included `r report::report_participants(df1, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 2**: Data from Murphy's (2020) [study 6](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769), downloaded from OSF, included `r report::report_participants(df2, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 3**: Data from Campos' (2022) [study 1](https://osf.io/j6ef3), downloaded from OSF, included `r report::report_participants(df3, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 4**: Data from Lin's (2023) [study 1 and 3](#ttps://osf.io/3eztd), downloaded from OSF, included `r report::report_participants(df4, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 5**: Data from Lin's (2023) [study 2](#ttps://osf.io/3eztd), downloaded from OSF, included `r report::report_participants(df5, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 6**: New dataset from [Makowski (under preparation)](https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data.csv), included `r report::report_participants(df6, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 7**: New dataset from [Makowski (under preparation)](https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv), included `r report::report_participants(df7, age = "Age", sex = NA, gender = "Gender")`.

Total N = `r nrow(df1) + nrow(df2) + nrow(df3) + nrow(df4) + nrow(df5) + nrow(df6) + nrow(df7)`.

### Descriptive

#### Distribution

```{r}
vars <- c(
  "Heart", "Hungry", "Breathing", "Thirsty", "Urinate", "Defecate", "Taste", "Vomit", "Sneeze", "Cough", "Temp",
  "Sex_arousal", "Wind", "Burp", "Muscles", "Bruise", "Pain", "Blood_Sugar", "Affective_touch", "Tickle", "Itch"
)

dens1 <- estimate_density(select(df1, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 1")
dens2 <- estimate_density(select(df2, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 2")
dens3 <- estimate_density(select(df3, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 3")
dens4 <- estimate_density(select(df4, all_of(setdiff(vars, "Tickle"))), method = "kernSmooth") |>
  mutate(Sample = "Sample 4")
dens5 <- estimate_density(select(df5, all_of(setdiff(vars, "Tickle"))), method = "kernSmooth") |>
  mutate(Sample = "Sample 5")
dens6 <- estimate_density(select(df6, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 6")
dens7 <- estimate_density(select(df7, all_of(setdiff(vars, c("Taste", "Cough", "Blood_Sugar")))), method = "kernSmooth") |>
  mutate(Sample = "Sample 7")

rbind(dens1, dens2, dens3, dens4, dens5, dens6, dens7) |>
  ggplot(aes(x = x, y = y, color = Parameter)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Distribution of Responses", x = "Response", y = "Density", color = "Item") +
  facet_wrap(~Sample, scales = "free", nrow = 4)
```

```{r}
data1 <- normalize(select(df1, all_of(dens1$Parameter)), range = c(1, 5))
data2 <- normalize(select(df2, all_of(dens2$Parameter)), range = c(1, 5))
data3 <- normalize(select(df3, all_of(dens3$Parameter)), range = c(1, 5))
data4 <- normalize(select(df4, all_of(dens4$Parameter)), range = c(1, 5))
data5 <- normalize(select(df5, all_of(dens5$Parameter)), range = c(1, 5))
data6 <- select(df6, all_of(dens6$Parameter))
data7 <- select(df7, all_of(dens7$Parameter))

data_all <- rbind(
  data1, data2, data3,
  mutate(data4, Tickle = NA), mutate(data5, Tickle = NA),
  data6, mutate(data7, Taste = NA, Cough = NA, Blood_Sugar = NA)
)
```

#### Correlations

An overall positive intercorrelation pattern, with no clear structure emerging.

::: {.panel-tabset}

##### All data

```{r}
make_correlation <- function(df) {
  correlation::correlation(df, redundant = TRUE) |>
    correlation::cor_sort() |>
    correlation::cor_lower() |>
    mutate(val = paste0(insight::format_value(r), format_p(p, stars_only = TRUE))) |>
    mutate(Parameter2 = fct_rev(Parameter2)) |>
    ggplot(aes(x = Parameter1, y = Parameter2)) +
    geom_tile(aes(fill = r), color = "white") +
    geom_text(aes(label = val), size = 3) +
    labs(title = "Correlation Matrix") +
    scale_fill_metro_c(limit = c(0, 0.75), guide = guide_colourbar(ticks = FALSE)) +
    theme_minimal() +
    theme(
      legend.title = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

make_correlation(data_all)
```

##### Sample 1

```{r}
make_correlation(data1)
```

##### Sample 2

```{r}
make_correlation(data2)
```

##### Sample 3

```{r}
make_correlation(data3)
```

##### Sample 4

```{r}
make_correlation(data4)
```

##### Sample 5

```{r}
make_correlation(data5)
```

##### Sample 6

```{r}
make_correlation(data6)
```

##### Sample 7

```{r}
make_correlation(data7)
```

:::

### Graph Analysis

Exploratory Graph Analysis (EGA) is a recently developed framework for psychometric assessment, that can be used to estimate the number of dimensions in questionnaire data using network estimation methods and community detection algorithms, and assess the stability of dimensions and items using bootstrapping.
See https://r-ega.net/articles/ega.html for details.

#### Unique Variable Analysis (UVA)

Unique Variable Analysis (Christensen, Garrido, & Golino, 2023) uses the weighted topological overlap measure (Nowick et al., 2009) on an estimated network. Values greater than 0.25 are determined to have considerable local dependence (i.e., redundancy) that should be handled (variables with the highest maximum weighted topological overlap to all other variables (other than the one it is redundant with) should be removed).

::: {.panel-tabset}

##### All data

```{r}
uva <- EGAnet::UVA(data = data_all, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 1

```{r}
uva <- EGAnet::UVA(data = data1, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 2

```{r}
uva <- EGAnet::UVA(data = data2, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 3

```{r}
uva <- EGAnet::UVA(data = data3, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 4

```{r}
uva <- EGAnet::UVA(data = data4, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 5

```{r}
uva <- EGAnet::UVA(data = data5, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 6

```{r}
uva <- EGAnet::UVA(data = data6, cut.off = 0.3)
uva
uva$keep_remove
```

##### Sample 7

```{r}
uva <- EGAnet::UVA(data = data7, cut.off = 0.3)
uva
uva$keep_remove
```

:::

#### Networks

::: {.panel-tabset}

##### All data

```{r ega1}
#| fig-height: 20
#| fig-width: 12

make_egas <- function(data) {
  egas <- list()
  for (model in c("glasso", "TMFG")) {
    for (algo in c("walktrap", "louvain")) {
      for (type in c("ega", "ega.fit")) { # "hierega"
        if (type == "ega.fit" & algo == "louvain") next # Too slow
        egas[[paste0(model, "_", algo, "_", type)]] <- EGAnet::bootEGA(
          data = data,
          seed = 123,
          model = model,
          algorithm = algo,
          EGA.type = type,
          type = "resampling",
          plot.itemStability = FALSE,
          verbose = FALSE
        )
      }
    }
  }

  p <- EGAnet::compare.EGA.plots(
    egas$glasso_walktrap_ega, egas$glasso_walktrap_ega.fit,
    egas$glasso_louvain_ega, egas$TMFG_louvain_ega,
    egas$TMFG_walktrap_ega, egas$TMFG_walktrap_ega.fit,
    labels = c(
      "glasso_walktrap_ega", "glasso_walktrap_ega.fit",
      "glasso_louvain_ega", "TMFG_louvain_ega",
      "TMFG_walktrap_ega", "TMFG_walktrap_ega.fit"
    ),
    rows = 3,
    plot.all = FALSE
  )$all
  list(egas = egas, p = p)
}

egas_all <- make_egas(data_all)
egas_all$p
```

##### Sample 1

```{r}
egas1 <- make_egas(data1)
egas1$p
```

##### Sample 2

```{r}
egas2 <- make_egas(data2)
egas2$p
```

##### Sample 3

```{r}
egas3 <- make_egas(data3)
egas3$p
```

##### Sample 4

```{r}
egas4 <- make_egas(data4)
egas4$p
```

##### Sample 5

```{r}
egas5 <- make_egas(data5)
egas5$p
```

##### Sample 6

```{r}
egas6 <- make_egas(data6)
egas6$p
```

##### Sample 7

```{r}
egas7 <- make_egas(data7)
egas7$p
```

:::

#### Structure Stability

::: {.panel-tabset}

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas_all$egas, plot), nrow = 3)
```

:::

#### Scores

```{r}
ega <- egas_all$egas$glasso_louvain_ega$EGA
plot(ega)

ega_scores_1 <- EGAnet::net.scores(data1, ega)$scores$std.scores |>
  as.data.frame() |>
  setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))

ega_scores_2 <- EGAnet::net.scores(data2, ega)$scores$std.scores |>
  as.data.frame() |>
  setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))

ega_scores_3 <- EGAnet::net.scores(data3, ega)$scores$std.scores |>
  as.data.frame() |>
  setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))

ega_scores_4 <- EGAnet::net.scores(data4, ega)$scores$std.scores |>
  as.data.frame() |>
  setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))

ega_scores_5 <- EGAnet::net.scores(data5, ega)$scores$std.scores |>
  as.data.frame() |>
  setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))

ega_scores_6 <- EGAnet::net.scores(data6, ega)$scores$std.scores |>
  as.data.frame() |>
  setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))

ega_scores_7 <- EGAnet::net.scores(data7, ega)$scores$std.scores |>
  as.data.frame() |>
  setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))
```

### Factor Analysis

#### How Many Factors

::: {.panel-tabset}

##### All data

```{r}
n <- parameters::n_factors(data_all, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4, 5), ], row.names = FALSE)
```

##### Sample 1

```{r}
n <- parameters::n_factors(data1, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 2

```{r}
n <- parameters::n_factors(data2, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 3

```{r}
n <- parameters::n_factors(data3, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 4

```{r}
n <- parameters::n_factors(data4, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 5

```{r}
n <- parameters::n_factors(data5, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 6

```{r}
n <- parameters::n_factors(data6, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

##### Sample 7

```{r}
n <- parameters::n_factors(data7, n_max = 12)

plot(n)

knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 4), ], row.names = FALSE)
```

:::

#### EFA

::: {.panel-tabset}

##### All data

```{r}
#| fig-height: 10

efa5 <- parameters::factor_analysis(data_all, n = 5, rotation = "oblimin", sort = TRUE)
plot(efa5)

display(efa5)
```

##### Sample 1

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data2, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

##### Sample 2

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data2, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

##### Sample 3

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data3, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

##### Sample 4

```{r}
#| fig-height: 10

efa5 <- parameters::factor_analysis(data4, n = 5, rotation = "oblimin", sort = TRUE)
plot(efa5)

display(efa5)
```

##### Sample 5

```{r}
#| fig-height: 10

efa5 <- parameters::factor_analysis(data5, n = 5, rotation = "oblimin", sort = TRUE)
plot(efa5)

display(efa5)
```

##### Sample 6

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data6, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

##### Sample 7

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data7, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

:::

### CFA

#### Structure

::: {.panel-tabset}

##### All data

```{r}
# Ambiguous: Temp + Vomit + Affective_touch + Sex_arousal + Taste
# Discard: Tickle + Blood_Sugar

m1 <- "
Interoception =~ Hungry + Thirsty + Urinate + Defecate + Itch + Bruise + Muscles + Pain + Breathing + Heart + Cough + Sneeze + Wind + Burp
"

m4 <- "
Sustenance =~ Hungry + Thirsty + Urinate + Defecate + Muscles + Pain
Skin =~ Itch + Bruise
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m5 <- "
Sustenance =~ Hungry + Thirsty + Pain
UrinateDefecate =~ Urinate + Defecate
Skin =~ Itch + Bruise + Muscles
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m6 <- "
HungryThirsty =~ Hungry + Thirsty
UrinateDefecate =~ Urinate + Defecate
ItchBruise =~ Itch + Bruise
MusclesPain =~ Muscles + Pain
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m7 <- "
HungryThirsty =~ Hungry + Thirsty
UrinateDefecate =~ Urinate + Defecate
ItchBruise =~ Itch + Bruise
MusclesPain =~ Muscles + Pain
HeartBreathing =~ Breathing + Heart
CoughSneeze =~ Cough + Sneeze
WindBurp =~ Wind + Burp
"

m1_all <- lavaan::cfa(m1, data = data_all)
m4_all <- lavaan::cfa(m4, data = data_all)
m5_all <- lavaan::cfa(m5, data = data_all)
m6_all <- lavaan::cfa(m6, data = data_all)
m7_all <- lavaan::cfa(m7, data = data_all)

anova(m1_all, m4_all, m5_all, m6_all, m7_all) |>
  parameters::parameters() |>
  display()

parameters::parameters(m7_all, standardize=TRUE) |>
  display()
```

##### Sample 1

```{r}
anova(
  lavaan::cfa(m1, data = data1),
  lavaan::cfa(m4, data = data1),
  lavaan::cfa(m5, data = data1),
  lavaan::cfa(m6, data = data1),
  lavaan::cfa(m7, data = data1)
  ) |>
  parameters::parameters() |>
  display()
```

##### Sample 2

```{r}
anova(
  lavaan::cfa(m1, data = data2),
  lavaan::cfa(m4, data = data2),
  lavaan::cfa(m5, data = data2),
  lavaan::cfa(m6, data = data2),
  lavaan::cfa(m7, data = data2)
  ) |>
  parameters::parameters() |>
  display()
```

##### Sample 3

```{r}
anova(
  lavaan::cfa(m1, data = data3),
  lavaan::cfa(m4, data = data3),
  lavaan::cfa(m5, data = data3),
  lavaan::cfa(m6, data = data3),
  lavaan::cfa(m7, data = data3)
  ) |>
  parameters::parameters() |>
  display()
```

##### Sample 4

```{r}
anova(
  lavaan::cfa(m1, data = data4),
  lavaan::cfa(m4, data = data4),
  lavaan::cfa(m5, data = data4),
  lavaan::cfa(m6, data = data4),
  lavaan::cfa(m7, data = data4)
  ) |>
  parameters::parameters() |>
  display()
```

##### Sample 5

```{r}
anova(
  lavaan::cfa(m1, data = data5),
  lavaan::cfa(m4, data = data5),
  lavaan::cfa(m5, data = data5),
  lavaan::cfa(m6, data = data5),
  lavaan::cfa(m7, data = data5)
  ) |>
  parameters::parameters() |>
  display()
```

##### Sample 6

```{r}
anova(
  lavaan::cfa(m1, data = data6),
  lavaan::cfa(m4, data = data6),
  lavaan::cfa(m5, data = data6),
  lavaan::cfa(m6, data = data6),
  lavaan::cfa(m7, data = data6)
  ) |>
  parameters::parameters() |>
  display()
```

##### Sample 7

```{r}
anova(
  lavaan::cfa(str_remove(m5, fixed("Cough + ")), data = data7),
  lavaan::cfa(str_remove(m1, fixed("Cough + ")), data = data7),
  lavaan::cfa(str_remove(m4, fixed("Cough + ")), data = data7),
  lavaan::cfa(str_remove(m6, fixed("Cough + ")), data = data7),
  lavaan::cfa(str_remove(m7, fixed("\nCoughSneeze =~ Cough + Sneeze")), data = data7)
  ) |>
  parameters::parameters() |>
  display()
```

:::

#### Higher Order Factors

::: {.panel-tabset}

##### All data

```{r}
m7h1 <- paste0(m7, "
              Interoception =~ HungryThirsty + UrinateDefecate + ItchBruise + MusclesPain + HeartBreathing + CoughSneeze + WindBurp
              ")

m7h3 <- paste0(m7, "
              Valenced =~ MusclesPain + HeartBreathing + ItchBruise
              Expulsion =~ CoughSneeze + WindBurp
              Homeostasis =~ HungryThirsty + UrinateDefecate
              ")

m7h1_all <- lavaan::cfa(m7h1, data = data_all)
m7h3_all <- lavaan::cfa(m7h3, data = data_all)
anova(m7_all, m7h1_all, m7h3_all) |>
  parameters::parameters() |>
  display()
```

##### Sample 1

```{r}
anova(
  lavaan::cfa(m7, data = data1), 
  lavaan::cfa(m7h1, data = data1), 
  lavaan::cfa(m7h3, data = data1)) |>
  parameters::parameters() |>
  display()
```

##### Sample 2

```{r}
anova(
  lavaan::cfa(m7, data = data2), 
  lavaan::cfa(m7h1, data = data2), 
  lavaan::cfa(m7h3, data = data2)) |>
  parameters::parameters() |>
  display()
```

##### Sample 3

```{r}
anova(
  lavaan::cfa(m7, data = data3), 
  lavaan::cfa(m7h1, data = data3), 
  lavaan::cfa(m7h3, data = data3)) |>
  parameters::parameters() |>
  display()
```

##### Sample 4

```{r}
anova(
  lavaan::cfa(m7, data = data4), 
  lavaan::cfa(m7h1, data = data4), 
  lavaan::cfa(m7h3, data = data4)) |>
  parameters::parameters() |>
  display()
```

##### Sample 5

```{r}
anova(
  lavaan::cfa(m7, data = data5), 
  lavaan::cfa(m7h1, data = data5), 
  lavaan::cfa(m7h3, data = data5)) |>
  parameters::parameters() |>
  display()
```

##### Sample 6

```{r}
anova(
  lavaan::cfa(m7, data = data6), 
  lavaan::cfa(m7h1, data = data6), 
  lavaan::cfa(m7h3, data = data6)) |>
  parameters::parameters() |>
  display()
```

##### Sample 7

None of the model can be properly estimated (variances are negative).

```{r}
NA
# anova(
#   lavaan::cfa(str_remove(m7, fixed("\nCoughSneeze =~ Cough + Sneeze")), data = data7), 
#   lavaan::cfa(str_remove(str_remove(m7h1, fixed("\nCoughSneeze =~ Cough + Sneeze")), fixed("CoughSneeze +")), data = data7), 
#   lavaan::cfa(str_remove(str_remove(m7h3, fixed("\nCoughSneeze =~ Cough + Sneeze")), fixed("CoughSneeze +")), data = data7)) |>
#   parameters::parameters() |>
#   display()
```

:::

#### Model Performance

```{r}
#| warning: false

rbind(
  performance::performance(m7_all, metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "All"),
  performance::performance(update(m7_all, data = data1), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 1"),
  performance::performance(update(m7_all, data = data2), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 2"),
  performance::performance(update(m7_all, data = data3), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 3"),
  performance::performance(update(m7_all, data = data4), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 4"),
  performance::performance(update(m7_all, data = data5), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 5"),
  performance::performance(update(m7_all, data = data6), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 6"),
  performance::performance(lavaan::cfa(str_remove(m7, fixed("\nCoughSneeze =~ Cough + Sneeze")), data = data7), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 7")
) |>
  display()
```

No evidence for higher order factors.

### Scores

Note the usage of descriptive factor names relating directly to the items to avoid abstraction.

```{r}
#| warning: false

scores <- list(
  sample1 = cbind(ega_scores_1, as.data.frame(predict(lavaan::cfa(m7, data = data1)))) |>
    mutate(Sample = "Sample 1"),
  sample2 = cbind(ega_scores_2, as.data.frame(predict(lavaan::cfa(m7, data = data2)))) |>
    mutate(Sample = "Sample 2"),
  sample3 = cbind(ega_scores_3, as.data.frame(predict(lavaan::cfa(m7, data = data3)))) |>
    mutate(Sample = "Sample 3"),
  sample4 = cbind(ega_scores_4, as.data.frame(predict(lavaan::cfa(m7, data = data4)))) |>
    mutate(Sample = "Sample 4"),
  sample5 = cbind(ega_scores_5, as.data.frame(predict(lavaan::cfa(m7, data = data5)))) |>
    mutate(Sample = "Sample 5"),
  sample6 = cbind(ega_scores_6, as.data.frame(predict(lavaan::cfa(m7, data = data6)))) |>
    mutate(Sample = "Sample 6"),
  sample7 = cbind(ega_scores_7, as.data.frame(predict(lavaan::cfa(str_remove(m7, fixed("\nCoughSneeze =~ Cough + Sneeze")), data = data7)))) |>
    mutate(Sample = "Sample 7", CoughSneeze = NA)
)

save(scores, file = "../data/scores.RData")
```

## Study 2 - Correlates

```{r}
#| warning: false

load("../data/scores.RData")

df1 <- cbind(df1, scores$sample1)
df2 <- cbind(df2, scores$sample2)
df3 <- cbind(df3, scores$sample3)
df4 <- cbind(df4, scores$sample4)
df5 <- cbind(df5, scores$sample5)
df6 <- cbind(df6, scores$sample6)
df7 <- cbind(df7, scores$sample7)

vars_intero <- names(select(scores$sample1, -Sample))
vars_cfa <- names(select(scores$sample1, -Sample, -starts_with("EGA")))
vars_cfa7 <- vars_cfa[vars_cfa != "CoughSneeze"]

# names(df1)
# names(df2)
# names(df3)  # TAS, BPQ
# names(df4)  # BPQ
# names(df5)  # PHQ, TAS
# names(df6)  # BDI, STAI
# names(df7)  # MAIA, PHQ4, PI, LIE, IPIP, PID, SPQ
```

### EGA and CFA

```{r}
#| warning: false

df_scores <- rbind(
  select(df1, all_of(vars_intero)),
  select(df2, all_of(vars_intero)),
  select(df3, all_of(vars_intero)),
  select(df4, all_of(vars_intero)),
  select(df5, all_of(vars_intero)),
  select(df6, all_of(vars_intero)),
  select(df7, all_of(vars_intero))
)

correlation::correlation(df_scores, redundant = TRUE) |>
    correlation::cor_sort() |>
    correlation::cor_lower() |>
    mutate(val = paste0(insight::format_value(r), format_p(p, stars_only = TRUE))) |>
    mutate(Parameter2 = fct_rev(Parameter2)) |>
    ggplot(aes(x = Parameter1, y = Parameter2)) +
    geom_tile(aes(fill = r), color = "white") +
    geom_text(aes(label = val), size = 3) +
    labs(title = "Correlation Matrix") +
    scale_fill_metro_c(limit = c(0, 1), guide = guide_colourbar(ticks = FALSE)) +
    theme_minimal() +
    theme(
      legend.title = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
```

### Demographics

```{r}
make_diff <- function(df, vars, component=NA) {
  dat <- data.frame()
  for (resp in vars) {
    f <- as.formula(paste0(resp, "~ Gender"))
    d <- filter(df, Gender %in% c("Male", "Female")) |> 
      mutate(Gender = fct_relevel(Gender, "Male", "Female"))
    dat <- effectsize::cohens_d(f, data=d) |> 
      mutate(var_Correlate = "Gender", var_Interoception = resp, 
             p=t.test(f, data=d)$p.val) |> 
      as.data.frame() |> 
      rbind(dat)
  }

  dat$r <- effectsize::d_to_r(dat$Cohens_d, n1=sum(df$Gender=="Male"), n2=sum(df$Gender=="Female"))
  dat$CI_low <- effectsize::d_to_r(dat$CI_low, n1=sum(df$Gender=="Male"), n2=sum(df$Gender=="Female"))
  dat$CI_high <- effectsize::d_to_r(dat$CI_high, n1=sum(df$Gender=="Male"), n2=sum(df$Gender=="Female"))
  dat$Sample <- unique(df$Sample)
  dat$Component <- component
  select(dat, -Cohens_d)
}

data <- rbind(
  make_diff(df1, vars_cfa, component="Demographic"),
  make_diff(df2, vars_cfa, component="Demographic"),
  make_diff(df3, vars_cfa, component="Demographic"),
  make_diff(df4, vars_cfa, component="Demographic"),
  make_diff(df5, vars_cfa, component="Demographic"),
  make_diff(df6, vars_cfa, component="Demographic"),
  make_diff(df7, vars_cfa7, component="Demographic")
) |> 
  mutate(var_Correlate = "Gender (Male vs. Female)")
```

```{r}
make_cor <- function(df, vars, with = "Age", component=NA) {
  dat <- data.frame() 
  for (v in vars) {
    dat <- correlation::cor_test(df, v, with) |> 
      rename(var_Interoception = Parameter1, var_Correlate = Parameter2) |> 
      as.data.frame() |> 
      rbind(dat)
  }

  dat$Sample <- unique(df$Sample)
  dat$Component <- component
  select(dat, -df_error, -t, -Method, -n_Obs)
}

data <- rbind(
  data,
  make_cor(df1, vars_cfa, with = "Age", component="Demographic"),
  make_cor(df2, vars_cfa, with = "Age", component="Demographic"),
  make_cor(df3, vars_cfa, with = "Age", component="Demographic"),
  make_cor(df4, vars_cfa, with = "Age", component="Demographic"),
  make_cor(df5, vars_cfa, with = "Age", component="Demographic"),
  make_cor(df6, vars_cfa, with = "Age", component="Demographic"),
  make_cor(df7, vars_cfa7, with = "Age", component="Demographic")
) 
```

### Body

```{r}
# Reverse (TODO: Ana confirm why)
df3$BPQ_BodyAwareness <- -1*normalize(df3$BPQ_BodyAwareness)
df3$BPQ_AutonomicReactivity <- -1*normalize(df3$BPQ_AutonomicReactivity)

data <- rbind(
  data,
  make_cor(df3, vars_cfa, with = "BPQ_BodyAwareness", component="Body") |> 
    mutate(var_Correlate = "Body Awareness"),
  make_cor(df3, vars_cfa, with = "BPQ_AutonomicReactivity", component="Body") |> 
    mutate(var_Correlate = "Autonomic Reactivity"),
  make_cor(df4, vars_cfa, with = "BPQ_A", component="Body") |> 
    mutate(var_Correlate = "Body Awareness")
)
```

```{r}
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "MAIA_AttentionRegulation", component="MAIA")  |> 
    mutate(var_Correlate = "Attention Regulation"),
  make_cor(df7, vars_cfa7, with = "MAIA_BodyListening", component="MAIA") |> 
    mutate(var_Correlate = "Body Listening"),
  make_cor(df7, vars_cfa7, with = "MAIA_EmotionalAwareness", component="MAIA") |> 
    mutate(var_Correlate = "Emotional Awareness"),
  make_cor(df7, vars_cfa7, with = "MAIA_NotDistracting", component="MAIA") |> 
    mutate(var_Correlate = "Not Distracting"),
  make_cor(df7, vars_cfa7, with = "MAIA_Noticing", component="MAIA") |> 
    mutate(var_Correlate = "Noticing"),
  make_cor(df7, vars_cfa7, with = "MAIA_NotWorrying", component="MAIA") |> 
    mutate(var_Correlate = "Not Worrying"),
  make_cor(df7, vars_cfa7, with = "MAIA_SelfRegulation", component="MAIA") |> 
    mutate(var_Correlate = "Self Regulation"),
  make_cor(df7, vars_cfa7, with = "MAIA_Trusting", component="MAIA") |> 
    mutate(var_Correlate = "Trusting")
)
```

```{r}
# names(df3)[str_detect(names(df3), "TAS")]
# names(df5)[str_starts(names(df5), "TAS")]

data <- rbind(
  data,
  make_cor(df3, vars_cfa, with = "TAS_IdentifyingFeelings", component="Body") |> 
    mutate(var_Correlate = "Difficulties Identifying Feelings"),
  make_cor(df3, vars_cfa, with = "TAS_DescribingFeelings", component="Body") |> 
    mutate(var_Correlate = "Difficulties Describing Feelings"),
  make_cor(df3, vars_cfa, with = "TAS_ExternallyOrientedThinking", component="Body") |> 
    mutate(var_Correlate = "External Thinking"),
  make_cor(df5, vars_cfa, with = "TAS_DIF", component="Body") |> 
    mutate(var_Correlate = "Difficulties Identifying Feelings"),
  make_cor(df5, vars_cfa, with = "TAS_DDF", component="Body") |> 
    mutate(var_Correlate = "Difficulties Describing Feelings")
)
```

### Mood

```{r}
data <- rbind(
  data,
  make_cor(df5, vars_cfa, with = "PHQ15_sum", component="Mood") |> 
    mutate(var_Correlate = "Somatic Concerns (PHQ-15)"),
  make_cor(df5, vars_cfa, with = "PHQ9_sum", component="Mood")|> 
    mutate(var_Correlate = "Depression (PHQ-9)"),
  make_cor(df7, vars_cfa7, with = "PHQ4_Depression", component="Mood") |> 
    mutate(var_Correlate = "Depression (PHQ-4)"),
  make_cor(df7, vars_cfa7, with = "PHQ4_Anxiety", component="Mood") |> 
    mutate(var_Correlate = "Anxiety (PHQ-4)")
)
```

```{r}
data <- rbind(
  data,
  make_cor(df6, vars_cfa, with = "BDI2_Total", component="Mood") |> 
    mutate(var_Correlate = "Depression (BDI-II)")
)
```

```{r}
data <- rbind(
  data,
  make_cor(df6, vars_cfa, with = "STAI5_General", component="Mood") |> 
    mutate(var_Correlate = "Anxiety (STAI-5)")
)
```

```{r}
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "BPD", component="Mood") |> 
    mutate(var_Correlate = "Boderline Personality")
)
```

### Psychopathology

```{r}
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "PID5_Antagonism", component="Maladaptive") |> 
    mutate(var_Correlate = "Antagonism"),
  make_cor(df7, vars_cfa7, with = "PID5_Detachment", component="Maladaptive") |> 
    mutate(var_Correlate = "Detachment"),
  make_cor(df7, vars_cfa7, with = "PID5_Disinhibition", component="Maladaptive") |> 
    mutate(var_Correlate = "Disinhibition"),
  make_cor(df7, vars_cfa7, with = "PID5_NegativeAffect", component="Maladaptive") |> 
    mutate(var_Correlate = "Negative Affect"),
  make_cor(df7, vars_cfa7, with = "PID5_Psychoticism", component="Maladaptive") |> 
    mutate(var_Correlate = "Psychoticism")
)
```

```{r}
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "SPQ_ConstrictedAffect", component="Schizotypic") |> 
    mutate(var_Correlate = "Constricted Affect"),
  make_cor(df7, vars_cfa7, with = "SPQ_Eccentric", component="Schizotypic") |> 
    mutate(var_Correlate = "Eccentric"),
  make_cor(df7, vars_cfa7, with = "SPQ_MagicalThinking", component="Schizotypic") |> 
    mutate(var_Correlate = "Magical Thinking"),
  make_cor(df7, vars_cfa7, with = "SPQ_NoCloseFriends", component="Schizotypic") |> 
    mutate(var_Correlate = "No Close Friends"),
  make_cor(df7, vars_cfa7, with = "SPQ_OddSpeech", component="Schizotypic") |> 
    mutate(var_Correlate = "Odd Speech"),
  make_cor(df7, vars_cfa7, with = "SPQ_Reference", component="Schizotypic") |> 
    mutate(var_Correlate = "Reference"),
  make_cor(df7, vars_cfa7, with = "SPQ_SocialAnxiety", component="Schizotypic") |> 
    mutate(var_Correlate = "Social Anxiety"),
  make_cor(df7, vars_cfa7, with = "SPQ_Suspiciousness", component="Schizotypic") |> 
    mutate(var_Correlate = "Suspiciousness"),
  make_cor(df7, vars_cfa7, with = "SPQ_UnusualPerceptions", component="Schizotypic") |> 
    mutate(var_Correlate = "Unusual Perceptions")
)
```

```{r}
#  names(select(df7, starts_with("ASQ")))
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "ASQ_Imagination", component="Autistic") |> 
    mutate(var_Correlate = "Imagination"),
  make_cor(df7, vars_cfa7, with = "ASQ_LackSocialSkills", component="Autistic") |>
    mutate(var_Correlate = "Lack of Social Skills"),
  make_cor(df7, vars_cfa7, with = "ASQ_LowAttentionalSwitching", component="Autistic") |>
    mutate(var_Correlate = "Low Attentional Switching"),
  make_cor(df7, vars_cfa7, with = "ASQ_Patterns", component="Autistic") |>
    mutate(var_Correlate = "Patterns and Numbers"),
  make_cor(df7, vars_cfa7, with = "ASQ_Routine", component="Autistic") |>
    mutate(var_Correlate = "Routines")
)
```

### Personality

```{r}
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "IPIP6_Agreeableness", component="Personality") |> 
    mutate(var_Correlate = "Agreeableness"),
  make_cor(df7, vars_cfa7, with = "IPIP6_Conscientiousness", component="Personality") |>
    mutate(var_Correlate = "Conscientiousness"),
  make_cor(df7, vars_cfa7, with = "IPIP6_Neuroticism", component="Personality") |>
    mutate(var_Correlate = "Neuroticism"),
  make_cor(df7, vars_cfa7, with = "IPIP6_Openness", component="Personality") |>
    mutate(var_Correlate = "Openness"),
  make_cor(df7, vars_cfa7, with = "IPIP6_Extraversion", component="Personality") |>
    mutate(var_Correlate = "Extraversion"),
  make_cor(df7, vars_cfa7, with = "IPIP6_HonestyHumility", component="Personality") |>
    mutate(var_Correlate = "Honesty-Humility")
)
```

```{r}
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "PI_Alive", component="World Beliefs") |> 
    mutate(var_Correlate = "Alive"),
  make_cor(df7, vars_cfa7, with = "PI_Changing", component="World Beliefs") |>
    mutate(var_Correlate = "Changing"),
  make_cor(df7, vars_cfa7, with = "PI_Enticing", component="World Beliefs") |>
    mutate(var_Correlate = "Enticing"),
  make_cor(df7, vars_cfa7, with = "PI_Good", component="World Beliefs") |>
    mutate(var_Correlate = "Good"),
  make_cor(df7, vars_cfa7, with = "PI_Hierarchical", component="World Beliefs") |>
    mutate(var_Correlate = "Hierarchical"),
  make_cor(df7, vars_cfa7, with = "PI_Safe", component="World Beliefs") |>
    mutate(var_Correlate = "Safe")
)
```

```{r}
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "GCBS_Extraterrestrial", component="Conspiracy Beliefs") |> 
    mutate(var_Correlate = "Extraterrestrial"),
  make_cor(df7, vars_cfa7, with = "GCBS_GlobalConspiracies", component="Conspiracy Beliefs") |>
    mutate(var_Correlate = "Global Conspiracies"),
  make_cor(df7, vars_cfa7, with = "GCBS_GovernmentMalfeasance", component="Conspiracy Beliefs") |>
    mutate(var_Correlate = "Government Malfeasance"),
  make_cor(df7, vars_cfa7, with = "GCBS_InformationControl", component="Conspiracy Beliefs") |>
    mutate(var_Correlate = "Information Control"),
  make_cor(df7, vars_cfa7, with = "GCBS_PersonalWellbeing", component="Conspiracy Beliefs") |>
    mutate(var_Correlate = "Personal Wellbeing")
)
```

```{r}
data <- rbind(
  data,
  make_cor(df7, vars_cfa7, with = "LIE_Ability", component="Lying Profile") |> 
    mutate(var_Correlate = "Ability"),
  make_cor(df7, vars_cfa7, with = "LIE_Frequency", component="Lying Profile") |>
    mutate(var_Correlate = "Frequency"),
  make_cor(df7, vars_cfa7, with = "LIE_Contextuality", component="Lying Profile") |>
    mutate(var_Correlate = "Contextuality"),
  make_cor(df7, vars_cfa7, with = "LIE_Negativity", component="Lying Profile") |>
    mutate(var_Correlate = "Negativity")
)
```

### Summary

```{r}
#| fig-height: 20
#| fig-width: 15

p <- data |>
  mutate(sig = ifelse(p < .01, "p < .01", ifelse(p < .001, "p < .001", "N.S.")),
         dir = sign(r),
         var_Interoception = str_replace(var_Interoception, "CoughSneeze", "Cough/Sneeze"),
         var_Interoception = str_replace(var_Interoception, "HeartBreathing", "Heart/Breathing"),
         var_Interoception = str_replace(var_Interoception, "HungryThirsty", "Hungry/Thirsty"),
         var_Interoception = str_replace(var_Interoception, "ItchBruise", "Itch/Bruise"),
         var_Interoception = str_replace(var_Interoception, "MusclesPain", "Muscles/Pain"),
         var_Interoception = str_replace(var_Interoception, "UrinateDefecate", "Urinate/Defecate"),
         var_Interoception = str_replace(var_Interoception, "WindBurp", "Wind/Burp"),
         var_Interoception = str_replace(var_Interoception, "Expulsion_Sudden", "Expulsion - Sudden"),
         var_Interoception = fct_relevel(var_Interoception, "Heart/Breathing", "Muscles/Pain", "Itch/Bruise", "Hungry/Thirsty", "Urinate/Defecate", "Wind/Burp", "Cough/Sneeze"),
         Component = fct_relevel(
           Component, "Demographic", "MAIA", "Body", "Mood", "Maladaptive",
           "Schizotypic", "Autistic", "Personality", "Conspiracy Beliefs", "World Beliefs", "Lying Profile")) |>
  ggplot(aes(x = var_Correlate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  # geom_bar(aes(fill=Sample, alpha=sig, y=r, group=Sample), stat="identity", position = position_dodge2(width = 0.8)) +
  geom_linerange(aes(color=dir, alpha=sig, ymin=0, ymax=r, group=Sample), position = position_dodge2(width = 0.7), linewidth=2.5) +
  geom_linerange(aes(ymin = CI_low, ymax = CI_high, color = dir, alpha=sig, group=Sample), position = position_dodge2(width = 0.7)) +
  scale_fill_gradient(low = "#2196F3", high = "#FF5722", guide = "none") +
  scale_color_gradient(low = "#2196F3", high = "#FF5722", guide = "none") +
  scale_alpha_discrete(range = c(0.2, 0.7), guide = "none") +
  scale_y_continuous(expand = c(0, 0), breaks = c(-0.3, 0, 0.3)) +
  facet_grid(Component ~ var_Interoception, scales = "free_y", switch = "y") +
  coord_flip(ylim = c(-0.45, 0.45)) +
  theme_modern(axis.title.space = 5) +
  theme(panel.grid.major.y = element_line(),
        axis.title.y = element_blank(),
        axis.line.x = element_blank(),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(size = rel(0.8)),
        plot.title = element_text(face = "bold", hjust = 0),
        strip.placement = "outside",
        strip.background.y = element_rect(fill = "#E0E0E0", color = "white"),
        strip.background.x = element_rect(fill = "#F8BBD0", color = "white")) 
ggsave("figures/Figure2.png", p, width=21*0.6, height=29.7*0.45, dpi=200, bg="white")
p
```

## Discussion

Benefits of the IAS:
- Straightforward and sensation-centered items

Recommendations:
- Remove Itch (redundant + issue in Chinese)
- Use analog scales

Limitations:
- Not much clear theorethical or empirical structure (small grouping of items)
- Limited variability (clear mode at 4/5)
- Ambiguous items which grouping depends on the context (and its awareness)
  - E.g., heart beating fast, vomit when nauseaous
- Few items for some modalities (e.g., 1 for heart)
- Positive phrasing: benefits but also might exacerbate positivity bias (and thus unidimensionality)

Need for context-specific items (cross-modal when possible, i.E., cardioception, respiroception, etc.).

New Scale: Multimodal Interoceptive Sensitivity Scale (MISS)