### 2024-02-26 analysis/study1.qmd
separate 2 studies into 2 files

---
title: "Meta-analysis of the Interoceptive Accuracy Scale"
subtitle: "Study 1"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    fig-dpi: 100
    fig-width: 11.33
execute: 
  cache: true
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(lavaan)
library(ggraph)
library(tidySEM)
library(EGAnet)
```

## Data Preparation

::: {.panel-tabset}

### Sample 1

Murphy (https://osf.io/3m5nh)

```{r}
df1a <- haven::read_sav("../data/Murphy2020/Study 1.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))

df1b <- haven::read_sav("../data/Murphy2020/Study 6 IAS.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))
```

### Sample 2

Gaggero (2021) (https://osf.io/5x9sg)

```{r}
load("../data/Gaggero2021/data.rda")
df2 <- data |> 
  mutate(Gender = as.character(gender)) |> 
  rename(
    Age=age, Heart = `IAS 1`, Hungry = `IAS 2`, Breathing = `IAS 3`, Thirsty = `IAS 4`,
    Urinate = `IAS 5`, Defecate = `IAS 6`, Taste = `IAS 7`, Vomit = `IAS 8`, Sneeze = `IAS 9`,
    Cough = `IAS 10`, Temp = `IAS 11`, Sex_arousal = `IAS 12`, Wind = `IAS 13`, Burp = `IAS 14`,
    Muscles = `IAS 15`, Bruise = `IAS 16`, Pain = `IAS 17`, Blood_Sugar = `IAS 18`,
    Affective_touch = `IAS 19`, Tickle = `IAS 20`, Itch = `IAS 21`)
rm(data)
```

### Sample 3

Campos (2022) - Study 1 (https://osf.io/j6ef3)

```{r}
df3 <- haven::read_sav("../data/Campos2022/Dataset_Test.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Sex == 1, "Male", ifelse(Sex == 0, "Female", "Other")))) |>
  rename(
    Heart = IAS1, Hungry = IAS2, Breathing = IAS3, Thirsty = IAS4,
    Urinate = IAS5, Defecate = IAS6, Taste = IAS7, Vomit = IAS8, Sneeze = IAS9,
    Cough = IAS10, Temp = IAS11, Sex_arousal = IAS12, Wind = IAS13, Burp = IAS14,
    Muscles = IAS15, Bruise = IAS16, Pain = IAS17, Blood_Sugar = IAS18,
    Affective_touch = IAS19, Tickle = IAS20, Itch = IAS21
  )
```

### Sample 4

Todd (2022) - https://osf.io/ms354/

```{r}
df4 <- haven::read_sav("../data/Todd2022/CompleteData_new.sav") |>
  rename_with(\(x) str_remove(x, "IAS_"), .cols = starts_with("IAS_")) |>
  rename(
    Heart = IAS1, Hungry = IAS2, Breathing = IAS3, Thirsty = IAS4,
    Urinate = IAS5, Defecate = IAS6, Taste = IAS7, Vomit = IAS8, Sneeze = IAS9,
    Cough = IAS10, Temp = IAS11, Sex_arousal = IAS12, Wind = IAS13, Burp = IAS14,
    Muscles = IAS15, Bruise = IAS16, Pain = IAS17, Blood_Sugar = IAS18,
    Affective_touch = IAS19, Tickle = IAS20, Itch = IAS21
  )
```

### Sample 5

Arslanova (2022) - https://osf.io/mp3cy/

```{r}
df5 <- readxl::read_excel("../data/Arslanova2022/Murphy_Iacc_scale.xlsx")  |> 
  filter(str_detect(Question.Key, "IAC_")) |> 
  filter(str_detect(Question.Key, "-quantised")) |> 
  pivot_wider(names_from = Question.Key, values_from = Response) |>
  rename_with(\(x) str_remove(x, "IAC_"), .cols = starts_with("IAC_")) |> 
  rename_with(\(x) str_remove(x, "-quantised"), .cols = everything()) |> 
  rename(Sex_arousal = SexuallyAroused, Itch = Itchy, Sex_arousal = SexuallyAroused,
         Temp = HotCold, Tickle = Ticklish, Breathing= BreatheFast, 
         Affective_touch= PleasantAffectionate, Blood_Sugar= BloodSugar, 
         Muscles=TiredMuscles, Heart= FastHeart, Taste=Tastes) |> 
  select(-starts_with("Participant")) |> 
  mutate_all(as.numeric)
```

### Sample 6

Brand (2022) - https://osf.io/xwz6g/

```{r}
df6 <- haven::read_sav("../data/Brand2022/LatentVariableApproach.sav") |> 
  select(-SERIAL) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(GENDER == 2, "Male", ifelse(GENDER == 1, "Female", "Other")))) |> 
  rename(
    Age=AGE, Heart = IAS_01, Hungry = IAS_02, Breathing = IAS_03, Thirsty = IAS_04,
    Urinate = IAS_05, Defecate = IAS_06, Taste = IAS_07, Vomit = IAS_08, Sneeze = IAS_09,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )
```

### Sample 7

Brand (2023) - https://osf.io/3f2h6/

```{r}
df7a <- haven::read_sav("../data/Brand2023/Data_Giessen.sav") |> 
  select(-COUNTRY_OTHER) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(SEX == 1, "Male", ifelse(SEX == 2, "Female", "Other")))) |> 
  rename(
    Age=AGE, Heart = IAS_01, Hungry = IAS_02, Breathing = IAS_03, Thirsty = IAS_04,
    Urinate = IAS_05, Defecate = IAS_06, Taste = IAS_07, Vomit = IAS_08, Sneeze = IAS_09,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )

df7b <- haven::read_sav("../data/Brand2023/Data_Mainz.sav") |> 
  select(-COUNTRY_OTHER, -EDUCATION_OTHER, -Sample) |> 
  mutate_all(as.numeric) |> 
  mutate(Gender = as.character(ifelse(GENDER == 2, "Male", ifelse(GENDER == 1, "Female", "Other")))) |> 
  rename(
    Age=AGE, Heart = IAS_01, Hungry = IAS_02, Breathing = IAS_03, Thirsty = IAS_04,
    Urinate = IAS_05, Defecate = IAS_06, Taste = IAS_07, Vomit = IAS_08, Sneeze = IAS_09,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )

df7c <- haven::read_sav("../data/Brand2023/Data_PotVie.sav") |> 
  select(-ID) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(GENDER == 2, "Male", ifelse(GENDER == 1, "Female", "Other")))) |> 
  rename(
    Age=AGE, Heart = IA02_01, Hungry = IA02_02, Breathing = IA02_03, Thirsty = IA02_04,
    Urinate = IA02_05, Defecate = IA02_06, Taste = IA02_07, Vomit = IA02_08, Sneeze = IA02_09,
    Cough = IA02_10, Temp = IA02_11, Sex_arousal = IA02_12, Wind = IA02_13, Burp = IA02_14,
    Muscles = IA02_15, Bruise = IA02_16, Pain = IA02_17, Blood_Sugar = IA02_18,
    Affective_touch = IA02_19, Tickle = IA02_20, Itch = IA02_21
  )
```

### Sample 8

Lin (2023) - https://osf.io/3eztd
- Note: No tickle because same Chinese character

```{r}
df8a <- haven::read_sav("../data/Lin2023/Study 1 & 3.sav") |>
  select(-sex) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(sex_dummy == 1, "Male", ifelse(sex_dummy == 0, "Female", "Other")))) |>
  rename(
    Age = age,
    Heart = Heart, Hungry = HUNGR, Breathing = BREAT, Thirsty = Thirs,
    Urinate = URINA, Defecate = Defec, Taste = TASTE, Vomit = VOMIT, Sneeze = Sneez,
    Cough = COUGH, Temp = TEMPE, Sex_arousal = SEXAR, Wind = WIND, Burp = Burp,
    Muscles = MUSCL, Bruise = Bruis, Pain = PAIN, Blood_Sugar = BloSu,
    Affective_touch = Touch, Itch = ITCH
  )

df8b <- haven::read_sav("../data/Lin2023/Study 2.sav") |>
  mutate(Gender = as.character(ifelse(Sex == "男", "Male", ifelse(Sex == "女", "Female", "Other")))) |>
  rename(
    Heart = Heart, Hungry = HUNGR, Breathing = BREAT, Thirsty = Thirs,
    Urinate = URINA, Defecate = Defec, Taste = TASTE, Vomit = VOMIT, Sneeze = Sneez,
    Cough = COUGH, Temp = TEMPE, Sex_arousal = SEXAR, Wind = WIND, Burp = Burp,
    Muscles = MUSCL, Bruise = Bruis, Pain = PAIN, Blood_Sugar = BloSu,
    Affective_touch = Touch, Itch = ITCH)
```

### Sample 9

VonMohr (2023) - Study 3 (https://osf.io/7p9u5/)

```{r}
df9 <- read.csv("../data/VonMohr2023/DataSet_study3.csv")
df9 <- df9[complete.cases(select(df9, starts_with("IAS_"))), ]
df9 <- df9 |>
  mutate(Gender = as.character(ifelse(GENDER == "Man", "Male", ifelse(GENDER == "Woman", "Female", "Other")))) |>
  rename(
    Age=AGE, Heart = IAS_1, Hungry = IAS_2, Breathing = IAS_3, Thirsty = IAS_4,
    Urinate = IAS_5, Defecate = IAS_6, Taste = IAS_7, Vomit = IAS_8, Sneeze = IAS_9,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )
```

<!-- ##### Sample 10 -->

<!-- Bowling (2024) - https://osf.io/tx8h9/ -->

<!-- ```{r} -->
<!-- df10 <- haven::read_sav("../data/Bowling2024/TouchAttitudesData.sav") |> -->
<!--   mutate_all(as.numeric) |> -->
<!--   mutate(Gender = as.character(ifelse(Gender_coded == 1, "Male", ifelse(Gender_coded == 2, "Female", "Other")))) |> -->
<!--   rename( -->
<!--     Heart = IAS1, Hungry = IAS2, Breathing = IAS3, Thirsty = IAS4, -->
<!--     Urinate = IAS5, Defecate = IAS6, Taste = IAS7, Vomit = IAS8, Sneeze = IAS9, -->
<!--     Cough = IAS10, Temp = IAS11, Sex_arousal = IAS12, Wind = IAS13, Burp = IAS14, -->
<!--     Muscles = IAS15, Bruise = IAS16, Pain = IAS17, Blood_Sugar = IAS18, -->
<!--     Affective_touch = IAS19, Tickle = IAS20, Itch = IAS21 -->
<!--   )  -->

<!-- ``` -->

### Sample 10

Makowski

- Note: First sample has some missing items
- Note: Analog scales

```{r}
df10a <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |>
  rename(
    Gender = Sex, Heart = Item_IAS_Interoception_1, Hungry = Item_IAS_Interoception_2,
    Breathing = Item_IAS_Interoception_3, Thirsty = Item_IAS_Interoception_4,
    Temp = Item_IAS_Interoception_5, Sex_arousal = Item_IAS_Interoception_6,
    Urinate = Item_IAS_Elimination_1, Defecate = Item_IAS_Elimination_2,
    Vomit = Item_IAS_Elimination_3, Wind = Item_IAS_Expulsion_1,
    Burp = Item_IAS_Expulsion_2, Sneeze = Item_IAS_Expulsion_3,
    Muscles = Item_IAS_Nociception_1, Bruise = Item_IAS_Nociception_2,
    Pain = Item_IAS_Nociception_3, Affective_touch = Item_IAS_Skin_1,
    Tickle = Item_IAS_Skin_2, Itch = Item_IAS_Skin_3
  ) |>
  filter(!is.na(Urinate))

df10b <- read.csv("https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data.csv") |>
  rename(
    Heart = IAS_1, Hungry = IAS_2, Breathing = IAS_3, Thirsty = IAS_4,
    Urinate = IAS_5, Defecate = IAS_6, Taste = IAS_7, Vomit = IAS_8, Sneeze = IAS_9,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )

# Save all
data <- list(df1a=df1a, df1b=df1b, df2=df2, df3=df3, df4=df4, df5=df5, df6=df6, df7a=df7a, df7b=df7b, df7c=df7c, df8a=df8a, df8b=df8b, df9=df9, df10a=df10a, df10b=df10b)
save(data, file = "../data/data.RData")
```

:::

## Participants

- **Sample 1a**: Data from [Murphy's (2020)](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769) study 1, downloaded from OSF, included `r report::report_participants(df1a, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 1b**: Data from [Murphy's (2020)](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769) study 6, downloaded from OSF, included `r report::report_participants(df1b, age = "Age", sex = NA, gender = "Gender")`.
- **TODO: ADD INFO**

Total N = `r nrow(df1a) + nrow(df1b) + nrow(df2) + nrow(df3) + nrow(df4) + nrow(df5) + nrow(df6) + nrow(df7a) + nrow(df7b) + nrow(df7c) + nrow(df8a) + nrow(df8b) + nrow(df9) + nrow(df10a) + nrow(df10b)`.

## Descriptive

### Distribution

The items with the differing distribution pattern (with a lower mode around 2/5) are "Affective Touch", "Blood Sugar" and "Bruise".

```{r}
vars <- c(
  "Heart", "Hungry", "Breathing", "Thirsty", "Urinate", "Defecate", "Taste", "Vomit", "Sneeze", "Cough", "Temp",
  "Sex_arousal", "Wind", "Burp", "Muscles", "Bruise", "Pain", "Blood_Sugar", "Affective_touch", "Tickle", "Itch"
)

dens1a <- estimate_density(select(df1a, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 1a")
dens1b <- estimate_density(select(df1b, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 1b")
dens2 <- estimate_density(select(df2, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 2")
dens3 <- estimate_density(select(df3, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 3")
dens4 <- estimate_density(select(df4, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 4")
dens5 <- estimate_density(select(df5, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 5")
dens6 <- estimate_density(select(df6, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 6")
dens7a <- estimate_density(select(df7a, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 7a")
dens7b <- estimate_density(select(df7b, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 7b")
dens7c <- estimate_density(select(df7c, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 7c")
dens8a <- estimate_density(select(df8a, all_of(setdiff(vars, "Tickle"))), method = "kernSmooth") |>
  mutate(Sample = "Sample 8a")
dens8b <- estimate_density(select(df8b, all_of(setdiff(vars, "Tickle"))), method = "kernSmooth") |>
  mutate(Sample = "Sample 8b")
dens9 <- estimate_density(select(df9, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 9")
dens10a <- estimate_density(select(df10a, all_of(setdiff(vars, c("Taste", "Cough", "Blood_Sugar")))), method = "kernSmooth") |>
  mutate(Sample = "Sample 10a",
         x = datawizard::rescale(x, to=c(1, 5), range=c(0, 1)))
dens10b <- estimate_density(select(df10b, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 10b",
         x = datawizard::rescale(x, to=c(1, 5), range=c(0, 1)))

p1 <- rbind(dens1a, dens1b, dens2, dens3, dens4, dens5, dens6, dens7a, dens7b, dens7c, dens8a, dens8b, dens9, dens10a, dens10b) |>
  mutate(Sample = fct_relevel(Sample, "Sample 1a", "Sample 1b", "Sample 2", "Sample 3", "Sample 4", "Sample 5", "Sample 6", "Sample 7a", "Sample 7b", "Sample 7c", "Sample 8a", "Sample 8b", "Sample 9", "Sample 10a", "Sample 10b")) |>
  # mutate(Dashed = ifelse(Parameter %in% c("Affective_touch", "Blood_Sugar", "Bruise"), TRUE, FALSE)) |> 
  ggplot(aes(x = x, y = y, color = Parameter)) +
  geom_line() +
  scale_color_material_d() +
  scale_linetype_manual(values = c("solid", "dashed"), guide="none") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        plot.title = element_text(face="bold")) +
  labs(title = "Distribution of responses for all items across various datasets", x = "Response", y = "Density", color = "Item") +
  facet_wrap(~Sample, scales = "free_y", nrow = 5)

ggsave("figures/Figure1.png", p1, width=10, height=10, dpi=200, bg="white")
p1
```

```{r}
data1a <- normalize(select(df1a, all_of(dens1a$Parameter)), range = c(1, 5))
data1b <- normalize(select(df1b, all_of(dens1b$Parameter)), range = c(1, 5))
data2 <- normalize(select(df2, all_of(dens2$Parameter)), range = c(1, 5))
data3 <- normalize(select(df3, all_of(dens3$Parameter)), range = c(1, 5))
data4 <- normalize(select(df4, all_of(dens4$Parameter)), range = c(1, 5))
data5 <- normalize(select(df5, all_of(dens5$Parameter)), range = c(1, 5))
data6 <- normalize(select(df6, all_of(dens6$Parameter)), range = c(1, 5))
data7a <- normalize(select(df7a, all_of(dens7a$Parameter)), range = c(1, 5))
data7b <- normalize(select(df7b, all_of(dens7b$Parameter)), range = c(1, 5))
data7c <- normalize(select(df7c, all_of(dens7c$Parameter)), range = c(1, 5))
data8a <- normalize(select(df8a, all_of(dens8a$Parameter)), range = c(1, 5))
data8b <- normalize(select(df8b, all_of(dens8b$Parameter)), range = c(1, 5))
data9 <- normalize(select(df9, all_of(dens9$Parameter)), range = c(1, 5))
data10a <- select(df10a, all_of(dens10a$Parameter))
data10b <- select(df10b, all_of(dens10b$Parameter))

data_all <- rbind(
  data1a, data1b, data2, data3, data4, data5, data6, data7a, data7b, data7c,
  mutate(data8a, Tickle = NA), mutate(data8b, Tickle = NA), data9,
  mutate(data10a, Taste = NA, Cough = NA, Blood_Sugar = NA), data10b
) 
```

### Correlations

An overall positive intercorrelation pattern, with no clear structure emerging.

::: {.panel-tabset}

#### All data

```{r}
make_correlation <- function(df) {
  correlation::correlation(df, redundant = TRUE) |>
    correlation::cor_sort() |>
    correlation::cor_lower() |>
    mutate(val = paste0(insight::format_value(r), format_p(p, stars_only = TRUE))) |>
    mutate(Parameter2 = fct_rev(Parameter2)) |>
    ggplot(aes(x = Parameter1, y = Parameter2)) +
    geom_tile(aes(fill = r), color = "white") +
    geom_text(aes(label = val), size = 3) +
    labs(title = "Correlation Matrix") +
    scale_fill_metro_c(limit = c(0, 0.75), guide = guide_colourbar(ticks = FALSE)) +
    theme_minimal() +
    theme(
      legend.title = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

make_correlation(data_all)
```

#### Sample 1a

```{r}
make_correlation(data1a)
```

#### Sample 1b

```{r}
make_correlation(data1b)
```

#### Sample 2

```{r}
make_correlation(data2)
```

#### Sample 3

```{r}
make_correlation(data3)
```

#### Sample 4

```{r}
make_correlation(data4)
```

#### Sample 5

```{r}
make_correlation(data5)
```

#### Sample 6

```{r}
make_correlation(data6)
```

#### Sample 7a

```{r}
make_correlation(data7a)
```

#### Sample 7b

```{r}
make_correlation(data7b)
```

#### Sample 7c

```{r}
make_correlation(data7c)
```

#### Sample 8a

```{r}
make_correlation(data8a)
```

#### Sample 8b

```{r}
make_correlation(data8b)
```

#### Sample 9

```{r}
make_correlation(data9)
```

#### Sample 10a

```{r}
make_correlation(data10a)
```

#### Sample 10b

```{r}
make_correlation(data10b)
```

:::

## Graph Analysis

Exploratory Graph Analysis (EGA) is a recently developed framework for psychometric assessment, that can be used to estimate the number of dimensions in questionnaire data using network estimation methods and community detection algorithms, and assess the stability of dimensions and items using bootstrapping.
See https://r-ega.net/articles/ega.html for details.

### Unique Variable Analysis (UVA)

Unique Variable Analysis (Christensen, Garrido, & Golino, 2023) uses the weighted topological overlap measure (Nowick et al., 2009) on an estimated network. Values greater than 0.25 are determined to have considerable local dependence (i.e., redundancy) that should be handled (variables with the highest maximum weighted topological overlap to all other variables (other than the one it is redundant with) should be removed).

::: {.panel-tabset}

#### All data

```{r}
uva <- EGAnet::UVA(data = data_all, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 1a

```{r}
uva <- EGAnet::UVA(data = data1a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 1b

```{r}
uva <- EGAnet::UVA(data = data1b, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 3

```{r}
uva <- EGAnet::UVA(data = data3, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 4

```{r}
uva <- EGAnet::UVA(data = data4, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 5

```{r}
uva <- EGAnet::UVA(data = data5, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 6

```{r}
uva <- EGAnet::UVA(data = data6, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 7a

```{r}
uva <- EGAnet::UVA(data = data7a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 7b

```{r}
uva <- EGAnet::UVA(data = data7b, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 7c

```{r}
uva <- EGAnet::UVA(data = data7c, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 8a

```{r}
uva <- EGAnet::UVA(data = data8a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 8b

```{r}
uva <- EGAnet::UVA(data = data8b, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 9

```{r}
uva <- EGAnet::UVA(data = data9, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 10a

```{r}
uva <- EGAnet::UVA(data = data10a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 10b

```{r}
uva <- EGAnet::UVA(data = data10b, cut.off = 0.3)
uva
uva$keep_remove
```

:::

### Networks

::: {.panel-tabset}

#### All data

```{r ega_all}
#| fig-height: 20
#| fig-width: 12

make_egas <- function(data) {
  egas <- list()
  for (model in c("glasso")) {  # , "TMFG"
    for (algo in c("walktrap", "louvain")) {
      for (type in c("ega", "ega.fit")) { # "" # "hierega"
        if (type == "ega.fit" & algo == "louvain") next # Too slow
        egas[[paste0(model, "_", algo, "_", type)]] <- EGAnet::bootEGA(
          data = data,
          seed = 123,
          model = model,
          algorithm = algo,
          EGA.type = type,
          type = "resampling",
          plot.itemStability = FALSE,
          verbose = FALSE
        )
      }
    }
  }

  p <- EGAnet::compare.EGA.plots(
    egas$glasso_walktrap_ega, 
    # egas$glasso_walktrap_ega.fit,
    egas$glasso_louvain_ega, 
    # egas$TMFG_louvain_ega,
    # egas$TMFG_walktrap_ega, 
    # egas$TMFG_walktrap_ega.fit,
    labels = c(
      "glasso_walktrap_ega", 
      # "glasso_walktrap_ega.fit",
      "glasso_louvain_ega" 
      # "TMFG_louvain_ega",
      # "TMFG_walktrap_ega", 
      # "TMFG_walktrap_ega.fit"
    ),
    rows = 3,
    plot.all = FALSE
  )$all
  list(egas = egas, p = p)
}

egas_all <- make_egas(data_all)
egas_all$p
```

#### Sample 1a

```{r ega1a}
#| fig-height: 20
#| fig-width: 12

egas1a <- make_egas(data1a)
egas1a$p
```

#### Sample 1b

```{r ega1b}
#| fig-height: 20
#| fig-width: 12

egas1b <- make_egas(data1b)
egas1b$p
```

#### Sample 2

```{r ega2}
#| fig-height: 20
#| fig-width: 12

egas2 <- make_egas(data2)
egas2$p
```

#### Sample 3

```{r ega3}
#| fig-height: 20
#| fig-width: 12

egas3 <- make_egas(data3)
egas3$p
```

#### Sample 4

```{r ega4}
#| fig-height: 20
#| fig-width: 12

egas4 <- make_egas(data4)
egas4$p
```

#### Sample 5

```{r ega5}
#| fig-height: 20
#| fig-width: 12

egas5 <- make_egas(data5)
egas5$p
```

#### Sample 6

```{r ega6}
#| fig-height: 20
#| fig-width: 12

egas6 <- make_egas(data6)
egas6$p
```

#### Sample 7a

```{r ega7&}
#| fig-height: 20
#| fig-width: 12

egas7a <- make_egas(data7a)
egas7a$p
```

#### Sample 7b

```{r ega7b}
#| fig-height: 20
#| fig-width: 12

egas7b <- make_egas(data7b)
egas7b$p
```

#### Sample 7c

```{r ega7c}
#| fig-height: 20
#| fig-width: 12

egas7c <- make_egas(data7c)
egas7c$p
```

#### Sample 8a

```{r ega8a}
#| fig-height: 20
#| fig-width: 12

egas8a <- make_egas(data8a)
egas8a$p
```

#### Sample 8b

```{r ega8b}
#| fig-height: 20
#| fig-width: 12

egas8b <- make_egas(data8b)
egas8b$p
```

#### Sample 9

```{r ega9}
#| fig-height: 20
#| fig-width: 12

egas9 <- make_egas(data9)
egas9$p
```

#### Sample 10a

```{r ega10a}
#| fig-height: 20
#| fig-width: 12

egas10a <- make_egas(data10a)
egas10a$p
```

#### Sample 10b

```{r ega10b}
#| fig-height: 20
#| fig-width: 12

egas10b <- make_egas(data10b)
egas10b$p
```

:::

### Structure Stability

::: {.panel-tabset}

#### All data

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas_all$egas, plot), nrow = 3)
```

#### Sample 1a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas1a$egas, plot), nrow = 3)
```

#### Sample 1b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas1b$egas, plot), nrow = 3)
```

#### Sample 2

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas2$egas, plot), nrow = 3)
```

#### Sample 3

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas3$egas, plot), nrow = 3)
```

#### Sample 4

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas4$egas, plot), nrow = 3)
```

#### Sample 5

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas5$egas, plot), nrow = 3)
```

#### Sample 6

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas6$egas, plot), nrow = 3)
```

#### Sample 7a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas7a$egas, plot), nrow = 3)
```

#### Sample 7b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas7b$egas, plot), nrow = 3)
```

#### Sample 7c

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas7c$egas, plot), nrow = 3)
```

#### Sample 8a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas8a$egas, plot), nrow = 3)
```

#### Sample 8b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas8b$egas, plot), nrow = 3)
```

#### Sample 9

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas9$egas, plot), nrow = 3)
```

#### Sample 10a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas10a$egas, plot), nrow = 3)
```

#### Sample 10b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas10b$egas, plot), nrow = 3)
```

:::

### Scores

```{r}
ega <- egas_all$egas$glasso_louvain_ega$EGA
plot(ega)

get_scores <- function(data, ega) {
  EGAnet::net.scores(data, ega)$scores$std.scores |>
    as.data.frame() |>
    setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))
}

ega_scores_1a <- get_scores(data1a, ega)
ega_scores_1b <- get_scores(data1b, ega)
ega_scores_2 <- get_scores(data2, ega)
ega_scores_3 <- get_scores(data3, ega)
ega_scores_4 <- get_scores(data4, ega)
ega_scores_5 <- get_scores(data5, ega)
ega_scores_6 <- get_scores(data6, ega)
ega_scores_7a <- get_scores(data7a, ega)
ega_scores_7b <- get_scores(data7b, ega)
ega_scores_7c <- get_scores(data7c, ega)
ega_scores_8a <- get_scores(data8a, ega)
ega_scores_8b <- get_scores(data8b, ega)
ega_scores_9 <- get_scores(data9, ega)
ega_scores_10a <- get_scores(data10a, ega)
ega_scores_10b <- get_scores(data10b, ega)
```

## Factor Analysis

### How Many Factors

::: {.panel-tabset}

#### All data

```{r}
n <- parameters::n_factors(data_all, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 1a

```{r}
n <- parameters::n_factors(data1a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 1b

```{r}
n <- parameters::n_factors(data1b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 2

```{r}
n <- parameters::n_factors(data2, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 3

```{r}
n <- parameters::n_factors(data3, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 4

```{r}
n <- parameters::n_factors(data4, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 5

```{r}
n <- parameters::n_factors(data5, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 6

```{r}
n <- parameters::n_factors(data6, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 7a

```{r}
n <- parameters::n_factors(data7a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 7b

```{r}
n <- parameters::n_factors(data7b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 7c

```{r}
n <- parameters::n_factors(data7c, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 8a

```{r}
n <- parameters::n_factors(data8a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 8b

```{r}
n <- parameters::n_factors(data8b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 9

```{r}
n <- parameters::n_factors(data9, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 10a

```{r}
n <- parameters::n_factors(data10a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 10b

```{r}
n <- parameters::n_factors(data10b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

:::

### EFA

::: {.panel-tabset}

#### All data

```{r}
#| fig-height: 10

efa5 <- parameters::factor_analysis(data_all, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa5)

display(efa5)
```

#### Sample 1a

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data1a, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 1b

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data1b, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

:::

## CFA

We discarded the following items:
- Tickle: not present in the same dataset and consistently flagged as redundant.
<!-- - Blood_Sugar -->

Ambiguous:
- Temp
- Vomit
- Affective_touch
- Sex_arousal
- Taste

### Lower-Level Factors

::: {.panel-tabset}

#### All data

```{r}
m1 <- "
Interoception =~ Hungry + Thirsty + Urinate + Defecate + Itch + Bruise + Muscles + Pain + Breathing + Heart + Cough + Sneeze + Wind + Burp
"

m4 <- "
Sustenance =~ Hungry + Thirsty + Urinate + Defecate + Muscles + Pain
Skin =~ Itch + Bruise
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m5 <- "
Sustenance =~ Hungry + Thirsty + Pain
UrinateDefecate =~ Urinate + Defecate
Skin =~ Itch + Bruise + Muscles
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m6 <- "
HungryThirsty =~ Hungry + Thirsty
UrinateDefecate =~ Urinate + Defecate
ItchBruise =~ Itch + Bruise
MusclesPain =~ Muscles + Pain
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m7 <- "
HungryThirsty =~ Hungry + Thirsty
UrinateDefecate =~ Urinate + Defecate
ItchBruise =~ Itch + Bruise
MusclesPain =~ Muscles + Pain
HeartBreathing =~ Breathing + Heart
CoughSneeze =~ Cough + Sneeze
WindBurp =~ Wind + Burp
"

is_converged <- function(m) {
  ev <- eigen(lavaan::lavTech(m, "cov.lv")[[1]], symmetric=TRUE, only.values=TRUE)$values
  if (any(ev < -1 * .Machine$double.eps^(3/4))) {
    "Not converged"
  } else {
    ""
  }
}

compare_mods <- function(data) {
  m1 <- lavaan::cfa(m1, data = data)
  m4 <- lavaan::cfa(m4, data = data)
  m5 <- lavaan::cfa(m5, data = data)
  m6 <- lavaan::cfa(m6, data = data)
  m7 <- lavaan::cfa(m7, data = data)

  anova(m1, m4, m5, m6, m7) |>
    parameters::parameters() |>
    mutate(AIC = format(round(AIC, 0), nsmall = 0),
           BIC = format(round(BIC, 0), nsmall = 0),
           Chi2 = format(round(Chi2, 2), nsmall = 2),
           Converged = sapply(list(m1, m4, m5, m6, m7), is_converged))
}

display(compare_mods(data_all))

m7_all <- lavaan::cfa(m7, data = data_all)
parameters::parameters(m7_all, standardize=TRUE) |>
  display()
```

#### Sample 1a

```{r}
#| warning: false

display(compare_mods(data1a))
```

:::

Conclusion: The 7th factor solution was preferred in most datasets, including by indices that penalize increased number of parameters (such as BIC).

### Higher Order Factors

::: {.panel-tabset}

#### All data

```{r}
m7h1 <- paste0(m7, "
              Interoception =~ HungryThirsty + UrinateDefecate + ItchBruise + MusclesPain + HeartBreathing + CoughSneeze + WindBurp
              ")

m7h2 <- paste0(m7, "
              Expulsion =~ CoughSneeze + WindBurp
              Interoception =~ HungryThirsty + UrinateDefecate + HeartBreathing
              ")

m7h3 <- paste0(m7, "
              Valenced =~ MusclesPain + HeartBreathing + ItchBruise
              Expulsion =~ CoughSneeze + WindBurp
              Homeostasis =~ HungryThirsty + UrinateDefecate
              ")

compare_mods_h <- function(data) {
  m7 <- lavaan::cfa(m7, data = data)
  m7h2 <- lavaan::cfa(m7h2, data = data)
  m7h1 <- lavaan::cfa(m7h1, data = data)
  m7h3 <- lavaan::cfa(m7h3, data = data)

  anova(m7, m7h1, m7h2, m7h3) |>
    parameters::parameters() |>
    mutate(AIC = format(round(AIC, 0), nsmall = 0),
           BIC = format(round(BIC, 0), nsmall = 0),
           Chi2 = format(round(Chi2, 2), nsmall = 2),
           Converged = sapply(list(m7, m7h1, m7h2, m7h3), is_converged))
}

display(compare_mods_h(data_all))
```

No evidence for higher order factors.

#### Sample 1a

```{r}
display(compare_mods_h(data1a))
```

:::

### Model Performance

```{r}
#| warning: false

rbind(
  performance::performance(m7_all, metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "All"),
  performance::performance(update(m7_all, data = data1a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 1a"),
  performance::performance(update(m7_all, data = data1b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 1b"),
  performance::performance(update(m7_all, data = data2), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 2"),
  performance::performance(update(m7_all, data = data3), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 3"),
  performance::performance(update(m7_all, data = data4), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 4"),
  performance::performance(update(m7_all, data = data5), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 5"),
  performance::performance(update(m7_all, data = data6), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 6"),
  performance::performance(update(m7_all, data = data7a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 7a"),
  performance::performance(update(m7_all, data = data7b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 7b"),
  performance::performance(update(m7_all, data = data7c), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 7c"),
  performance::performance(update(m7_all, data = data8a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 8a"),
  performance::performance(update(m7_all, data = data8b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 8b"),
  performance::performance(update(m7_all, data = data9), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 9"),
  performance::performance(lavaan::cfa(str_remove(m7, fixed("\nCoughSneeze =~ Cough + Sneeze")), data = data10a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 10a"),
  performance::performance(update(m7_all, data = data10b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 10b")
) |>
  display()
```

## Scores

Note the usage of descriptive factor names relating directly to the items to avoid abstraction.

```{r}
#| warning: false

# Add empirical variables
add_empirical <- function(data, df) {
  x <- data |>
    mutate(
      HungryThirsty = (df$Hungry + df$Thirsty) / 2,
      UrinateDefecate = (df$Urinate + df$Defecate) / 2,
      MusclesPain = (df$Muscles + df$Pain) / 2,
      HeartBreathing = (df$Heart + df$Breathing) / 2,
      ItchBruise = (df$Itch + df$Bruise) / 2,
      WindBurp = (df$Wind + df$Burp) / 2)
  if("Cough" %in% names(df)) {
    mutate(x, CoughSneeze = (df$Cough + df$Sneeze) / 2)
  } else {
    mutate(x, CoughSneeze = df$Sneeze)
  }
}

# Deal with missing data
temp <- predict(m7_all, newdata=data7c)
scores7c <- data.frame(matrix(ncol = 7, nrow = nrow(ega_scores_7c))) 
scores7c[!is.na(ega_scores_7c$EGA1), ] <- temp
names(scores7c) <- names(as.data.frame(temp))

scores <- list(
  sample1a = cbind(ega_scores_1a, data_addprefix(as.data.frame(predict(m7_all, newdata=data1a)), "CFA_")) |>
    add_empirical(data1a) |> 
    mutate(Sample = "Sample 1a"),
  sample1b = cbind(ega_scores_1b, data_addprefix(as.data.frame(predict(m7_all, newdata=data1b)), "CFA_")) |>
    add_empirical(data1b) |> 
    mutate(Sample = "Sample 1b"),
  sample2 = cbind(ega_scores_2, data_addprefix(as.data.frame(predict(m7_all, newdata=data2)), "CFA_")) |>
    add_empirical(data2) |> 
    mutate(Sample = "Sample 2"),
  sample3 = cbind(ega_scores_3, data_addprefix(as.data.frame(predict(m7_all, newdata=data3)), "CFA_")) |>
    add_empirical(data3) |> 
    mutate(Sample = "Sample 3"),
  sample4 = cbind(ega_scores_4, data_addprefix(as.data.frame(predict(m7_all, newdata=data4)), "CFA_")) |>
    add_empirical(data4) |> 
    mutate(Sample = "Sample 4"),
  sample5 = cbind(ega_scores_5, data_addprefix(as.data.frame(predict(m7_all, newdata=data5)), "CFA_")) |>
    add_empirical(data5) |> 
    mutate(Sample = "Sample 5"),
  sample6 = cbind(ega_scores_6, data_addprefix(as.data.frame(predict(m7_all, newdata=data6)), "CFA_")) |>
    add_empirical(data6) |> 
    mutate(Sample = "Sample 6"),
  sample7a = cbind(ega_scores_7a, data_addprefix(as.data.frame(predict(m7_all, newdata=data7a)), "CFA_")) |>
    add_empirical(data7a) |> 
    mutate(Sample = "Sample 7a"),
  sample7b = cbind(ega_scores_7b, data_addprefix(as.data.frame(predict(m7_all, newdata=data7b)), "CFA_")) |>
    add_empirical(data7b) |> 
    mutate(Sample = "Sample 7b"),
  sample7c = cbind(ega_scores_7c, data_addprefix(scores7c, "CFA_")) |> 
    add_empirical(data7c) |> 
    mutate(Sample = "Sample 7c"),
  sample8a = cbind(ega_scores_8a, data_addprefix(as.data.frame(predict(m7_all, newdata=data8a)), "CFA_")) |>
    add_empirical(data8a) |> 
    mutate(Sample = "Sample 8a"),
  sample8b = cbind(ega_scores_8b, data_addprefix(as.data.frame(predict(m7_all, newdata=data8b)), "CFA_")) |>
    add_empirical(data8b) |> 
    mutate(Sample = "Sample 8b"),
  sample9 = cbind(ega_scores_9, data_addprefix(as.data.frame(predict(m7_all, newdata=data9)), "CFA_")) |>
    add_empirical(data9) |> 
    mutate(Sample = "Sample 9"),
  sample10a = cbind(ega_scores_10a, data_addprefix(as.data.frame(predict(lavaan::cfa(str_remove(m7, fixed("\nCoughSneeze =~ Cough + Sneeze")), data = data_all), newdata=data10a)), "CFA_")) |>
    add_empirical(data10a) |> 
    mutate(Sample = "Sample 10a", CFA_CoughSneeze = NA),
  sample10b = cbind(ega_scores_10b, data_addprefix(as.data.frame(predict(m7_all, newdata=data10b)), "CFA_")) |>
    add_empirical(data10b) |> 
    mutate(Sample = "Sample 10b")
)

save(scores, file = "../data/scores.RData")
```