### 2025-01-03 analysis/Stud1_a.qmd
rendered html with added data

---
title: "Meta-analysis of the Interoceptive Accuracy Scale (IAS)"
subtitle: "Study 1"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    fig-dpi: 100
    fig-width: 11.33
execute: 
  cache: true
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(lavaan)
library(ggraph)
library(tidySEM)
library(EGAnet)
```

## Data Preparation

::: {.panel-tabset}

### Sample 1

- Murphy (2020)
  - https://osf.io/3m5nh

```{r}
df1a <- haven::read_sav("../data/Murphy2020/Study 1.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))

df1b <- haven::read_sav("../data/Murphy2020/Study 6 IAS.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Gender == 1, "Male", ifelse(Gender == 2, "Female", "Other"))))
```

### Sample 2

- Gaggero (2021) 
  - https://osf.io/5x9sg

```{r}
load("../data/Gaggero2021/data.rda")
df2 <- data |> 
  mutate(Gender = as.character(gender)) |> 
  mutate(across(starts_with("IAS "), as.numeric)) |>
  rename(
    Age=age, Heart = `IAS 1`, Hungry = `IAS 2`, Breathing = `IAS 3`, Thirsty = `IAS 4`,
    Urinate = `IAS 5`, Defecate = `IAS 6`, Taste = `IAS 7`, Vomit = `IAS 8`, Sneeze = `IAS 9`,
    Cough = `IAS 10`, Temp = `IAS 11`, Sex_arousal = `IAS 12`, Wind = `IAS 13`, Burp = `IAS 14`,
    Muscles = `IAS 15`, Bruise = `IAS 16`, Pain = `IAS 17`, Blood_Sugar = `IAS 18`,
    Affective_touch = `IAS 19`, Tickle = `IAS 20`, Itch = `IAS 21`)
rm(data)
```

### Sample 3

- Campos (2022) - Study 1 
  - https://osf.io/j6ef3

```{r}
df3 <- haven::read_sav("../data/Campos2022/Dataset_Test.sav") |>
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(Sex == 1, "Male", ifelse(Sex == 0, "Female", "Other")))) |>
  rename(
    Heart = IAS1, Hungry = IAS2, Breathing = IAS3, Thirsty = IAS4,
    Urinate = IAS5, Defecate = IAS6, Taste = IAS7, Vomit = IAS8, Sneeze = IAS9,
    Cough = IAS10, Temp = IAS11, Sex_arousal = IAS12, Wind = IAS13, Burp = IAS14,
    Muscles = IAS15, Bruise = IAS16, Pain = IAS17, Blood_Sugar = IAS18,
    Affective_touch = IAS19, Tickle = IAS20, Itch = IAS21
  )
```

### Sample 4

- Todd (2022) 
  - https://osf.io/ms354/

```{r}
df4 <- haven::read_sav("../data/Todd2022/CompleteData_new.sav") |>
  rename_with(\(x) str_remove(x, "IAS_"), .cols = starts_with("IAS_")) |>
  rename(
    Heart = IAS1, Hungry = IAS2, Breathing = IAS3, Thirsty = IAS4,
    Urinate = IAS5, Defecate = IAS6, Taste = IAS7, Vomit = IAS8, Sneeze = IAS9,
    Cough = IAS10, Temp = IAS11, Sex_arousal = IAS12, Wind = IAS13, Burp = IAS14,
    Muscles = IAS15, Bruise = IAS16, Pain = IAS17, Blood_Sugar = IAS18,
    Affective_touch = IAS19, Tickle = IAS20, Itch = IAS21
  )
```

### Sample 5

- Arslanova (2022)
  - https://osf.io/mp3cy/
  - note: Final.xlsx includes information about which participants passed the attention checks

```{r}

df5_attention <- readxl::read_excel("C:/Users/asf25/Documents/studies/InteroceptionIAS/data/Arslanova2022/Participants2.xlsx", 
    sheet = "FINAL") |>
  filter(ActorHR == 1) |>
  select(Subj.ID, Gender, Age) |>
  mutate(Gender = as.numeric(Gender), Age = as.numeric(Age)) |>
  filter(!is.na(Age)) |> #error on the documentation of this participant - attention check failed but not noted
  mutate(Gender = case_when(
    Gender== 0 ~ "Male",
    Gender== 1 ~ "Female" # based on paper reporting 65 women
  ))

df5 <- readxl::read_excel("../data/Arslanova2022/Murphy_Iacc_scale.xlsx")  |>
  filter(str_detect(Question.Key, "IAC_")) |> 
  filter(str_detect(Question.Key, "-quantised")) |> 
  pivot_wider(names_from = Question.Key, values_from = Response) |>
  rename_with(\(x) str_remove(x, "IAC_"), .cols = starts_with("IAC_")) |> 
  rename_with(\(x) str_remove(x, "-quantised"), .cols = everything()) |> 
  rename(Sex_arousal = SexuallyAroused, Itch = Itchy, Sex_arousal = SexuallyAroused,
         Temp = HotCold, Tickle = Ticklish, Breathing= BreatheFast, 
         Affective_touch= PleasantAffectionate, Blood_Sugar= BloodSugar, 
         Muscles=TiredMuscles, Heart= FastHeart, Taste=Tastes) |> 
  rename(Subj.ID = "Participant.Public.ID") |>
  select(-starts_with("Participant"))

df5 <- df5 |>
  filter(Subj.ID %in% df5_attention$Subj.ID) |>
  select(-Subj.ID) |>
  mutate(across(everything(), as.numeric))

```

### Sample 6

- Brand (2022) 
  - https://osf.io/xwz6g/

```{r}
df6 <- haven::read_sav("../data/Brand2022/LatentVariableApproach.sav") |> 
  select(-SERIAL) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(GENDER == 2, "Male", ifelse(GENDER == 1, "Female", "Other")))) |> 
  rename(
    Age=AGE, Heart = IAS_01, Hungry = IAS_02, Breathing = IAS_03, Thirsty = IAS_04,
    Urinate = IAS_05, Defecate = IAS_06, Taste = IAS_07, Vomit = IAS_08, Sneeze = IAS_09,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )
```

### Sample 7

- Brand (2023) 
  - https://osf.io/3f2h6/

```{r}
df7a <- haven::read_sav("../data/Brand2023/Data_Giessen.sav") |> 
  select(-COUNTRY_OTHER) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(SEX == 1, "Male", ifelse(SEX == 2, "Female", "Other")))) |> 
  rename(
    Age=AGE, Heart = IAS_01, Hungry = IAS_02, Breathing = IAS_03, Thirsty = IAS_04,
    Urinate = IAS_05, Defecate = IAS_06, Taste = IAS_07, Vomit = IAS_08, Sneeze = IAS_09,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )

df7b <- haven::read_sav("../data/Brand2023/Data_Mainz.sav") |> 
  select(-COUNTRY_OTHER, -EDUCATION_OTHER, -Sample) |> 
  mutate_all(as.numeric) |> 
  mutate(Gender = as.character(ifelse(GENDER == 2, "Male", ifelse(GENDER == 1, "Female", "Other")))) |> 
  rename(
    Age=AGE, Heart = IAS_01, Hungry = IAS_02, Breathing = IAS_03, Thirsty = IAS_04,
    Urinate = IAS_05, Defecate = IAS_06, Taste = IAS_07, Vomit = IAS_08, Sneeze = IAS_09,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )

df7c <- haven::read_sav("../data/Brand2023/Data_PotVie.sav") |> 
  select(-ID) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(GENDER == 2, "Male", ifelse(GENDER == 1, "Female", "Other")))) |> 
  rename(
    Age=AGE, Heart = IA02_01, Hungry = IA02_02, Breathing = IA02_03, Thirsty = IA02_04,
    Urinate = IA02_05, Defecate = IA02_06, Taste = IA02_07, Vomit = IA02_08, Sneeze = IA02_09,
    Cough = IA02_10, Temp = IA02_11, Sex_arousal = IA02_12, Wind = IA02_13, Burp = IA02_14,
    Muscles = IA02_15, Bruise = IA02_16, Pain = IA02_17, Blood_Sugar = IA02_18,
    Affective_touch = IA02_19, Tickle = IA02_20, Itch = IA02_21
  )
```

### Sample 8

- Lin (2023)
  - https://osf.io/3eztd
  - Note: No tickle because same Chinese character

```{r}
df8a <- haven::read_sav("../data/Lin2023/Study 1 & 3.sav") |>
  select(-sex) |> 
  mutate_all(as.numeric) |>
  mutate(Gender = as.character(ifelse(sex_dummy == 1, "Male", ifelse(sex_dummy == 0, "Female", "Other")))) |>
  rename(
    Age = age,
    Heart = Heart, Hungry = HUNGR, Breathing = BREAT, Thirsty = Thirs,
    Urinate = URINA, Defecate = Defec, Taste = TASTE, Vomit = VOMIT, Sneeze = Sneez,
    Cough = COUGH, Temp = TEMPE, Sex_arousal = SEXAR, Wind = WIND, Burp = Burp,
    Muscles = MUSCL, Bruise = Bruis, Pain = PAIN, Blood_Sugar = BloSu,
    Affective_touch = Touch, Itch = ITCH
  )

df8b <- haven::read_sav("../data/Lin2023/Study 2.sav") |>
  mutate(Gender = as.character(ifelse(Sex == "男", "Male", ifelse(Sex == "女", "Female", "Other")))) |>
  rename(
    Heart = Heart, Hungry = HUNGR, Breathing = BREAT, Thirsty = Thirs,
    Urinate = URINA, Defecate = Defec, Taste = TASTE, Vomit = VOMIT, Sneeze = Sneez,
    Cough = COUGH, Temp = TEMPE, Sex_arousal = SEXAR, Wind = WIND, Burp = Burp,
    Muscles = MUSCL, Bruise = Bruis, Pain = PAIN, Blood_Sugar = BloSu,
    Affective_touch = Touch, Itch = ITCH)
```

### Sample 9

- VonMohr (2023) - Study 3 
  - https://osf.io/7p9u5/

```{r}
df9 <- read.csv("../data/VonMohr2023/DataSet_study3.csv")
df9 <- df9[complete.cases(select(df9, starts_with("IAS_"))), ]
df9 <- df9 |>
  mutate(Gender = as.character(ifelse(GENDER == "Man", "Male", ifelse(GENDER == "Woman", "Female", "Other")))) |>
  rename(
    Age=AGE, Heart = IAS_1, Hungry = IAS_2, Breathing = IAS_3, Thirsty = IAS_4,
    Urinate = IAS_5, Defecate = IAS_6, Taste = IAS_7, Vomit = IAS_8, Sneeze = IAS_9,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )
```

### Sample 10

- Makowski
  - Note: First sample has some missing items
  - Note: Analog scales
  - note df10c = primals data

```{r}
df10a <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |>
  rename(
    Gender = Sex, Heart = Item_IAS_Interoception_1, Hungry = Item_IAS_Interoception_2,
    Breathing = Item_IAS_Interoception_3, Thirsty = Item_IAS_Interoception_4,
    Temp = Item_IAS_Interoception_5, Sex_arousal = Item_IAS_Interoception_6,
    Urinate = Item_IAS_Elimination_1, Defecate = Item_IAS_Elimination_2,
    Vomit = Item_IAS_Elimination_3, Wind = Item_IAS_Expulsion_1,
    Burp = Item_IAS_Expulsion_2, Sneeze = Item_IAS_Expulsion_3,
    Muscles = Item_IAS_Nociception_1, Bruise = Item_IAS_Nociception_2,
    Pain = Item_IAS_Nociception_3, Affective_touch = Item_IAS_Skin_1,
    Tickle = Item_IAS_Skin_2, Itch = Item_IAS_Skin_3
  ) |>
  filter(!is.na(Urinate))

df10b <- read.csv("https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data.csv") |>
  rename(
    Heart = IAS_1, Hungry = IAS_2, Breathing = IAS_3, Thirsty = IAS_4,
    Urinate = IAS_5, Defecate = IAS_6, Taste = IAS_7, Vomit = IAS_8, Sneeze = IAS_9,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  )

df10c <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionPrimals/refs/heads/main/data/data_participants.csv") |>
  select("Participant" = "participant_id", "Gender", "Age", "Ethnicity", starts_with("IAS")) |>
   rename(
    Heart = IAS_1, Hungry = IAS_2, Breathing = IAS_3, Thirsty = IAS_4,
    Urinate = IAS_5, Defecate = IAS_6, Taste = IAS_7, Vomit = IAS_8, Sneeze = IAS_9,
    Cough = IAS_10, Temp = IAS_11, Sex_arousal = IAS_12, Wind = IAS_13, Burp = IAS_14,
    Muscles = IAS_15, Bruise = IAS_16, Pain = IAS_17, Blood_Sugar = IAS_18,
    Affective_touch = IAS_19, Tickle = IAS_20, Itch = IAS_21
  ) |>
  filter(!is.na(Heart)) # participant is outlier and had total IAS scores below 0.4

```

### Sample 11

- Giulia

```{r}

df11a <- df11a <- haven::read_sav("../data/Giulia/Interoceptive Attention ESM.sav") |> 
  select(AGE, GENDER, contains("IAS_ACC"), -IAS_ACC_20) |>
  mutate_all(as.numeric) |>
  rename(
    Age = AGE, Gender = GENDER,
    Heart = IAS_ACC_1, Hungry = IAS_ACC_2, Breathing = IAS_ACC_3, Thirsty = IAS_ACC_4,
    Urinate = IAS_ACC_5, Defecate = IAS_ACC_6, Taste = IAS_ACC_7, Vomit = IAS_ACC_8, Sneeze = IAS_ACC_9,
    Cough = IAS_ACC_10, Temp = IAS_ACC_11, Sex_arousal = IAS_ACC_12, Wind = IAS_ACC_13, Burp = IAS_ACC_14,
    Muscles = IAS_ACC_15, Bruise = IAS_ACC_16, Pain = IAS_ACC_17, Blood_Sugar = IAS_ACC_18,
    Affective_touch = IAS_ACC_19, Tickle =IAS_ACC_21, Itch = IAS_ACC_22
  ) |>
   mutate(Gender = case_when(
    Gender == 1 ~ "Male",
    Gender == 2 ~ "Female",
    Gender %in% c(3, 5) ~ "Other",
    Gender == 4 ~ "NA"
  ))

df11b <- haven::read_sav("../data/Giulia/Sleep and Exteroceptive Interoceptive Sensitivity.sav") |> # 165 participants
  filter(Total_Attention_Fail == 0) |> # 32 participants removed based on failing one or more checks
  select(AGE, GENDER, contains("IAS_ACC")) |> 
  mutate_all(as.numeric) |> 
  rename( Age = AGE, Gender = GENDER, 
          Heart = IAS_ACC_1, Hungry = IAS_ACC_2, Breathing = IAS_ACC_3, Thirsty = IAS_ACC_4, Urinate = IAS_ACC_5, Defecate = IAS_ACC_6, Taste = IAS_ACC_7,
          Vomit = IAS_ACC_8, Sneeze = IAS_ACC_9, Cough = IAS_ACC_10, Temp = IAS_ACC_11, Sex_arousal = IAS_ACC_12, Wind = IAS_ACC_13, Burp = IAS_ACC_14,
          Muscles = IAS_ACC_15, Bruise = IAS_ACC_16, Pain = IAS_ACC_17, Blood_Sugar = IAS_ACC_18, Affective_touch = IAS_ACC_19, Tickle =IAS_ACC_20, 
          Itch = IAS_ACC_21 )  |> 
  mutate(Gender = case_when(
    Gender == 1 ~ "Male", 
    Gender == 2 ~"Female", 
    Gender %in% c(3, 5) ~ "Other", 
    Gender == 4 ~ "NA", ))

# Save all
data <- list(df1a=df1a, df1b=df1b, df2=df2, df3=df3, df4=df4, df5=df5, df6=df6, df7a=df7a, df7b=df7b, df7c=df7c, df8a=df8a, df8b=df8b, df9=df9, df10a=df10a, df10b=df10b, df10c=df10c, df11a=df11a, df11b=df11b)
save(data, file = "../data/data.RData")
```

:::

## Participants

- **Sample 1a**: Data from [Murphy's (2020)](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769) study 1, downloaded from OSF, included `r report::report_participants(df1a, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 1b**: Data from [Murphy's (2020)](https://osf.io/3m5nh/?view_only=a68051df4abe4ecb992f22dc8c17f769) study 6, downloaded from OSF, included `r report::report_participants(df1b, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 2**: Data from [Gaggero's(2020)](https://osf.io/5x9sg/) study, downloaded from OSF, included `r report::report_participants(df2[c("Age", "Gender")], age = "Age", sex = NA, gender = "Gender")`.
- **Sample 3**: Data from [Campos's(2022)](https://osf.io/j6ef3) study, downloaded from OSF, included `r report::report_participants(df3, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 4**: Data from [Todd's(2022)]( https://osf.io/ms354/) study, downloaded from OSF, included `r report::report_participants(df4, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 5**: Data from [Arslanova (2022)](https://osf.io/mp3cy/) study, downloaded from OSF, included `r report::report_participants(df5_attention, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 6**: Data from [Brand's(2022)](https://osf.io/xwz6g/) study, downloaded from OSF, included `r report::report_participants(df6, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 7a**: Data from [Brand's(2023)](https://osf.io/3f2h6/) study, downloaded from OSF, included `r report::report_participants(df7a, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 7b**: Data from [Brand's(2023)](https://osf.io/3f2h6/) study, downloaded from OSF, included `r report::report_participants(df7b, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 7c**: Data from [Brand's(2023)](https://osf.io/3f2h6/) study, downloaded from OSF, included `r report::report_participants(df7c, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 8a**: Data from [Lin's(2023)](https://osf.io/3eztd) study, downloaded from OSF, included `r report::report_participants(df8a, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 8b**: Data from [Lin's(2023)](https://osf.io/3eztd) study, downloaded from OSF, included `r report::report_participants(df8b, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 9**: Data from [VonMohr's(2023)](https://osf.io/7p9u5/) study 3, downloaded from OSF, included `r report::report_participants(df9, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 10a**: Data from [Makowski's(2023)] study , downloaded from OSF, included `r report::report_participants(df10a, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 10b**: Data from [Makowski's(2023)] study , downloaded from OSF, included `r report::report_participants(df10b, age = "Age", sex = NA, gender = "Gender")`.
- **Sample 10c**: Data from [Makowski's(2023)] study , downloaded from OSF, included `r report::report_participants(df10c, age = "Age", sex = NA, gender = "Gender")`.
-- **Sample 11a**: Data from [Giulia's (2024)](https://osf.io/49wbv) study , included `r report::report_participants(df11a, age = "Age", sex = NA, gender = "Gender")`
- **Sample 11b**: Data from [Giulia] study , included `r report::report_participants(df11b, age = "Age", sex = NA, gender = "Gender")`

Total N = `r nrow(df1a) + nrow(df1b) + nrow(df2) + nrow(df3) + nrow(df4) + nrow(df5) + nrow(df6) + nrow(df7a) + nrow(df7b) + nrow(df7c) + nrow(df8a) + nrow(df8b) + nrow(df9) + nrow(df10a) + nrow(df10b) + nrow(df10c) +nrow(df11a) + nrow(df11b)`.

```{r}

desc <- function(df){
  summary <- df |> summarise(
      n = n(),
      mean = mean(Age, na.rm = TRUE),
      sd = sd(Age, na.rm = TRUE),
      min = min(Age, na.rm = TRUE),
      max = max(Age, na.rm = TRUE),
      female = sum(Gender == "Female", na.rm = TRUE)
    )
    return(summary)
}

desc_df1a <- desc(df1a)
desc_df1b <- desc(df1b)
desc_df2 <- desc(df2)
desc_df3 <- desc(df3)
#desc_df4 <- desc(df4)
desc_df5 <- desc(df5_attention)
desc_df6 <- desc(df6)
desc_df7a <- desc(df7a)
desc_df7b <- desc(df7b)
desc_df7c <- desc(df7c)
desc_df8a <- desc(df8a)
desc_df8b <- desc(df8b)
desc_df9 <- desc(df9)
desc_df10a <- desc(df10a)
desc_df10b <- desc(df10b)
desc_df10c <- desc(df10c)
desc_df11a <- desc(df11a)
desc_df11b <- desc(df11b)

desc_total <- rbind(desc_df1a, desc_df1b, desc_df2, desc_df3, desc_df5, desc_df6, desc_df7a, desc_df7b, desc_df7c, desc_df8a, desc_df8b, desc_df9, desc_df10a, desc_df10b, desc_df10c, desc_df11a, desc_df11b)

## calculate weighted mean

total_mean = sum(desc_total$mean * desc_total$n) / sum(desc_total$n)
total_sd = sum(desc_total$sd * desc_total$n) / sum(desc_total$n)
perce_female = sum(desc_total$female)/sum(desc_total$n) *100

```

```{r}
#| output: false
#| warning: false

library(gt)

# APA style ####
gt_apastyle <- function(gt_table, font.size=12) {
  gt_table  |> 
    gt::opt_table_lines(extent = "none") %>%
    gt::tab_options(
      heading.border.bottom.width = 2,
      heading.border.bottom.color = "black",
      heading.border.bottom.style = "solid",
      table.border.top.color = "black",
      table.border.top.style = "solid",
      table.border.top.width = 2,  
      table_body.hlines.color = "white",
      table_body.border.top.color = "black",
      table_body.border.top.style = "solid",
      table_body.border.top.width = 2,
      heading.title.font.size = font.size,
      table.font.size = font.size,
      heading.subtitle.font.size = font.size,
      table_body.border.bottom.color = "black",
      table_body.border.bottom.width = 2,
      table_body.border.bottom.style = "solid",
      column_labels.border.bottom.color = "black",
      column_labels.border.bottom.style = "solid",
      column_labels.border.bottom.width = 1
    ) |> 
      gt::opt_table_font(font = "times")
}

make_age <- function(age) {
  age <- as.numeric(age)

  mean(age, na.rm = TRUE) |> 
    insight::format_value(digits=1) |> 
    paste0(" ± ", 
           insight::format_value(sd(age, na.rm = TRUE), digits=1)) 
}

# Create the dataframe
table <- data.frame(
   Sample = c('Murphy et al., (2020)', '', '', 'Gaggero et al., (2021)', 'Campos et al., (2022)', 'Todd et al., (2022)', 'Arslanova et al., (2022)', 'Brand et al., (2022)', 'Brand et al., (2023)', '', '', '', 'Lin et al., (2023)', '', '', 'VonMohr et al., (2023)', 'Makowski et al., (2023a)', 'Makowski et al., (2023b)', 'Makowski et al., (2023c)', 'Poreiro et al., (2024)', 'Poreiro et al., unpublished', 'Total'),
  Subsample = c('', 'Sample 1', 'Sample 2', '', '', '', '', '', '', 'Sample 1', 'Sample 2', 'Sample 3', '', 'Sample 1', 'Sample 2', '', '', '', '', '','', ''),   
  Language = c('', 'English', 'English' , 'English and Italian', 'Portuguese', 'English', 'English', 'German', '', 'German', 'German', 'German', '', 'Chinese', 'Chinese', 'English', 'English', 'English', 'English', 'English', 'English',''),
  N = c('', nrow(df1a), nrow(df1b), nrow(df2), nrow(df3), nrow(df4), nrow(df5_attention), nrow(df6),'', nrow(df7a), nrow(df7b), 802,'', nrow(df8a), nrow(df8b), nrow(df9), nrow(df10a), nrow(df10b), nrow(df10c), nrow(df11a), nrow(df11b), 32216),
  Age = c('', make_age(df1a$Age), make_age(df1b$Age), make_age(df2$Age), make_age(df3$Age), "48.6.6 ± 14.1*", make_age(df5_attention$Age), make_age(df6$Age),'', make_age(df7a$Age), make_age(df7b$Age), make_age(df7c$Age),'', make_age(df8a$Age), make_age(df8b$Age), make_age(df9$Age), make_age(df10a$Age), make_age(df10b$Age), make_age(df10c$Age), make_age(df11a$Age), make_age(df11b$Age), '48.6 ± 13.1'),
  Range = c('', '18-69', '18-91', '18-58', '18-72', '18-92*', '18-73', '18-78', '', '18-79', '16-81', '18-72','', '16-60', '20-60', '18-93', '18-73', '17-76','18-50', '18-57', '18-60', '17-93'),
  Female_Percentage = c('', '69.4%', '70.1%', '60.3%', '59.6%', '50%*', '46.8%', '78.7%', '', '79.5%', '77.7%', '68.9%', '', '57.0%', '56.2%', '73.2%', '50.3%', '53.0%', '76%', '74.8%', '75.9%', '71.6'),
  Availability= c('', '', 'https://osf.io/3m5nh', 'https://osf.io/5x9sg', 'https://osf.io/j6ef3', 'https://osf.io/ms354', 'https://osf.io/mp3cy', 'https://osf.io/xwz6g', '', '', 'https://osf.io/3f2h6', '','','', 'https://osf.io/3eztd', 'https://osf.io/7p9u5', 'https://github.com/RealityBending/IllusionGameReliability', 'https://github.com/DominiqueMakowski/PHQ4R', 'https://github.com/RealityBending/InteroceptionPrimals', 'https://osf.io/49wbv', '','')
)

table_apa <- table |> 
  gt() |>
  cols_align(align = c("right"), columns = "Age") |> 
  cols_label(Age = "Age (Mean  ± SD)", Female_Percentage = "Female %") |> 
  tab_footnote("* Information taken from the sample description of relevant paper rather than recomputed.") |> 
  gt_apastyle()

table_apa
gtsave(table_apa, "figures/table1.tex")
```

## Descriptive

### Distribution

The items with the differing distribution pattern (with a lower mode around 2/5) are "Affective Touch", "Blood Sugar" and "Bruise".

```{r}
vars <- c(
  "Heart", "Hungry", "Breathing", "Thirsty", "Urinate", "Defecate", "Taste", "Vomit", "Sneeze", "Cough", "Temp",
  "Sex_arousal", "Wind", "Burp", "Muscles", "Bruise", "Pain", "Blood_Sugar", "Affective_touch", "Tickle", "Itch"
)

dens1a <- estimate_density(select(df1a, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 1a")
dens1b <- estimate_density(select(df1b, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 1b")
dens2 <- estimate_density(select(df2, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 2")
dens3 <- estimate_density(select(df3, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 3")
dens4 <- estimate_density(select(df4, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 4")
dens5 <- estimate_density(select(df5, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 5")
dens6 <- estimate_density(select(df6, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 6")
dens7a <- estimate_density(select(df7a, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 7a")
dens7b <- estimate_density(select(df7b, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 7b")
dens7c <- estimate_density(select(df7c, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 7c")
dens8a <- estimate_density(select(df8a, all_of(setdiff(vars, "Tickle"))), method = "kernSmooth") |>
  mutate(Sample = "Sample 8a")
dens8b <- estimate_density(select(df8b, all_of(setdiff(vars, "Tickle"))), method = "kernSmooth") |>
  mutate(Sample = "Sample 8b")
dens9 <- estimate_density(select(df9, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 9")
dens10a <- estimate_density(select(df10a, all_of(setdiff(vars, c("Taste", "Cough", "Blood_Sugar")))), method = "kernSmooth") |>
  mutate(Sample = "Sample 10a",
         x = datawizard::rescale(x, to=c(1, 5), range=c(0, 1)))
dens10b <- estimate_density(select(df10b, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 10b",
         x = datawizard::rescale(x, to=c(1, 5), range=c(0, 1)))
dens10c <- estimate_density(select(df10c, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 10c",
         x = datawizard::rescale(x, to=c(1, 5), range=c(0, 1)))
dens11a <- estimate_density(select(df11a, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 11a")
dens11b <- estimate_density(select(df11b, all_of(vars)), method = "kernSmooth") |>
  mutate(Sample = "Sample 11b")

p1 <- rbind(dens1a, dens1b, dens2, dens3, dens4, dens5, dens6, dens7a, dens7b, dens7c, dens8a, dens8b, dens9, dens10a, dens10b, dens10c, dens11a, dens11b) |>
  mutate(Sample = fct_relevel(Sample, "Sample 1a", "Sample 1b", "Sample 2", "Sample 3", "Sample 4", "Sample 5", "Sample 6", "Sample 7a", "Sample 7b", "Sample 7c", "Sample 8a", "Sample 8b", "Sample 9", "Sample 10a", "Sample 10b", "Sample 10c", "Sample 11a", "Sample 11b")) |>
  # mutate(Dashed = ifelse(Parameter %in% c("Affective_touch", "Blood_Sugar", "Bruise"), TRUE, FALSE)) |> 
  ggplot(aes(x = x, y = y, color = Parameter)) +
  geom_line() +
  scale_color_material_d() +
  scale_linetype_manual(values = c("solid", "dashed"), guide="none") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        plot.title = element_text(face="bold")) +
  labs(title = "Distribution of responses for all items across various datasets", x = "Response", y = "Density", color = "Item") +
  facet_wrap(~Sample, scales = "free_y", nrow = 5)

ggsave("figures/Figure1.png", p1, width=10, height=10, dpi=200, bg="white")
p1

```

```{r}
data1a <- normalize(select(df1a, all_of(dens1a$Parameter)), range = c(1, 5))
data1b <- normalize(select(df1b, all_of(dens1b$Parameter)), range = c(1, 5))
data2 <- normalize(select(df2, all_of(dens2$Parameter)), range = c(1, 5))
data3 <- normalize(select(df3, all_of(dens3$Parameter)), range = c(1, 5))
data4 <- normalize(select(df4, all_of(dens4$Parameter)), range = c(1, 5))
data5 <- normalize(select(df5, all_of(dens5$Parameter)), range = c(1, 5))
data6 <- normalize(select(df6, all_of(dens6$Parameter)), range = c(1, 5))
data7a <- normalize(select(df7a, all_of(dens7a$Parameter)), range = c(1, 5))
data7b <- normalize(select(df7b, all_of(dens7b$Parameter)), range = c(1, 5))
data7c <- normalize(select(df7c, all_of(dens7c$Parameter)), range = c(1, 5))
data8a <- normalize(select(df8a, all_of(dens8a$Parameter)), range = c(1, 5))
data8b <- normalize(select(df8b, all_of(dens8b$Parameter)), range = c(1, 5))
data9 <- normalize(select(df9, all_of(dens9$Parameter)), range = c(1, 5))
data10a <- select(df10a, all_of(dens10a$Parameter))
data10b <- select(df10b, all_of(dens10b$Parameter))
data10c <- select(df10c, all_of(dens10c$Parameter))
data11a <- normalize(select(df11a, all_of(dens11a$Parameter)), range = c(1, 5))
data11b <- na.omit(normalize(select(df11b, all_of(dens11b$Parameter)), range = c(1, 5)))
row.names(data11b) <- NULL

data_all <- rbind(
  data1a, data1b, data2, data3, data4, data5, data6, data7a, data7b, data7c,
  mutate(data8a, Tickle = NA), mutate(data8b, Tickle = NA), data9,
  mutate(data10a, Taste = NA, Cough = NA, Blood_Sugar = NA),  data10b,  data10c, data11a, data11b
) 
```

### Correlations

An overall positive intercorrelation pattern, with no clear structure emerging.

::: {.panel-tabset}

#### All data

```{r}
make_correlation <- function(df) {
  correlation::correlation(df, redundant = TRUE) |>
    correlation::cor_sort() |>
    correlation::cor_lower() |>
    mutate(val = paste0(insight::format_value(r), format_p(p, stars_only = TRUE))) |>
    mutate(Parameter2 = fct_rev(Parameter2)) |>
    ggplot(aes(x = Parameter1, y = Parameter2)) +
    geom_tile(aes(fill = r), color = "white") +
    geom_text(aes(label = val), size = 3) +
    labs(title = "Correlation Matrix") +
    scale_fill_metro_c(limit = c(0, 0.75), guide = guide_colourbar(ticks = FALSE)) +
    theme_minimal() +
    theme(
      legend.title = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

make_correlation(data_all)
```

#### Sample 1a

```{r}
make_correlation(data1a)
```

#### Sample 1b

```{r}
make_correlation(data1b)
```

#### Sample 2

```{r}
make_correlation(data2)
```

#### Sample 3

```{r}
make_correlation(data3)
```

#### Sample 4

```{r}
make_correlation(data4)
```

#### Sample 5

```{r}
make_correlation(data5)
```

#### Sample 6

```{r}
make_correlation(data6)
```

#### Sample 7a

```{r}
make_correlation(data7a)
```

#### Sample 7b

```{r}
make_correlation(data7b)
```

#### Sample 7c

```{r}
make_correlation(data7c)
```

#### Sample 8a

```{r}
make_correlation(data8a)
```

#### Sample 8b

```{r}
make_correlation(data8b)
```

#### Sample 9

```{r}
make_correlation(data9)
```

#### Sample 10a

```{r}
make_correlation(data10a)
```

#### Sample 10b

```{r}
make_correlation(data10b)
```

#### Sample 10c

```{r}
make_correlation(data10c)
```

#### Sample 11a

```{r}
make_correlation(data11a)
```

#### Sample 11b

```{r}
make_correlation(data11b)
```

:::

## Graph Analysis

Exploratory Graph Analysis (EGA) is a recently developed framework for psychometric assessment, that can be used to estimate the number of dimensions in questionnaire data using network estimation methods and community detection algorithms, and assess the stability of dimensions and items using bootstrapping.
See https://r-ega.net/articles/ega.html for details.

### Unique Variable Analysis (UVA)

Unique Variable Analysis (Christensen, Garrido, & Golino, 2023) uses the weighted topological overlap measure (Nowick et al., 2009) on an estimated network. Values greater than 0.25 are determined to have considerable local dependence (i.e., redundancy) that should be handled (variables with the highest maximum weighted topological overlap to all other variables (other than the one it is redundant with) should be removed).

::: {.panel-tabset}

#### All data

```{r}
uva <- EGAnet::UVA(data = data_all, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 1a

```{r}
uva <- EGAnet::UVA(data = data1a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 1b

```{r}
uva <- EGAnet::UVA(data = data1b, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 3

```{r}
uva <- EGAnet::UVA(data = data3, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 4

```{r}
uva <- EGAnet::UVA(data = data4, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 5

```{r}
uva <- EGAnet::UVA(data = data5, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 6

```{r}
uva <- EGAnet::UVA(data = data6, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 7a

```{r}
uva <- EGAnet::UVA(data = data7a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 7b

```{r}
uva <- EGAnet::UVA(data = data7b, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 7c

```{r}
uva <- EGAnet::UVA(data = data7c, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 8a

```{r}
uva <- EGAnet::UVA(data = data8a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 8b

```{r}
uva <- EGAnet::UVA(data = data8b, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 9

```{r}
uva <- EGAnet::UVA(data = data9, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 10a

```{r}
uva <- EGAnet::UVA(data = data10a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 10b

```{r}
uva <- EGAnet::UVA(data = data10b, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 10c

```{r}
uva <- EGAnet::UVA(data = data10c, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 11a

```{r}
uva <- EGAnet::UVA(data = data11a, cut.off = 0.3)
uva
uva$keep_remove
```

#### Sample 11b

```{r}
uva <- EGAnet::UVA(data = data11b, cut.off = 0.3)
uva
uva$keep_remove
```

:::

### Networks

::: {.panel-tabset}

#### All data

```{r ega_all}
#| fig-height: 20
#| fig-width: 12

make_egas <- function(data) {
  egas <- list()
  for (model in c("glasso")) {  # , "TMFG"
    for (algo in c("walktrap", "louvain")) {
      for (type in c("ega", "ega.fit")) { # "" # "hierega"
        if (type == "ega.fit" & algo == "louvain") next # Too slow
        egas[[paste0(model, "_", algo, "_", type)]] <- EGAnet::bootEGA(
          data = data,
          seed = 123,
          model = model,
          algorithm = algo,
          EGA.type = type,
          type = "resampling",
          plot.itemStability = FALSE,
          verbose = FALSE
        )
      }
    }
  }

  p <- EGAnet::compare.EGA.plots(
    egas$glasso_walktrap_ega, 
    # egas$glasso_walktrap_ega.fit,
    egas$glasso_louvain_ega, 
    # egas$TMFG_louvain_ega,
    # egas$TMFG_walktrap_ega, 
    # egas$TMFG_walktrap_ega.fit,
    labels = c(
      "glasso_walktrap_ega", 
      # "glasso_walktrap_ega.fit",
      "glasso_louvain_ega" 
      # "TMFG_louvain_ega",
      # "TMFG_walktrap_ega", 
      # "TMFG_walktrap_ega.fit"
    ),
    rows = 3,
    plot.all = FALSE
  )$all
  list(egas = egas, p = p)
}

egas_all <- make_egas(data_all)
egas_all$p
```

#### Sample 1a

```{r ega1a}
#| fig-height: 20
#| fig-width: 12

egas1a <- make_egas(data1a)
egas1a$p
```

#### Sample 1b

```{r ega1b}
#| fig-height: 20
#| fig-width: 12

egas1b <- make_egas(data1b)
egas1b$p
```

#### Sample 2

```{r ega2}
#| fig-height: 20
#| fig-width: 12

egas2 <- make_egas(data2)
egas2$p
```

#### Sample 3

```{r ega3}
#| fig-height: 20
#| fig-width: 12

egas3 <- make_egas(data3)
egas3$p
```

#### Sample 4

```{r ega4}
#| fig-height: 20
#| fig-width: 12

egas4 <- make_egas(data4)
egas4$p
```

#### Sample 5

```{r ega5}
#| fig-height: 20
#| fig-width: 12

egas5 <- make_egas(data5)
egas5$p
```

#### Sample 6

```{r ega6}
#| fig-height: 20
#| fig-width: 12

egas6 <- make_egas(data6)
egas6$p
```

#### Sample 7a

```{r ega7&}
#| fig-height: 20
#| fig-width: 12

egas7a <- make_egas(data7a)
egas7a$p
```

#### Sample 7b

```{r ega7b}
#| fig-height: 20
#| fig-width: 12

egas7b <- make_egas(data7b)
egas7b$p
```

#### Sample 7c

```{r ega7c}
#| fig-height: 20
#| fig-width: 12

egas7c <- make_egas(data7c)
egas7c$p
```

#### Sample 8a

```{r ega8a}
#| fig-height: 20
#| fig-width: 12

egas8a <- make_egas(data8a)
egas8a$p
```

#### Sample 8b

```{r ega8b}
#| fig-height: 20
#| fig-width: 12

egas8b <- make_egas(data8b)
egas8b$p
```

#### Sample 9

```{r ega9}
#| fig-height: 20
#| fig-width: 12

egas9 <- make_egas(data9)
egas9$p
```

#### Sample 10a

```{r ega10a}
#| fig-height: 20
#| fig-width: 12

egas10a <- make_egas(data10a)
egas10a$p
```

#### Sample 10b

```{r ega10b}
#| fig-height: 20
#| fig-width: 12

egas10b <- make_egas(data10b)
egas10b$p
```

#### Sample 10c

```{r ega10c}
#| fig-height: 20
#| fig-width: 12

egas10c <- make_egas(data10c)
egas10c$p
```

#### Sample 11a

```{r ega11a}
#| fig-height: 20
#| fig-width: 12

egas11a <- make_egas(data11a)
egas11a$p
```

#### Sample 11b

```{r ega11c}
#| fig-height: 20
#| fig-width: 12

egas11b <- make_egas(data11b)
egas11b$p
```

:::

### Structure Stability

::: {.panel-tabset}

#### All data

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas_all$egas, plot), nrow = 3)
```

#### Sample 1a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas1a$egas, plot), nrow = 3)
```

#### Sample 1b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas1b$egas, plot), nrow = 3)
```

#### Sample 2

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas2$egas, plot), nrow = 3)
```

#### Sample 3

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas3$egas, plot), nrow = 3)
```

#### Sample 4

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas4$egas, plot), nrow = 3)
```

#### Sample 5

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas5$egas, plot), nrow = 3)
```

#### Sample 6

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas6$egas, plot), nrow = 3)
```

#### Sample 7a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas7a$egas, plot), nrow = 3)
```

#### Sample 7b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas7b$egas, plot), nrow = 3)
```

#### Sample 7c

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas7c$egas, plot), nrow = 3)
```

#### Sample 8a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas8a$egas, plot), nrow = 3)
```

#### Sample 8b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas8b$egas, plot), nrow = 3)
```

#### Sample 9

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas9$egas, plot), nrow = 3)
```

#### Sample 10a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas10a$egas, plot), nrow = 3)
```

#### Sample 10b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas10b$egas, plot), nrow = 3)
```

#### Sample 10c

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas10c$egas, plot), nrow = 3)
```

#### Sample 11a

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas11a$egas, plot), nrow = 3)
```

#### Sample 11b

```{r}
#| fig-height: 20
#| fig-width: 12

patchwork::wrap_plots(lapply(egas11b$egas, plot), nrow = 3)
```

:::

### Scores

```{r}
ega <- egas_all$egas$glasso_louvain_ega$EGA
plot(ega)

get_scores <- function(data, ega) {
  EGAnet::net.scores(data, ega)$scores$std.scores |>
    as.data.frame() |>
    setNames(c("EGA1", "EGA2", "EGA3", "EGA4"))
}

ega_scores_1a <- get_scores(data1a, ega)
ega_scores_1b <- get_scores(data1b, ega)
ega_scores_2 <- get_scores(data2, ega)
ega_scores_3 <- get_scores(data3, ega)
ega_scores_4 <- get_scores(data4, ega)
ega_scores_5 <- get_scores(data5, ega)
ega_scores_6 <- get_scores(data6, ega)
ega_scores_7a <- get_scores(data7a, ega)
ega_scores_7b <- get_scores(data7b, ega)
ega_scores_7c <- get_scores(data7c, ega)
ega_scores_8a <- get_scores(data8a, ega)
ega_scores_8b <- get_scores(data8b, ega)
ega_scores_9 <- get_scores(data9, ega)
ega_scores_10a <- get_scores(data10a, ega)
ega_scores_10b <- get_scores(data10b, ega)
ega_scores_10c <- get_scores(data10c, ega)
ega_scores_11a <- get_scores(data11a, ega)
ega_scores_11b <- na.omit(get_scores(data11b, ega)) # remove two rows with NAS
row.names(ega_scores_11b) <- NULL
```

## Factor Analysis

### How Many Factors

::: {.panel-tabset}

#### All data

```{r}
n <- parameters::n_factors(data_all, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 1a

```{r}
n <- parameters::n_factors(data1a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 1b

```{r}
n <- parameters::n_factors(data1b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 2

```{r}
n <- parameters::n_factors(data2, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 3

```{r}
n <- parameters::n_factors(data3, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 4

```{r}
n <- parameters::n_factors(data4, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 5

```{r}
n <- parameters::n_factors(data5, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 6

```{r}
n <- parameters::n_factors(data6, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 7a

```{r}
n <- parameters::n_factors(data7a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 7b

```{r}
n <- parameters::n_factors(data7b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 7c

```{r}
n <- parameters::n_factors(data7c, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 8a

```{r}
n <- parameters::n_factors(data8a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 8b

```{r}
n <- parameters::n_factors(data8b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 9

```{r}
n <- parameters::n_factors(data9, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 10a

```{r}
n <- parameters::n_factors(data10a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 10b

```{r}
n <- parameters::n_factors(data10b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 10c

```{r}
n <- parameters::n_factors(data10c, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 11a

```{r}
n <- parameters::n_factors(data11a, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```

#### Sample 11b

```{r}
n <- parameters::n_factors(data11b, n_max = 12)

plot(n)
knitr::kable(as.data.frame(n)[n$n_Factors %in% c(1, 3, 4, 5), ], row.names = FALSE)
```
:::

### EFA

::: {.panel-tabset}

#### All data

```{r}
#| fig-height: 10

efa5 <- parameters::factor_analysis(data_all, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa5)

display(efa5)
```

#### Sample 1a

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data1a, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 1b

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data1b, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 2

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data2, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)

```

#### Sample 3

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data3, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### sample 4

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data4, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 5

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data5, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 6

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data6, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 7a

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data7a, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 7b

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data7b, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 7c

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data7c, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 8a

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data8a, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 8b

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data8b, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 9

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data9, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 10a

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data10a, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 10b

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data10b, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 10c

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data10c, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 11a

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data11a, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

#### Sample 11b

```{r}
#| fig-height: 10

efa4 <- parameters::factor_analysis(data11b, n = 4, rotation = "oblimin", sort = TRUE)
plot(efa4)

display(efa4)
```

:::

## CFA

We discarded the following items:
- Tickle: not present in the same dataset and consistently flagged as redundant.
<!-- - Blood_Sugar -->

Ambiguous:
- Temp
- Vomit
- Affective_touch
- Sex_arousal
- Taste

### Lower-Level Factors

::: {.panel-tabset}

#### All data

```{r}
m1 <- "
Interoception =~ Hungry + Thirsty + Urinate + Defecate + Itch + Bruise + Muscles + Pain + Breathing + Heart + Cough + Sneeze + Wind + Burp
"

m4 <- "
Sustenance =~ Hungry + Thirsty + Urinate + Defecate + Muscles + Pain
Skin =~ Itch + Bruise
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m5 <- "
Sustenance =~ Hungry + Thirsty + Pain
UrinateDefecate =~ Urinate + Defecate
Skin =~ Itch + Bruise + Muscles
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m6 <- "
HungryThirsty =~ Hungry + Thirsty
UrinateDefecate =~ Urinate + Defecate
ItchBruise =~ Itch + Bruise
MusclesPain =~ Muscles + Pain
HeartBreathing =~ Breathing + Heart
Expulsion =~ Cough + Sneeze + Wind + Burp
"

m7 <- "
HungryThirsty =~ Hungry + Thirsty
UrinateDefecate =~ Urinate + Defecate
ItchBruise =~ Itch + Bruise
MusclesPain =~ Muscles + Pain
HeartBreathing =~ Breathing + Heart
CoughSneeze =~ Cough + Sneeze
WindBurp =~ Wind + Burp
"

is_converged <- function(m) {
  ev <- eigen(lavaan::lavTech(m, "cov.lv")[[1]], symmetric=TRUE, only.values=TRUE)$values
  if (any(ev < -1 * .Machine$double.eps^(3/4))) {
    "Not converged"
  } else {
    ""
  }
}

compare_mods <- function(data) {

  fit_model <- function(f, data) {
    if(!"Cough" %in% names(data)) {
      f <- str_remove(f, fixed("Cough +"))
    }
    lavaan::cfa(f, data = data)
  }

  cfa1 <- fit_model(m1, data = data)
  cfa4 <- fit_model(m4, data = data)
  cfa5 <- fit_model(m5, data = data)
  cfa6 <- fit_model(m6, data = data)
  cfa7 <- fit_model(m7, data = data)

  anova(cfa1, cfa4, cfa5, cfa6, cfa7) |> 
    parameters::parameters() |>
    mutate(AIC = format(round(AIC, 0), nsmall = 0),
           BIC = format(round(BIC, 0), nsmall = 0),
           Chi2 = format(round(Chi2, 2), nsmall = 2),
           Converged = sapply(list(cfa1, cfa4, cfa5, cfa6, cfa7), is_converged))
}

display(compare_mods(data_all))

m7_all <- lavaan::cfa(m7, data = data_all)
parameters::parameters(m7_all, standardize=TRUE) |>
  display()
```

#### Sample 1a

```{r}
#| warning: false

display(compare_mods(data1a))
```

#### Sample 1b

```{r}
#| warning: false

display(compare_mods(data1b))

```

#### Sample 2

```{r}
#| warning: false

display(compare_mods(data2))

```

#### Sample 3

```{r}
#| warning: false

display(compare_mods(data3))

```

#### Sample 4

```{r}
#| warning: false

display(compare_mods(data4))

```

#### Sample 5

```{r}
#| warning: false

display(compare_mods(data5))

```

#### Sample 6

```{r}
#| warning: false

display(compare_mods(data6))
```

#### Sample 7a

```{r}
#| warning: false

display(compare_mods(data7a))
```

#### Sample 7b

```{r}
#| warning: false

display(compare_mods(data7b))

```

#### Sample 7c

```{r}
#| warning: false

display(compare_mods(data7c))

```

#### Sample 8a

```{r}
#| warning: false

display(compare_mods(data8a))
```

#### Sample 8b

```{r}
#| warning: false

display(compare_mods(data8b))
```

#### Sample 9

```{r}
#| warning: false

display(compare_mods(data9))

```

#### Sample 10a

```{r}
#| warning: false

display(compare_mods(data10a))

```

#### Sample 10b

```{r}
#| warning: false

display(compare_mods(data10b))

```

#### Sample 10c

```{r}
#| warning: false

display(compare_mods(data10c))

```

#### Sample 11a

```{r}
#| warning: false

display(compare_mods(data11a))

```

#### Sample 11b

```{r}
#| warning: false

display(compare_mods(data11b))

```

:::

Conclusion: The 7th factor solution was preferred in most datasets, including by indices that penalize increased number of parameters (such as BIC).

### Higher Order Factors

::: {.panel-tabset}

#### All data

```{r}
m7h1 <- paste0(m7, "
              Interoception =~ HungryThirsty + UrinateDefecate + ItchBruise + MusclesPain + HeartBreathing + CoughSneeze + WindBurp
              ")

m7h2 <- paste0(m7, "
              Expulsion =~ CoughSneeze + WindBurp
              Interoception =~ HungryThirsty + UrinateDefecate + HeartBreathing
              ")

m7h3 <- paste0(m7, "
              Valenced =~ MusclesPain + HeartBreathing + ItchBruise
              Expulsion =~ CoughSneeze + WindBurp
              Homeostasis =~ HungryThirsty + UrinateDefecate
              ")

compare_mods_h <- function(data) {

  fit_model_h <- function(f, data) {
    if(!"Cough" %in% names(data)) {
      f <- str_remove(f, fixed("Cough +"))
    }
    lavaan::cfa(f, data = data)
  }

  m7 <- fit_model_h(m7, data = data)
  m7h2 <- fit_model_h(m7h2, data = data)
  m7h1 <- fit_model_h(m7h1, data = data)
  m7h3 <- fit_model_h(m7h3, data = data)

  anova(m7, m7h1, m7h2, m7h3) |>
    parameters::parameters() |>
    mutate(AIC = format(round(AIC, 0), nsmall = 0),
           BIC = format(round(BIC, 0), nsmall = 0),
           Chi2 = format(round(Chi2, 2), nsmall = 2),
           Converged = sapply(list(m7, m7h1, m7h2, m7h3), is_converged))
}

display(compare_mods_h(data_all))
```

No evidence for higher order factors.

#### Sample 1a

```{r}
display(compare_mods_h(data1a))
```

#### Sample 1b

```{r}
display(compare_mods_h(data1b))
```

#### Sample 2

```{r}
display(compare_mods_h(data2))
```

#### Sample 3

```{r}
display(compare_mods_h(data3))
```

#### Sample 4

```{r}
display(compare_mods_h(data4))
```

#### Sample 5

```{r}
display(compare_mods_h(data5))
```

#### Sample 6

```{r}
display(compare_mods_h(data6))
```

#### Sample 7a

```{r}
display(compare_mods_h(data7a))
```

#### Sample 7b

```{r}
display(compare_mods_h(data7b))
```

#### Sample 7c

```{r}
display(compare_mods_h(data7c))
```

#### Sample 8a

```{r}
display(compare_mods_h(data8a))
```

#### Sample 8b

```{r}
display(compare_mods_h(data8b))
```

#### Sample 9

```{r}
display(compare_mods_h(data9))
```

#### Sample 10a

```{r}
display(compare_mods_h(data10a))
```

#### Sample 10b

```{r}
display(compare_mods_h(data10b))
```

#### Sample 10c

```{r}
display(compare_mods_h(data10c))
```

#### Sample 11a

```{r}
display(compare_mods_h(data11a))
```

#### Sample 11b

```{r}
display(compare_mods_h(data11b))
```

:::

### Model Performance

```{r}
#| warning: false

rbind(
  performance::performance(m7_all, metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "All"),
  performance::performance(update(m7_all, data = data1a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 1a"),
  performance::performance(update(m7_all, data = data1b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 1b"),
  performance::performance(update(m7_all, data = data2), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 2"),
  performance::performance(update(m7_all, data = data3), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 3"),
  performance::performance(update(m7_all, data = data4), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 4"),
  performance::performance(update(m7_all, data = data5), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 5"),
  performance::performance(update(m7_all, data = data6), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 6"),
  performance::performance(update(m7_all, data = data7a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 7a"),
  performance::performance(update(m7_all, data = data7b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 7b"),
  performance::performance(update(m7_all, data = data7c), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 7c"),
  performance::performance(update(m7_all, data = data8a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 8a"),
  performance::performance(update(m7_all, data = data8b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 8b"),
  performance::performance(update(m7_all, data = data9), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 9"),
  performance::performance(lavaan::cfa(str_remove(m7, fixed("\nCoughSneeze =~ Cough + Sneeze")), data = data10a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 10a"),
  performance::performance(update(m7_all, data = data10b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 10b"),
  performance::performance(update(m7_all, data = data10c), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 10c"),
  performance::performance(update(m7_all, data = data11a), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 11a"),
  performance::performance(update(m7_all, data = data11b), metrics = c("Chi2", "RMSEA", "CFI", "SRMR")) |>
    mutate(Sample = "Sample 11b")
) |>
  display()
```

## Scores

Note the usage of descriptive factor names relating directly to the items to avoid abstraction.

```{r}
#| warning: false

# Add empirical variables
add_empirical <- function(data, df) {
  x <- data |>
    mutate(
      HungryThirsty = (df$Hungry + df$Thirsty) / 2,
      UrinateDefecate = (df$Urinate + df$Defecate) / 2,
      MusclesPain = (df$Muscles + df$Pain) / 2,
      HeartBreathing = (df$Heart + df$Breathing) / 2,
      ItchBruise = (df$Itch + df$Bruise) / 2,
      WindBurp = (df$Wind + df$Burp) / 2)
  if("Cough" %in% names(df)) {
    mutate(x, CoughSneeze = (df$Cough + df$Sneeze) / 2)
  } else {
    mutate(x, CoughSneeze = df$Sneeze)
  }
}

# Deal with missing data
temp <- predict(m7_all, newdata=data7c)
scores7c <- data.frame(matrix(ncol = 7, nrow = nrow(ega_scores_7c))) 
scores7c[!is.na(ega_scores_7c$EGA1), ] <- temp
names(scores7c) <- names(as.data.frame(temp))

scores <- list(
  sample1a = cbind(ega_scores_1a, data_addprefix(as.data.frame(predict(m7_all, newdata=data1a)), "CFA_")) |>
    add_empirical(data1a) |> 
    mutate(Sample = "Sample 1a"),
  sample1b = cbind(ega_scores_1b, data_addprefix(as.data.frame(predict(m7_all, newdata=data1b)), "CFA_")) |>
    add_empirical(data1b) |> 
    mutate(Sample = "Sample 1b"),
  sample2 = cbind(ega_scores_2, data_addprefix(as.data.frame(predict(m7_all, newdata=data2)), "CFA_")) |>
    add_empirical(data2) |> 
    mutate(Sample = "Sample 2"),
  sample3 = cbind(ega_scores_3, data_addprefix(as.data.frame(predict(m7_all, newdata=data3)), "CFA_")) |>
    add_empirical(data3) |> 
    mutate(Sample = "Sample 3"),
  sample4 = cbind(ega_scores_4, data_addprefix(as.data.frame(predict(m7_all, newdata=data4)), "CFA_")) |>
    add_empirical(data4) |> 
    mutate(Sample = "Sample 4"),
  sample5 = cbind(ega_scores_5, data_addprefix(as.data.frame(predict(m7_all, newdata=data5)), "CFA_")) |>
    add_empirical(data5) |> 
    mutate(Sample = "Sample 5"),
  sample6 = cbind(ega_scores_6, data_addprefix(as.data.frame(predict(m7_all, newdata=data6)), "CFA_")) |>
    add_empirical(data6) |> 
    mutate(Sample = "Sample 6"),
  sample7a = cbind(ega_scores_7a, data_addprefix(as.data.frame(predict(m7_all, newdata=data7a)), "CFA_")) |>
    add_empirical(data7a) |> 
    mutate(Sample = "Sample 7a"),
  sample7b = cbind(ega_scores_7b, data_addprefix(as.data.frame(predict(m7_all, newdata=data7b)), "CFA_")) |>
    add_empirical(data7b) |> 
    mutate(Sample = "Sample 7b"),
  sample7c = cbind(ega_scores_7c, data_addprefix(scores7c, "CFA_")) |> 
    add_empirical(data7c) |> 
    mutate(Sample = "Sample 7c"),
  sample8a = cbind(ega_scores_8a, data_addprefix(as.data.frame(predict(m7_all, newdata=data8a)), "CFA_")) |>
    add_empirical(data8a) |> 
    mutate(Sample = "Sample 8a"),
  sample8b = cbind(ega_scores_8b, data_addprefix(as.data.frame(predict(m7_all, newdata=data8b)), "CFA_")) |>
    add_empirical(data8b) |> 
    mutate(Sample = "Sample 8b"),
  sample9 = cbind(ega_scores_9, data_addprefix(as.data.frame(predict(m7_all, newdata=data9)), "CFA_")) |>
    add_empirical(data9) |> 
    mutate(Sample = "Sample 9"),
  sample10a = cbind(ega_scores_10a, data_addprefix(as.data.frame(predict(lavaan::cfa(str_remove(m7, fixed("\nCoughSneeze =~ Cough + Sneeze")), data = data_all), newdata=data10a)), "CFA_")) |>
    add_empirical(data10a) |> 
    mutate(Sample = "Sample 10a", CFA_CoughSneeze = NA),
  sample10b = cbind(ega_scores_10b, data_addprefix(as.data.frame(predict(m7_all, newdata=data10b)), "CFA_")) |>
    add_empirical(data10b) |> 
    mutate(Sample = "Sample 10b"),
  sample10c = cbind(ega_scores_10c, data_addprefix(as.data.frame(predict(m7_all, newdata=data10c)), "CFA_")) |>
    add_empirical(data10c) |> 
    mutate(Sample = "Sample 10c"),
  sample11a = cbind(ega_scores_11a, data_addprefix(as.data.frame(predict(m7_all, newdata=data11a)), "CFA_")) |>
    add_empirical(data11a) |> 
    mutate(Sample = "Sample 11a"),
  sample11b = cbind(ega_scores_11b, data_addprefix(as.data.frame(predict(m7_all, newdata=data11b)), "CFA_")) |>
    add_empirical(data11b) |> 
    mutate(Sample = "Sample 11b")
)

save(scores, file = "../data/scores.RData")
```