### 2025-02-10 analysis/figures/table2.tex
updated manuscript and analysis script

\begin{table}
\fontsize{6.8pt}{8.1pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lrlrr}
\toprule
Questionnaire & Number of Dimensions & Assessment & Number of Items & Scoring \\ 
\midrule\addlinespace[2.5pt]
{\bfseries Interoceptive Related} &  &  &  &  \\ 
MAIA-2 & 8 & Interoception & 37 & 6-point Likert scale \\ 
BPQ & 2 & Body awareness and autonomic reactivity & 49 & 5-point Likert scale \\ 
TAS-20 & 3 & Alexithymia & 20 & 5-point Likert scale \\ 
BVAQ & 2 & Alexithymia & 40 & 5-point Likert scale \\ 
{\bfseries Mood} &  &  &  &  \\ 
BDI-II & 1 & Severity of depressive symptoms & 21 & 0-to-3-point values \\ 
PHQ-4 & 2 & Anxiety and depressive symptoms & 4 & 5-point Likert scale \\ 
STAI-T & 1 & Trait anxiety & 20 & 4-point Likert scale \\ 
GAD-2 & 1 & General Anxiety & 2 & 4-point Likert scale \\ 
{\bfseries Personality} &  &  &  &  \\ 
NEO-FFI & 1 & Neuroticism & 12 & 5-point Likert scale \\ 
Mini IPIP6 & 6 & Personality & 24 & Analogue scales \\ 
BFI & 5 & Personality & 44 & 5-point Likert scale \\ 
PID-5-SF & 5 & Dysfunctional personality traits & 25 & 4-point Likert scale \\ 
{\bfseries Psychopathology} &  &  &  &  \\ 
SPQ-BRU & 4 & Schizotypy & 32 & 5-point Likert scale \\ 
MSI-BPD & 1 & Borderline personality disorder & 10 & Analogue scales \\ 
ASQ - Short & 5 & Autistic Traits & 28 & 4-point Likert scale \\ 
{\bfseries Beliefs and Misbeliefs} &  &  &  &  \\ 
GCB & 5 & Conspiracy beliefs & 15 & 5-point Likert scale \\ 
PI-18 & 1 & Beliefs about the world & 99 & 6-point Likert scale \\ 
LIE scale & 4 & Lying tendencies & 16 & Visual analogue scales \\ 
\bottomrule
\end{tabular*}
\end{table}