### 2025-03-17 analysis/figures/table1.tex
major changes to manuscript
@DominiqueMakowski

please check the latest changes (lot has been done till end of result section on study 1)

\begin{table}[!t]
\fontsize{6.8pt}{8.1pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lllrlrrrl}
\toprule
Sample & Subsample & Language & N & Difference & Age (Mean  ± SD) & Range & Female \% & Availability \\ 
\midrule\addlinespace[2.5pt]
Murphy et al., (2020) &  &  &  &  &  &  &  & osf.io/3m5nh \\ 
 & Sample 1 & English & 451 &  & 25.8 ± 8.4 & 18-69 & 69.4\% &  \\ 
 & Sample 2 & English & 375 &  & 35.3 ± 16.9 & 18-91 & 70.1\% &  \\ 
Gaggero et al., (2021) &  & English and Italian & 814 &  & 24.9 ± 5.3 & 18-58 & 60.3\% & osf.io/5x9sg \\ 
Campos et al., (2022) &  & Portuguese & 515 &  & 30.7 ± 10.5 & 18-72 & 59.6\% & osf.io/j6ef3 \\ 
Todd et al., (2022) &  & English & 802 &  & 48.6.6 ± 14.1* & 18-92* & 50\%* & osf.io/ms354 \\ 
Arslanova et al., (2022) &  & English & 143 &  & 28.5 ± 7.6 & 18-73 & 46.8\% & osf.io/mp3cy \\ 
Brand et al., (2022) &  & German & 619 &  & 43.9 ± 14.5 & 18-78 & 78.7\% & osf.io/xwz6g \\ 
Brand et al., (2023) &  &  &  &  &  &  &  & osf.io/3f2h6 \\ 
 & Sample 1 & German & 522 &  & 23.4 ± 6.7 & 18-79 & 79.5\% &  \\ 
 & Sample 2 & German & 1993 &  & 32.0 ± 12.6 & 16-81 & 77.7\% &  \\ 
 & Sample 3 & German & 802 &  & 27.3 ± 9.3 & 18-72 & 68.9\% &  \\ 
Lin et al., (2023) &  &  &  &  &  &  &  & osf.io/3eztd \\ 
 & Sample 1 & Chinese & 1166 & Collapsed "Itch" and "Tingling" & 32.5 ± 8.4 & 16-60 & 57.0\% &  \\ 
 & Sample 2 & Chinese & 500 & Collapsed "Itch" and "Tingling" & 37.4 ± 7.4 & 20-60 & 56.2\% &  \\ 
VonMohr et al., (2023) &  & English & 21843 &  & 56.5 ± 14.4 & 18-93 & 73.2\% & osf.io/7p9u5 \\ 
Makowski et al., (2023a) &  & English & 485 & Analog scales & 30.1 ± 10.1 & 18-73 & 50.3\% & github.com/RealityBending/IllusionGameReliability \\ 
Makowski et al., (2023b) &  & English & 836 & Analog scales & 25.1 ± 11.3 & 17-76 & 53.0\% & github.com/DominiqueMakowski/PHQ4R \\ 
Makowski et al., (2023c) &  & English & 146 & Analog scales & 21.1 ± 4.3 & 18-50 & 76\% & github.com/RealityBending/InteroceptionPrimals \\ 
Poerio et al., (2024) &  & English & 107 &  & 26.8 ± 9.2 & 18-57 & 74.8\% & osf.io/49wbv \\ 
Poerio et al., unpublished &  & English & 131 &  & 30.9 ± 12.0 & 18-60 & 75.9\% &  \\ 
Total &  &  & 32214 &  & 48.6 ± 13.1 & 17-93 & 71.6\% &  \\ 
\bottomrule
\end{tabular*}
\begin{minipage}{\linewidth}
* Information taken from the sample description of relevant paper rather than recomputed.\\
\end{minipage}
\end{table}