### 2025-08-27 analysis/figures/table2.tex
table 2 updated
need to add references

\begin{table}[t]
\fontsize{6.8pt}{8.1pt}\selectfont
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}lrlrrl}
\toprule
Questionnaire & Number of Dimensions & Assessment & Number of Items & Scoring & Example\_Items \\ 
\midrule\addlinespace[2.5pt]
{\bfseries Interoceptive Related} &  &  &  &  &  \\ 
MAIA-2 & 8 & Interoceptive awareness & 37 & 6-point Likert scale & I notice when I am uncomfortable in my body \\ 
BPQ-SF & 2 & Body awareness and autonomic reactivity & 46 & 5-point Likert scale & My heart often beats irregularily \\ 
BPQ-VSF & 1 & Body awareness & 12 & 5-point Likert scale & My mouth being dry \\ 
ICQ & 3 & Interoceptive confusion & 20 & 5-point Likert scale &  I cannot tell when my muscles are sore or tight \\ 
TAS-20 & 3 & Alexithymia & 20 & 5-point Likert scale & I have feelings I can't quite identify \\ 
BVAQ & 5 & Alexithymia & 40 & 5-point Likert scale & I like to tell others how I feel \\ 
{\bfseries Mood} &  &  &  &  &  \\ 
BDI-II & 2 & Depressive symptoms severity & 21 & 4-point scale & I feel sad much of the time \\ 
PHQ-4 & 2 & Depression and Anxiety & 4 & 4-point Likert scale & Little interest or pleasure in doing things \\ 
PHQ-9 & 1 & Depressive symptoms & 9 & 4-point Likert scale & Feeling tired or having little energy? \\ 
PHQ-15 & 1 & Somatic symptom distress & 15 & 3-point scale & Over the last week, how often have you been bothered by back pain? \\ 
MFQ & 1 & Mood and affect & 13 & 3-point scale & I felt miserable or unhappy \\ 
STAI-T & 1 & Trait anxiety & 20 & 4-point Likert scale & I worry too much about something that really doesn't matter \\ 
GAD-7 & 1 & General anxiety disorder symptoms & 7 & 4-point Likert scale & Not being able to stop or control worrying \\ 
{\bfseries Personality} &  &  &  &  &  \\ 
NEO-FFI (neuroticism subscale) & 1 & Neuroticism & 12 & 5-point Likert scale & I often feel inferior to others \\ 
Mini IPIP6 & 6 & Personality traits & 24 & Visual analogue scales & Am the life of the party \\ 
BFI-S & 5 & Personality traits & 10 & 7-point Likert scale & I see myself as someone who ...is considerate and kind to almost everyone \\ 
{\bfseries Psychopathology} &  &  &  &  &  \\ 
PID-5-SFF & 5 & Maladaptive personality traits & 25 & 4-point Likert scale & Plenty of people are out to get me \\ 
SPQ-BR & 9 & Schizotypy & 32 & 5-point Likert scale & I am an odd, unusual person \\ 
MSI-BPD & 1 & Borderline personality disorder & 9 & Dichotomous (yes/no) & Have any of your closest relationships been troubled by a lot of arguments or repeated breakups? \\ 
ASQ - Short & 5 & Autistic traits & 28 & 4-point Likert scale & I find it difficult to work out people’s intentions \\ 
{\bfseries Beliefs and Misbeliefs} &  &  &  &  &  \\ 
GCBS & 5 & Conspiracy beliefs & 15 & 5-point Likert scale & Secret organizations communicate with extraterrestrials, but keep this fact from the public \\ 
PI-18 & 3 & Primal world beliefs & 18 & 5-point Likert scale & Nearly everything in the world is beautiful \\ 
LIE scale & 4 & Lying tendencies & 16 & Visual analogue scales & I am a good liar \\ 
\bottomrule
\end{tabular*}
\end{table}