### 2025-10-20 analysis/study2.qmd
Update study2.qmd

---
title: "Meta-analysis of the Interoceptive Accuracy Scale"
subtitle: "Study 2"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    fig-dpi: 300
    fig-width: 11.33
execute: 
  cache: true
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
```

## Measures used

- Note, GAD2 and PHQ2 were collapsed into phq4
- PI-99 was collapsed into PI-18
- STait-5 collapsed into STait-t

```{r}
#| output: false

library(gt)

# APA style ####
gt_apastyle <- function(gt_table, font.size=12) {
  gt_table  |> 
    gt::opt_table_lines(extent = "none") %>%
    gt::tab_options(
      heading.border.bottom.width = 2,
      heading.border.bottom.color = "black",
      heading.border.bottom.style = "solid",
      table.border.top.color = "black",
      table.border.top.style = "solid",
      table.border.top.width = 2,  
      table_body.hlines.color = "white",
      table_body.border.top.color = "black",
      table_body.border.top.style = "solid",
      table_body.border.top.width = 2,
      heading.title.font.size = font.size,
      table.font.size = font.size,
      heading.subtitle.font.size = font.size,
      table_body.border.bottom.color = "black",
      table_body.border.bottom.width = 2,
      table_body.border.bottom.style = "solid",
      column_labels.border.bottom.color = "black",
      column_labels.border.bottom.style = "solid",
      column_labels.border.bottom.width = 1,
      latex.use_longtable = FALSE
    ) |> 
      gt::opt_table_font(font = "times")
}

table <- data.frame(
  Questionnaire = c(
    # Interoception
    "Interoceptive Related", "MAIA-2", "BPQ-SF", "BPQ-VSF", "ICQ", "TAS-20", "BVAQ",
    # Mood/Anxiety
    "Mood", "BDI-II", "PHQ-4", "PHQ-9", "PHQ-15", "MFQ", "STAI-T", "GAD-7",
    # Personality
    "Personality", "NEO-FFI (neuroticism subscale)", "Mini IPIP6", "BFI-S",
    # Psychopathology
    "Psychopathology", "PID-5-SFF", "SPQ-BR", "MSI-BPD", "ASQ - Short",
    # Beliefs and Misbeliefs
    "Beliefs and Misbeliefs", "GCBS", "PI-18", "LIE scale"
  ),
  Number_of_Dimensions = c(
    "", 8, 2, 1, 3, 3, 5,
    "", 2, 2, 1, 1, 1, 1, 1,
    "", 1, 6, 5,
    "", 5, 9, 1, 5,
    "", 5, 3, 4
  ),
  Assessment = c(
    "", "Interoceptive awareness", "Body awareness and autonomic reactivity", "Body awareness", "Interoceptive confusion", "Alexithymia", "Alexithymia", 
    "", "Depressive symptoms severity", "Depression and Anxiety", "Depressive symptoms", "Somatic symptom distress", "Mood and affect", "Trait anxiety", "General anxiety disorder symptoms",
    "", "Neuroticism", "Personality traits", "Personality traits",
    "", "Maladaptive personality traits", "Schizotypy", "Borderline personality disorder", "Autistic traits",
    "", "Conspiracy beliefs", "Primal world beliefs", "Lying tendencies"
  ),
  Number_of_Items = c(
    "", 37, 46, 12, 20, 20, 40,
    "", 21, 4, 9, 15, 13, 20, 7,
    "", 12, 24, 10,
    "", 25, 32, 9, 28,
    "", 15, 18, 16
  ),
  Scoring = c(
    "", "6-point Likert scale", "5-point Likert scale", "5-point Likert scale", "5-point Likert scale", "5-point Likert scale", "5-point Likert scale",
    "", "4-point scale", "4-point Likert scale", "4-point Likert scale", "3-point scale", "3-point scale", "4-point Likert scale", "4-point Likert scale",
    "", "5-point Likert scale", "Visual analogue scales", "7-point Likert scale",
    "", "4-point Likert scale", "5-point Likert scale", "Dichotomous (yes/no)", "4-point Likert scale",
    "", "5-point Likert scale", "5-point Likert scale", "Visual analogue scales"
  ),
  Example_Items = c(
    # Interoception
    "", "I notice when I am uncomfortable in my body", "My heart often beats irregularily", "My mouth being dry", " I cannot tell when my muscles are sore or tight", "I have feelings I can't quite identify", "I like to tell others how I feel",
    #Mood
    "", "I feel sad much of the time","Little interest or pleasure in doing things", "Feeling tired or having little energy?", "Over the last week, how often have you been bothered by back pain?", "I felt miserable or unhappy", "I worry too much about something that really doesn't matter", "Not being able to stop or control worrying",
    # Personality
    "", "I often feel inferior to others", "Am the life of the party", "I see myself as someone who ...is considerate and kind to almost everyone", 
   # Psychopathology
    "", "Plenty of people are out to get me", "I am an odd, unusual person", "Have any of your closest relationships been troubled by a lot of arguments or repeated breakups?", "I find it difficult to work out people’s intentions",
   # Beliefs and Misbeliefs
    "", "Secret organizations communicate with extraterrestrials, but keep this fact from the public", "Nearly everything in the world is beautiful", "I am a good liar"
),
   stringsAsFactors = FALSE
)

table_apa <- table |> 
  gt() |>
  cols_align(align = c("right"), columns = "Scoring") |> 
  cols_label(Number_of_Dimensions = "Number of Dimensions", Number_of_Items = "Number of Items", Example_Items = "Example Items")  |>
  # tab_footnote("* Information taken from the sample description of relevant paper rather than recomputed.") 
  tab_style( style = list(cell_text(weight = "bold")), locations = cells_body(columns = Questionnaire, rows = Questionnaire %in% c("Interoceptive Related", "Mood", "Personality", "Psychopathology", "Beliefs and Misbeliefs")))

gt_apastyle(table_apa, font.size=12)
gtsave(gt_apastyle(table_apa, font.size=9), "figures/table2.tex")
saveRDS(table, "figures/table2.rds")
```

## Data Preparation

```{r}
#| warning: false

load("../data/data.RData")
load("../data/scores.RData")

df1a <- cbind(data$df1a, scores$sample1a)
df1b <- cbind(data$df1b, scores$sample1b)
df2 <- cbind(data$df2, scores$sample2)
df3 <- cbind(data$df3, scores$sample3)
df4 <- cbind(data$df4, scores$sample4)
df5 <- cbind(data$df5, scores$sample5)
df6 <- cbind(data$df6, scores$sample6)
df7a <- cbind(data$df7a, scores$sample7a)
df7b <- cbind(data$df7b, scores$sample7b)
df7c <- cbind(data$df7c, scores$sample7c)
df8a <- cbind(data$df8a, scores$sample8a)
df8b <- cbind(data$df8b, scores$sample8b)
df9 <- cbind(data$df9, scores$sample9)
df10 <- cbind(data$df10, scores$sample10)
df11 <- cbind(data$df11, scores$sample11)
df12 <- cbind(data$df12, scores$sample12)
df13 <- cbind(data$df13, scores$sample13)
df14 <- cbind(data$df14, scores$sample14)
df15 <- cbind(data$df15, scores$sample15)
df16 <- cbind(data$df16, scores$sample16)

vars_intero <- names(select(scores$sample1a, -Sample))

# names(df1a)
# names(df1b)
# names(df2)  #TAS(20 items), MAIA(37 items), ICQ (20 items), BPQ(46 items), BVAQ (40 items)
# names(df3)  # TAS (20), BPQ (46)
# names(df4)  # BPQ(26), PBCS (private bodily consciousness?), MAIA (37), BAQ (18), BRS (4)
# names(df5)  
# names(df6)  # PHQ2 (2 items), GAD2 (2 items), SSS8, IATS
# names(df7a)  # STAIT (20), MAIA(37), CES (not reported in the paper), FFMQ (39), NEOFFI (12)
# names(df7b)  # BPQ (12), PHQ15 (15), PHQ9 (9), MAIA (37), STAIT(20), 
# names(df7c)  # CI01 (not reported in the paper), TAS, BPQ, MAIA, ASI, ICQ  - no individual items
# names(df8a)  # EDI (10), BPQ (12)
# names(df8b)  # PHQ9 (9), PHQ15 (15), TAS (20)
# names(df9)  # ECR (Experience in Close Relationship Scale), TEAQ (touch experiences and attitudes questionnaire), STQ, TACSEN, UCLA, BFI (15), DBIA, BO, EQ
# names(df10)  # ASQ(28), BPD(9), GCBS (15), IPIP6 (24), LIE (16), MAIA(37), PHQ4 (4), PI (32), PID (25), SPQ (32)
# names(df11)  # PHQ4 (4), STAIT5 (5), BDI (21), MAIA (37)
# names(df12) # MAIA (37), PI (99), 
# names(df13) # PI (18), MAIA (37)
# names(df14) # MAIA(37), TAS  (20), MFQ(DEP, 13), STAI(ANX, 20)
# names(df15) 
# names(df16) # MAIA (37), GAD (7), PHQ9 (9)

alldata <- list(df1a=df1a, df1b=df1b, df2=df2, df3=df3, df4=df4, 
                df5=df5, df6=df6, df7a=df7a, df7b=df7b, df7c=df7c, 
                df8a=df8a, df8b=df8b, df9=df9, df10=df10, df11=df11, 
                df12=df12, df14=df14, df15=df15, df16=df16)
```

## Demographics

### Gender

```{r}
#| warning: false

pool_coef <- function(with = "Age", component = "Demographic", category = "Demographic", name = with, unequalrange=FALSE) {
  rez <- data.frame()
  for(intero in vars_intero) {
    data <- data.frame() 
    for(dataset in alldata) {
      if (with %in% names(dataset)) {
        d <- dataset[c(with, intero, "Sample")]
        data <- rbind(data, d)
      }
    }

    # Make sure that min-max of "with" is the equal across all samples
    check <- summarize(data, 
                       min = min(!!sym(with), na.rm = TRUE),  
                       max = max(!!sym(with), na.rm = TRUE), .by = "Sample")
    if (with != "Gender" && (any(check$min != min(check$min)) || any(check$max != max(check$max)))) {
      if(unequalrange == FALSE) {
        print(check)
        stop(paste0("The range of '", with, "' is not equal across samples."))
      }
    }

    data <- standardize(data[complete.cases(data), ])

    if (with == "Gender") {
      data <- data[data$Gender != "Other",]
      data$Gender <- factor(data$Gender, levels = c("Male", "Female"))
      f <- as.formula(paste0(intero, "~", with, " + (", with, "|Sample)"))
      f2 <- as.formula(paste0(intero, "~", with, " + (1|Sample)"))
      f3 <- as.formula(paste0(intero, "~ Sample * ", with))
    } else {
      f <- as.formula(paste0(intero, "~", with, " + (", with, "|Sample)"))
      f2 <- as.formula(paste0(intero, "~", with, " + (1|Sample)"))
      f3 <- as.formula(paste0(intero, "~ Sample / ", with))
    }

    # Mixed models
    control1 <- glmmTMB::glmmTMBControl(parallel = list(n=4, autopar = TRUE))
    control2 <- glmmTMB::glmmTMBControl(optimizer=optim, optArgs=list(method="BFGS"), parallel = list(n=4, autopar = TRUE))
    m <- glmmTMB::glmmTMB(f, data = data, control=control1)
    type <- "Full"
    if(!insight::is_converged(m)) {
      m <- glmmTMB::glmmTMB(f, data = data, control=control2)
      type <- "Full"
    }
    if(!insight::is_converged(m)) {
      m <- glmmTMB::glmmTMB(f2, data = data, control=control1)
      type <- "Reduced"
    }
    if(!insight::is_converged(m)) {
      m <- glmmTMB::glmmTMB(f2, data = data, control=control2)
      type <- "Reduced"
    }

    # # Marginal
    # m <- lm(f3, data = data)
    # modelbased::estimate_slopes(m, trend = with)

    out <- parameters::parameters(m)[2, ] |> 
      select(r = Coefficient, CI_low, CI_high, p) |> 
      as.data.frame()
    out$var_Interoception <- intero
    out$var_Correlate <- with
    out$Category <- category
    out$Component <- component
    out$Type <- type
    out$Converged <- insight::is_converged(m)
    out$N <- nrow(data)
    rez <- rbind(rez, out)

  }
  rez$var_Correlate <- name
  rez
}

check_ranges <- function(var = "Neuroticism") {
  for(d in names(alldata)) {
    dataset <- alldata[[d]]
    if(!var %in% names(dataset)) next
    print(paste0(d, ": ", paste(round(range(dataset[[var]], na.rm=TRUE), 2), collapse = "-")))
  }
}

data <- pool_coef(with = "Gender", component = "Demographic", category = "General", name = "Gender (Female - Male)")
```

### Age

```{r}
#| warning: false

# check_ranges("Age")
data <- rbind(data, pool_coef(with = "Age", component = "Demographic", category = "General", unequalrange=TRUE))
```

## Personality

```{r}
#| warning: false

# names(df7a)[str_detect(names(df7a), "NEO")]
# names(df9)[str_detect(names(df9), "BFI_")]
# names(df10)[str_detect(names(df10), "IPIP6_")]

# check_ranges("Neuroticism")
# check_ranges("Agreeableness")
# check_ranges("Conscientiousness")
# check_ranges("HonestyHumility")

data <- rbind(
  data,
  pool_coef(with = "Neuroticism", component = "Personality", category = "General", name = "Neuroticism", unequalrange=TRUE),
  pool_coef(with = "Agreeableness", component="Personality", category = "General", name = "Agreeableness",unequalrange=TRUE),
  pool_coef(with = "Conscientiousness", component="Personality", category = "General", name = "Conscientiousness",unequalrange=TRUE), 
  pool_coef(with = "Openness", component="Personality", category = "General", name = "Openness",unequalrange=TRUE),
  pool_coef(with = "Extraversion", component="Personality", category = "General", name = "Extraversion", unequalrange=TRUE),
  pool_coef(with = "HonestyHumility", component="Personality", category = "General", name = "Honesty-Humility",unequalrange=TRUE)
)
```

## Body

### MAIA

```{r}
#| warning: false

# names(df2)[str_detect(names(df2), "MAIA")]  
# names(df4)[str_detect(names(df4), "MAIA")]  # individual items
# names(df7a)[str_detect(names(df7a), "MAIA")]  
# names(df7b)[str_detect(names(df7b), "maia")] 
# names(df7c)[str_detect(names(df7c), "MAIA")]
# names(df10)[str_detect(names(df10), "MAIA")]
# names(df11)[str_detect(names(df11), "MAIA2")] # all columns with NA values 
# names(df12)[str_detect(names(df12), "MAIA")]
# names(df14)[str_detect(names(df15), "MAIA")]

data <- rbind(
  data,
  pool_coef(with = "MAIA_AttentionReg", component = "MAIA", category = "Interoception", name = "Attention Regulation - MAIA", unequalrange=TRUE),
  pool_coef(with = "MAIA_BodyListening", component = "MAIA", category = "Interoception", name = "Listening - MAIA", unequalrange=TRUE),
  pool_coef(with = "MAIA_Trusting", component = "MAIA", category = "Interoception", name = "Trusting - MAIA", unequalrange=TRUE),
  pool_coef(with = "MAIA_EmoAwareness", component = "MAIA", category = "Interoception", name = "Emotional Awareness - MAIA", unequalrange=TRUE),
  pool_coef(with = "MAIA_NotDistracting", component = "MAIA", category = "Interoception", name = "Not Distracting - MAIA", unequalrange=TRUE),
  pool_coef(with = "MAIA_Noticing", component = "MAIA", category = "Interoception", name = "Noticing - MAIA", unequalrange=TRUE),
  pool_coef(with = "MAIA_NotWorrying", component = "MAIA", category = "Interoception", name = "Not Worrying - MAIA", unequalrange=TRUE),
  pool_coef(with = "MAIA_SelfReg", component = "MAIA", category = "Interoception", name = "Self Regulation - MAIA", unequalrange=TRUE)
)
```

### BPQ

```{r}
#| warning: false

# df3$BPQ_BodyAwareness <- -1*normalize(df3$BPQ_BodyAwareness)
# df3$BPQ_AutonomicReactivity <- -1*normalize(df3$BPQ_AutonomicReactivity)
# names(df2)[str_detect(names(df2), "BPQ")]  # BPQ_R_Supra and _Sub?
# names(df3)[str_detect(names(df3), "BPQ")]  
# names(df4)[str_detect(names(df4), "BPQ")]   # individual items 1-26
# names(df7b)[str_detect(names(df7b), "BPQ")]  # BPQ sum
# names(df8a)[str_detect(names(df8a), "BPQ")]  # individual items

data <- rbind(
  data,
  pool_coef(with = "BPQ_BodyAwareness", component="Interoception", category = "Interoception", name = "Body Awareness - BPQ",unequalrange=TRUE),
  pool_coef(with = "BPQ_AutonomicReactivity", component="Interoception", category = "Interoception", name = "Autonomic Reactivity - BPQ",unequalrange=TRUE)
  )
```

### ICQ

```{r}
#| warning: false

# names(df2)[str_detect(names(df2), "ICQ")] 
# names(df7c)[str_detect(names(df7c), "ICQ")]

data <- rbind(data, pool_coef(with = "ICQ", component="Interoception", category = "Interoception", name = "Interoceptive Confusion - ICQ", unequalrange=TRUE))
```

### Alexithymia

```{r}
#| warning: false

#  only one study assess alexithymia with BVAQ (df2)
# names(df2)[str_detect(names(df2), "TAS", "BVAQ)]
# names(df3)[str_starts(names(df3), "TAS")]
# names(df7c)[str_starts(names(df7c), "TAS")]
# names(df8b)[str_starts(names(df8b), "TAS")]
# names(df14)[str_starts(names(df14), "TAS")]

data <- rbind(
  data,
  pool_coef(with = "TAS_DIF", component="Interoception", category = "Interoception", name = "Difficulties Identifying Feelings - TAS", unequalrange=TRUE),
  pool_coef(with = "TAS_DDF", component="Interoception", category = "Interoception", name = "Difficulties Describing Feelings - TAS", unequalrange=TRUE),
  pool_coef(with = "TAS_EOT", component="Interoception", category = "Interoception", name = "External Thinking - TAS",unequalrange=TRUE),
  pool_coef(with = "BVAQ_Affective", component="Interoception", category = "Interoception", name = "Affective - BVAQ"), 
  pool_coef(with = "BVAQ_Cognitive", component="Interoception", category = "Interoception", name = "Cognitive - BVAQ")
)
```

## Mood

### Depression

```{r}
#| warning: false

# names(df6)[str_detect(names(df6), "PHQ")] 
# names(df7b)[str_detect(names(df7b), "PHQ")]
# names(df7c)[str_detect(names(df7c), "BDI")]
# names(df8b)[str_detect(names(df8b), "PHQ")]
# names(df9)
# names(df14)

check_ranges("PHQ15_Sum")
check_ranges("PHQ9_Sum")
check_ranges("BDI_Sum")
check_ranges("PHQ4_Depression")
check_ranges("DEP")

data <- rbind(
  data,
  pool_coef(with = "PHQ15_Sum", component="Depression", category = "Mood", name = "Somatic Concerns - PHQ-15",unequalrange=TRUE), 
  pool_coef(with = "PHQ9_Sum", component="Depression", category = "Mood", name = "Depression - PHQ-9", unequalrange=TRUE), 
  pool_coef(with = "BDI_Sum", component="Depression", category = "Mood", name = "Depression - BDI"), 
  pool_coef(with = "PHQ4_Depression", component="Depression", category = "Mood", name = "Depression - PHQ-4"),
  pool_coef(with = "DEP", component = "Depression", category = "Mood", name = "Depression - MFQ")
)
```

### Anxiety

```{r}
#| warning: false

# names(df6)[str_detect(names(df6), "GAD")] 
# names(df7a)[str_detect(names(df7a), "STAI")] 
# names(df7b)[str_detect(names(df7b), "STAI")] 
# names(df7c)[str_detect(names(df7c), "STAI")]
# names(df10)[]
# names(df11)[str_detect(names(df11), "STAI")] 
# names(df14)[str_detect(names(df14), "ANX")] # anx = stai-t

# check_ranges("STAIT_Mean")
# check_ranges("PHQ4_Anxiety")
# check_ranges("GAD7_Sum")

data <- rbind(
  data,
  pool_coef(with = "STAIT_Mean", component="Anxiety", category = "Mood", name = "Anxiety - STAI-T", unequalrange=TRUE), 
  pool_coef(with = "PHQ4_Anxiety", component="Anxiety", category = "Mood", name = "Anxiety - PHQ-4"),
  pool_coef(with = "GAD7_Sum", component="Anxiety", category = "Mood", name = "Anxiety - GAD-7")
)
```

## Psychopathology

```{r}
#| warning: false

# check_ranges("BPD")

data <- rbind(
  data,
  pool_coef(with = "BPD", component="Maladaptive", category = "Psychopathology", name = "Boderline Personality")
)
```

```{r}
#| warning: false

# check_ranges("PID5_Antagonism")

data <- rbind(
  data,
  pool_coef(with = "PID5_Antagonism", component="Maladaptive", category = "Psychopathology", name = "Antagonism - PID-5"),
  pool_coef(with = "PID5_Detachment", component="Maladaptive", category = "Psychopathology", name = "Detachment - PID-5"), 
  pool_coef(with = "PID5_Disinhibition", component="Maladaptive", category = "Psychopathology", name = "Disinhibition - PID-5"),
  pool_coef(with = "PID5_NegativeAffect", component="Maladaptive", category = "Psychopathology", name = "Negative Affect - PID-5"),
  pool_coef(with = "PID5_Psychoticism", component="Maladaptive", category = "Psychopathology", name = "Psychoticism - PID-5")
)
```

```{r}
#| warning: false

data <- rbind(
  data,
  pool_coef(with = "SPQ_ConstrictedAffect", component="Schizotypic", category = "Psychopathology", name = "Constricted Affect - SPQ"),
  pool_coef(with = "SPQ_Eccentric", component="Schizotypic", category = "Psychopathology", name = "Eccentric - SPQ"),
  pool_coef(with = "SPQ_MagicalThinking", component="Schizotypic", category = "Psychopathology", name = "Magical Thinking - SPQ"),
  pool_coef(with = "SPQ_NoCloseFriends", component="Schizotypic", category = "Psychopathology", name = "No Close Friends - SPQ"),
  pool_coef(with = "SPQ_OddSpeech", component="Schizotypic", category = "Psychopathology", name = "Odd Speech - SPQ"),
  pool_coef(with = "SPQ_Reference", component="Schizotypic", category = "Psychopathology", name = "Reference - SPQ"),
  pool_coef(with = "SPQ_SocialAnxiety", component="Schizotypic", category = "Psychopathology", name = "Social Anxiety - SPQ"),
  pool_coef(with = "SPQ_Suspiciousness", component="Schizotypic", category = "Psychopathology", name = "Suspiciousness - SPQ"),
  pool_coef(with = "SPQ_UnusualPerceptions", component="Schizotypic", category = "Psychopathology", name = "Unusual Perceptions - SPQ")
)
```

```{r}
#| warning: false

#  names(select(df10, starts_with("ASQ")))

data <- rbind(
  data,
  pool_coef(with = "ASQ_Imagination", component="Autistic", category = "Psychopathology", name = "Imagination - ASQ"),
  pool_coef(with = "ASQ_LackSocialSkills", component="Autistic", category = "Psychopathology", name = "Lack of Social Skills - ASQ"), 
  pool_coef(with = "ASQ_LowAttentionalSwitching", component="Autistic", category = "Psychopathology", name = "Low Attentional Switching - ASQ"),
  pool_coef(with = "ASQ_Patterns", component="Autistic", category = "Psychopathology", name = "Patterns and Numbers - ASQ"),
  pool_coef(with = "ASQ_Routine", component="Autistic", category = "Psychopathology", name = "Routines - ASQ")
)
```

## Beliefs

```{r}
#| warning: false

# names(df10)[str_detect(names(df10), "PI_Alive")]
# names(df12)[str_detect(names(df12), "PI_Alive")]
# names(df13)[str_detect(names(df13), "PI_Alive")]

# data <- rbind(
#   data, 
#   pool_coef(with = "PI_Alive", component="World Beliefs", name = "Alive", unequalrange=TRUE))

data <- rbind(
  data,
  pool_coef(with = "PI_Alive", component="World Beliefs", name = "Alive", unequalrange=TRUE),
  pool_coef(with = "PI_Changing", component="World Beliefs", name = "Changing", unequalrange=TRUE),
  pool_coef(with = "PI_Enticing", component="World Beliefs", name = "Enticing", unequalrange=TRUE),
  pool_coef(with = "PI_Good", component="World Beliefs", name = "Good", unequalrange=TRUE),
  pool_coef(with = "PI_Hierarchical", component="World Beliefs", name = "Hierarchical", unequalrange=TRUE),
  pool_coef(with = "PI_Safe", component="World Beliefs", name = "Safe",unequalrange=TRUE),
  pool_coef(with = "PI_Understandable", component="World Beliefs", name = "Understandable",unequalrange=TRUE),

  pool_coef(with = "PI_AboutMe", component="World Beliefs", name = "About me",unequalrange=TRUE),
  pool_coef(with = "PI_Abundant", component="World Beliefs", name = "Abundant",unequalrange=TRUE),
  pool_coef(with = "PI_Acceptable", component="World Beliefs", name = "Acceptable",unequalrange=TRUE),
  pool_coef(with = "PI_Beautiful", component="World Beliefs", name = "Beautiful",unequalrange=TRUE),
  pool_coef(with = "PI_Cooperative", component="World Beliefs", name = "Cooperative",unequalrange=TRUE),
  pool_coef(with = "PI_Funny", component="World Beliefs", name = "Funny",unequalrange=TRUE),
  pool_coef(with = "PI_Harmless", component="World Beliefs", name = "Harmless",unequalrange=TRUE),
  pool_coef(with = "PI_Improvable", component="World Beliefs", name = "Improvable",unequalrange=TRUE),
  pool_coef(with = "PI_Interconnected", component="World Beliefs", name = "Interconnected",unequalrange=TRUE),
  pool_coef(with = "PI_Interesting", component="World Beliefs", name = "Interesting",unequalrange=TRUE),
  pool_coef(with = "PI_Just", component="World Beliefs", name = "Just",unequalrange=TRUE),
  pool_coef(with = "PI_Meaningful", component="World Beliefs", name = "Meaningful",unequalrange=TRUE),
  pool_coef(with = "PI_NeedsMe", component="World Beliefs", name = "Needs Me",unequalrange=TRUE),
  pool_coef(with = "PI_Pleasurable", component="World Beliefs", name = "Pleasurable",unequalrange=TRUE),
  pool_coef(with = "PI_Progressing", component="World Beliefs", name = "Progressing",unequalrange=TRUE),
  pool_coef(with = "PI_Regenerative", component="World Beliefs", name = "Regenerative",unequalrange=TRUE),
  pool_coef(with = "PI_Stable", component="World Beliefs", name = "Stable",unequalrange=TRUE),
  pool_coef(with = "PI_WorthExploring", component="World Beliefs", name = "Worth Exploring",unequalrange=TRUE),
  pool_coef(with = "PI_Intentional", component="World Beliefs", name = "Intentional", unequalrange=TRUE)
  )
```

```{r}
#| warning: false

data <- rbind(
  data,
  pool_coef(with = "GCBS_Extraterrestrial", component="Conspiracy Beliefs", name = "Extraterrestrial"),
  pool_coef(with = "GCBS_GlobalConspiracies", component="Conspiracy Beliefs", name = "Global Conspiracies"),
  pool_coef(with = "GCBS_GovernmentMalfeasance", component="Conspiracy Beliefs", name  = "Government Malfeasance"),
  pool_coef(with = "GCBS_InformationControl", component="Conspiracy Beliefs", name = "Information Control"),
  pool_coef(with = "GCBS_PersonalWellbeing", component="Conspiracy Beliefs", name = "Personal Wellbeing")
)
```

```{r}
#| warning: false

data <- rbind(
  data,
  pool_coef(with = "LIE_Ability", component="Lying Profile", name = "Ability"),
  pool_coef(with = "LIE_Frequency", component="Lying Profile", name = "Frequency"),
  pool_coef(with = "LIE_Contextuality", component="Lying Profile", name = "Contextuality"),
  pool_coef(with = "LIE_Negativity", component="Lying Profile", name = "Negativity")
)
```

## Average Correlations

-   BPQ - sf - Body perception questionnaire short form 
-   BPQ - vsf - Body perception questionnaire very short form
-   BVAQ - Bermond–Vorst Alexithymia Questionnaire
-   MAIA-2 - Multidimensional Assessment of Interoceptive Awareness , Version-2
-   ICQ - Interocpetive Confusion Questionnaire
-   TAS - Toronto Alexithymia Scale
-   BAQ - Body awareness questionnaire
<!-- -   PBCS - The Private Bodily Consciousness Scale -->
<!-- -   BRS - the body responsiveness scale -->
-   PHQ-2 - Patient Health Questionnaire 2 items on major depressive episodes
-   PHQ-4 - atient Health Questionnaire 4 items assessing anxiety and depression
-   PHQ-9 - Patient Health Questionnaire 9-Item Version assesses depressive symptoms
-   PHQ-15 - Patient Health Questionnaire 15-Item Version assesses Somatic symptom distress
<!-- -   SSS-8 - Somatic Symptom Scale -->
<!-- -   IATS - Interoceptive Attention Scale -->
-   GAD-2 - Generalized Anxiety Disorder Scale 2 items on anxiety disorder
-   STAIT - State-Trait-Anxiety Inventory Trait subscale
-   FFMQ - Observation and Description subscales of the Five Facet Mindfulness Questionnaire
-   NEOFFI - Neuroticism Neo Five-Factor Inventory Neuroticism subscale
<!-- -   ASI - Anxiety Sensitivity Inventory 3 -->
-   BDI - The Beck Depression Inventory
<!-- -   EDI_IA - Eating Disorder Inventory, interoceptive awareness subscale -->
-   BFI - Big Five inventory-10
-   ASQ - Autism-Spectrum Quotient
<!-- -   MSI-BPD - Mclean Screening Instrument for Borderline Personality Disorder -->
-   GCB - Generic Conspiracist Beliefs scale
-   IPIP6 - Mini International Personality Item Pool
-   PI - Primal Inventory 18 items and 99 items
-   PID5-sf - Personality Inventory for DSM-5 Short Form
-   SPQ-BRU - Schizotypal Personality Questionnaire, brief revised updated
-   LIE - Lying Profile Questionnaire

```{r}
# names(df1a)
# names(df1b)
# names(df2)  #TAS, MAIA, ICQ, BPQ, BVAQ
# names(df3)  # TAS, BPQ
# names(df4)  # BPQ, PBCS (private bodily consciousness?), MAIA, BAQ, BRS
# names(df5)  
# names(df6)  # PHQ2, GAD2, SSS8, IATS
# names(df7a)  # STAIT, MAIA, CES (not reported in the paper), FFMQ, NEOFFI
# names(df7b)  # BPQ, PHQ15, PHQ9, MAIA, STAIT, 
# names(df7c)  # CI01 (not reported in the paper), TAS, BPQ, MAIA, ASI, ICQ 
# names(df8a)  # EDI, BPQ
# names(df8b)  # PHQ9, PHQ15, TAS
# names(df9)  # ECR (Experience in Close Relationship Scale), TEAQ (touch experiences and attitudes questionnaire), STQ, TACSEN, UCLA, BFI, DBIA, BO, EQ
# names(df10)  # ASQ, BPD, GCBS, IPIP6, LIE, MAIA, PHQ4, PI, PID, SPQ
# names(df11)  # PHQ4, STAIT5, BDI, MAIA
# names(df12) # MAIA, PI, 
# names(df14) # MAIA, TAS, MFQ(DEP), STAI(ANX)
# names(df15) #
# names(df16) #

average_corr <- data |>
  summarise(min_r = min(r, na.rm = TRUE), 
            mean_r = mean(r, na.rm = TRUE),
            max_r = max(r, na.rm = TRUE), 
            Max = var_Interoception[which.max(abs(r))],
            .by=c("Component", "var_Correlate")) 
#   
# average_corr <- data |>
#   summarise(
#     min_r = min(r, na.rm = TRUE), 
#     mean_r = mean(r, na.rm = TRUE),
#     max_r = max(r, na.rm = TRUE), 
#     Max = var_Interoception[which.max(abs(r))],
#     n_sig_001 = sum(p < .001, na.rm = TRUE),
#     n_sig_01  = sum(p < .01 & p >= .001, na.rm = TRUE),
#     n_ns      = sum(p >= .01 | is.na(p)),
#     total     = dplyr::n(),
#     .by=c("Component", "var_Correlate")
#   ) |>
#   mutate(
#     perc_sig = (n_sig_001 + n_sig_01) / total
#   )

sort(table(average_corr$Max), decreasing = TRUE)

average_corr |> 
  gt::gt() |> 
  gt::data_color(
    columns = c(mean_r, min_r, max_r),
    palette = c("#2196F3", "white", "#FF5722"),
    domain = c(-0.7, 0, 0.7)
  ) |>
  gt::data_color(
    columns = c(Component, Max)
  ) |> 
  gt::fmt_auto()
```

## Summary

```{r}
#| warning: false
#| fig-height: 20
#| fig-width: 15

p <- data |>
  mutate(sig = ifelse(p < .01, "p < .01", ifelse(p < .001, "p < .001", "N.S.")),
         dir = sign(r),
         var_Interoception = str_replace(var_Interoception, "CoughSneeze", "Cough/Sneeze"),
         var_Interoception = str_replace(var_Interoception, "HungryThirsty", "Hungry/Thirsty"),
         var_Interoception = str_replace(var_Interoception, "BruiseBlood", "Bruise/Blood sugar"),
         var_Interoception = str_replace(var_Interoception, "MusclesPain", "Muscles/Pain"),
         var_Interoception = str_replace(var_Interoception, "BreathingHeart", "Breathing/Heart"),
         var_Interoception = str_replace(var_Interoception, "UrinateDefecate", "Urinate/Defecate"),
         var_Interoception = str_replace(var_Interoception, "WindBurp", "Wind/Burp"),
         var_Interoception = fct_relevel(var_Interoception, "Original", "Hungry/Thirsty", "Bruise/Blood sugar", "Muscles/Pain", "Cough/Sneeze", "Wind/Burp", "Breathing/Heart", "Urinate/Defecate"),
         Category = fct_relevel(
           Category, "General", "Interoception", "Mood", "Psychopathology"),
         Component = fct_relevel(Component, "Personality", "Demographic", "Interoception", "MAIA", "Anxiety", "Depression", "Schizotypic", "Autistic", "Maladaptive")) |> 
  group_by(Component) |> 
  mutate(var_Correlate = fct_reorder(var_Correlate, r, .fun=mean)) |>
  ungroup() |>
  ggplot(aes(y = var_Correlate)) +
  annotate(geom = "rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax=Inf, fill = "#FFF8E1") +
  annotate(geom = "rect", xmin = c(-0.3, 0.1), xmax = c(-0.1, 0.3), ymin = -Inf, ymax=Inf, fill = "#F1F8E9") +
  annotate(geom = "rect", xmin = c(-0.6, 0.3), xmax = c(-0.3, 0.6), ymin = -Inf, ymax=Inf, fill = "#E8F5E9") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  # geom_point2(aes(y=r, color = dir, alpha = sig), size = 3) +
  geom_linerange(aes(color=dir, alpha=sig, xmax=r, xmin=0), linewidth = 1) +
  geom_linerange(aes(color = dir, alpha = sig, xmin = CI_low, xmax = CI_high, linewidth = log(N))) +
  # geom_linerange(aes(color=dir, alpha=sig, ymin=0, ymax=r, group=Sample), position = position_dodge2(width = 0.7), linewidth=2.5) +
  # geom_linerange(aes(ymin = CI_low, ymax = CI_high, color = dir, alpha=sig, group=Sample), position = position_dodge2(width = 0.7),
  #                linewidth = 1) +
  facet_grid(Category ~ var_Interoception, scales = "free", space ="free", switch = "y") +
  scale_fill_gradient(low = "#2196F3", high = "#FF5722", guide = "none") +
  scale_color_gradient(low = "#2196F3", high = "#FF5722", guide = "none") +
  scale_alpha_discrete(range = c(0.2, 0.9), guide = "none") +
  scale_x_continuous(expand = c(0, 0), breaks = c(-0.3, 0, 0.3)) +
  scale_linewidth_continuous(range = c(3, 7), guide = "none") +
  coord_cartesian(xlim = c(-0.47, 0.47)) +
  theme_modern(axis.title.space = 5) +
  labs(x = "Strength of Association (std. Beta)") +
  theme(panel.grid.major.y = element_line(),
        axis.title.y = element_blank(),
        axis.line.x = element_blank(),
        axis.text.y = element_text(size = rel(0.5)),
        axis.text.x = element_text(size = rel(0.8)),
        plot.title = element_text(face = "bold", hjust = 0),
        strip.placement = "outside",
        strip.background.y = element_rect(fill = "#E0E0E0", color = "white"),
        strip.background.x = element_rect(fill = "#F8BBD0", color = "white"))
p
ggsave("figures/Figure3.png", p, width=21*1, height=29.7*1, dpi=300, bg="white")
```

## Discussion

Benefits of the IAS: - Straightforward and sensation-centered items

Recommendations: - Remove Itch (redundant + issue in Chinese) - Use analog scales

Limitations: - Not much clear theorethical or empirical structure (small grouping of items) - Limited variability (clear mode at 4/5), very low prevalence of 5/5 responses. - Ambiguous items which grouping depends on the context (and its awareness) - E.g., heart beating fast, vomit when nauseaous - Few items for some modalities (e.g., 1 for heart) - Positive phrasing: benefits but also might exacerbate positivity bias (and thus unidimensionality)

Need for context-specific items (cross-modal when possible, i.e., cardioception, respiroception, etc.).