### 2025-10-22 paper/1_Initial/manuscript.qmd
Update manuscript.qmd

---
title: "Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates"
shorttitle: "IAS Meta-analysis"
author:
  - name: Ana Neves
    orcid: 0009-0006-0020-7599
    roles:
      - Data curation
      - Formal Analysis
      - Investigation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
        author:
  - name: Magdalena Pfaff
    orcid: 0009-0006-2386-7936
    roles:
      - Data curation
      - Writing – original draft
    affiliations:
      - ref: "id1"
  - name: Robyn Scharte
    orcid: blabla
    roles: 
      - bla bla
    affiliations:
      - ref: "id1"
  - name: Raquel Nogueira Arjona
    orcid: 0000-0002-5998-701X
    roles: 
      - bla bla
    affiliations:
      - ref: "id1"
      - name: "University of Sussex"
        department: "Sussex Addictions Research and Intervention Centre"
  - name: Giulia Poerio
    orcid: 0000-0002-2343-5109
    roles: 
      - bla bla
    affiliations:
      affiliations:
      - ref: "id1"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Project administration
      - Data curation
      - Formal Analysis
      - Investigation
      - Visualization
      - Writing – original draft
      - Writing – review & editing
    affiliations:
      - ref: "id1"
      - name: "University of Sussex"
        department:  "Sussex Centre for Consciousness Science"
# author-note:
#   status-changes: 
#     affiliation-change: null
#     deceased: null
#   disclosures:
#     study-registration: null
#     data-sharing: null
#     related-report: null
#     conflict-of-interest: null
#     financial-support: null
#     gratitude: null
#     authorship-agreements: null
    #   # - Project administration
    #   # - Data curation
    #   # - Formal Analysis
    #   # - Investigation
    #   # - Visualization
    #   # - Writing – original draft
    #   # - Writing – review & editing
    # Roles are optional. 
    # conceptualization, data curation, formal Analysis, funding acquisition, investigation, 
    # methodology, project administration, resources, software, supervision, validation, 
    # visualization, writing, editing
    #   - Conceptualization
    #   - Data curation
    #   - formal Analysis
    #   - Funding acquisition
    #   - Investigation
    #   - Methodology
    #   - Project administration
    #   - Resources
    #   - Software
    #   - Supervision
    #   - Validation
    #   - Visualization
    #   - Writing – original draft
author-note:
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      :::
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
abstract: |
  Blabla the abstract blabla.
keywords: [keyword1, keyword2, keyword3]
floatsintext: true
numbered-lines: false
bibliography: bibliography.bib
suppress-title-page: false
mask: false
# Language options. See https://quarto.org/docs/authoring/language.html
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  title-block-author-note: "Author Note"
  title-block-correspondence-note: "Correspondence concerning this article should be addressed to"
  title-block-role-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
  references-meta-analysis:
format:
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: jou
    include-in-header:
      - text: |
          \usepackage{lscape}
  apaquarto-docx: default
  apaquarto-html: default
editor: 
  markdown: 
    wrap: sentence
editor_options: 
  chunk_output_type: console
---

# Introduction

<!-- Interoception definition -->

Interoception refers to the process of sensing, interpreting and integrating information pertaining to internal organs, such as the heart, the lungs or the gut [@khalsa2018interoception].
While recent research emphasizes a key role of interoception in a variety of processes (e.g., emotion regulation, decision making) and of outcomes (physical and psychological well being), the field remains clouded by concerns about how interoception is assessed.

## The Interoceptive Assessment Puzzle

<!-- Interoception measure -->

Various measures of interoception have been developed (see @fig-measures), forming a combination of "objective" and "subjective" assessments (i.e., physiological tasks such as the heart beat counting or tracking *vs.* questionnaires and subjective scales involving metacognitive judgments), "explicit" and "implicit" paradigms (i.e., directing participants' awareness and attention to interoceptive processes *vs.* measuring interoception unbeknownst to them), various interoceptive modalities (e.g., cardioception, respiroception, gastroception) and theoretical dimensions (e.g., accuracy, sensitivity, awareness).
While there is no consensus as to which particular approach provides the most accurate and "pure" measure of interoception and interoceptive abilities (assuming it is a unidimensional construct), it is instead plausible that each measure has strengths and limitations, and a utility dependent on the context and goal at hand [@jahedi2014; @Desmedt2023accuracy].

```{r}
#| warning: false
#| label: "fig-measures"
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"
#| fig-cap: "The Interoceptive Assessment Puzzle. The different modalities of interoception (e.g., cardioception) can be assessed directly or indirectly. Direct assessments can further be subjective or objective, depending on whether they involve conscious metacognitive appraisals or more performance-based indices. Interoceptive tasks can be explicit (the participant is aware of the interoceptive nature of the task and must consciously attend to interoceptive signals; e.g., the heartbeat counting task) or implicit (measurements of interoception done unbeknowst to the participant; e.g., heartbeat evoked potentials measured during resting state). Indirect assessments evaluate constructs typically related (and ideally dependent on) to interoceptive processes or ability (or its deficit)."

knitr::include_graphics("figures/MeasuringInteroception.png")
```

<!-- self-reports -->

Although the use of subjective self-report questionnaires to measure deeply embodied functions might seem paradoxical at first, recent redefinitions of interoception emphasize the role of high-level and metacognitive elaboration of interoceptive information.
These redefinitions provide theoretical grounding to support the idea that some facets of interoception, including participants' metacognitive beliefs, can be assessed subjectively [@khalsa2018interoception; @suksasilp2022towards].
Moreover, the notion that self-reports might not reflect the same processes as other interoception tasks might be important to contextualize the apparent lack of convergence between measures in the field [@desmedt2022measures]. 
For instance, existing findings typically show weak or no correlations between questionnaires and objective measures, such as the Heartbeat Counting Task [HCT, @schandry1981heart] and the Heartbeat Detection Task [HDT, @kleckner2015methodological], including for measures of the same theoretical dimensions [e.g., task-based accuracy vs. self-reported accuracy, @murphy2019; @brand2023; @arslanova2022].
Additionally, even various objective measures assessing in theory the same interoceptive dimension, such as accuracy, either show no or weak correlation [respectively, @brand2023; @hickman2020relationship].
Perhaps more surprisingly, low correlations have been observed even among questionnaires, suggesting (in parallel to major validity concerns) the potential targeting of different facets related to interoception.

One striking example concerns the assessment of interoceptive sensibility, which is broadly defined as the self-reported tendency to focus on and detect internal sensations [@garfinkel2015knowing], but more narrowly as the subjective tendency to focus on interoceptive signals, without necessarily implying detection ability [@khalsa2018interoception].
A recent systematic review suggested that various questionnaires designed to assess interoceptive sensibility may, in fact, measure distinct constructs, with the risk of researchers treating them as equivalent despite overall low convergence [@desmedt2022measures].
Notably, this review adopted a broad definition of sensibility, incorporating both interoceptive sensibility and interoceptive self-report scales, following the eight-facet model by @khalsa2018interoception. 
<!-- In this framework, self-report measures are categorized into interoceptive sensibility and interoceptive self-report scales (i.e., the ability to evaluate one's personal experiences of internal bodily states, assess their consequences, and articulate them through speech or movement). -->
Several widely used questionnaires were included in the review, such as the Multidimensional Assessment of Interoceptive Awareness [MAIA, @mehling2012; MAIA-2, @mehling2018multidimensional], the Body Perception Questionnaire [BPQ, @porges1993body], the Private subscale of the Body Consciousness Questionnaire [PBCS, @miller1981consciousness], the Body Awareness Questionnaire [BAQ, @shields1989body], and the Eating Disorder Inventory [EDI, @garner1991eating; @garner1983].
The lack of correlations to moderate correlations among these questionnaires highlight the need for greater conceptual clarity regarding what each measure captures, how they relate to different dimensions of interoception, and their potential overlaps with other constructs, such as alexithymia and body awareness.

## The Interoceptive Accuracy Scale (IAS)

<!-- Introducing the IAS -->

Focusing on the *accuracy* dimension of interoception, a recently developed scale with a rapidly growing popularity is the Interoceptive Accuracy Scale [IAS, @murphy2019].
The IAS consists of 21 Likert-scale items that query how accurately one can perceive different bodily signals, with one item per physiological modality such as respiration (*"I can always accurately perceive when I am breathing fast"*), heart (*"I can always accurately perceive when my heart is beating fast"*), skin (*"I can always accurately perceive when something is going to be ticklish"*), arousal or bodily functions like coughing (*"I can always accurately perceive when I am going to cough"*) or urinating (*"I can always accurately perceive when I need to urinate"*).
Appealingly, the IAS' statements are about specific interoceptive behaviours and "objective" experiences (although still subjective reports of them), which is a distinct difference with other popular interoception questionnaires, such as the MAIA-2, which contains more general and metacognitive items (e.g., *"I trust my body sensations"*, *"I can notice an unpleasant body sensation without worrying about it"*), as well as dimensions related to attention regulation (e.g., Not-distracting) or emotion regulation (e.g., Not-worrying).

<!-- IAS structure  --> 
The original validation study suggested a two-factor structure for the IAS, one reflecting the perception of general interoceptive signals (urinate, hungry, defecate, thirsty, pain, heart, taste, breathing, temperature, muscles, affective touch, vomit, sexual arousal), and other relating to signals that may be difficult to perceive solely through interoceptive information (itch, tickle, cough, burp, bruise, blood sugar, sneeze, wind). 
The authors however underlined its acceptable but imperfect fit [@murphy2019, p. 127], and several follow-up studies have indeed identified different optimal solutions. 
For instance, @brand2023 reported a 1-factor solution, while @lin2023 and @campos2021 found bifactor solutions [i.e., one general factor above a set of lower-level factors, @rodriguez2016evaluating] to be the best fit. 
Using a 2-factors Exploratory Factor Analysis (EFA), @koike2023 suggested that the items could be grouped into cutaneous (itching, tickling, coughing, burping, affective touch, bruising, passing gas, sneezing, muscle sensations, sexual arousal, and taste) and visceral sensations (urination, defecation, hunger, thirst, pain, breathing, fatigue/blood sugar, temperature, vomiting, and heartbeat).

Discussions have also been focused on specific items.
For instance, @murphy2019 notes that some items might measure direct interoceptive signals such as cardioception, while others might capture phenomena not perceivable through interoceptive signals alone (e.g., "bruising"; p. 119). 
@lin2023 additionally highlights five locally dependent pairs and three items (touch, blood sugar, bruise) with exceptionally high difficulty and low discrimination, and @campos2021 reported "tickle" to be the only item that reflected more specific factors than the general factor. Furthermore, localization issues also arose, with both "itch" and "tickle" corresponding to the same Chinese character, leading to their collapse into a single item [@lin2023].

<!-- Relationship with other interoception measures -->
Regarding its validity, the IAS has naturally been compared to other interoception-related measures, and shows a positive correlations with most facets of the MAIA [@mehling2018multidimensional; @gaggero2021], except for the Not-Distracting and Not-Worrying subscales [@brand2023] - which were previously highlighted as related to non-interoceptive abilities [@ferentzi2021examining].
Interestingly, findings on the correlation between the IAS and the body awareness dimension of the BPQ (i.e., BPQ-A) have been mixed: some studies report small positive correlations [@brand2023; @campos2021; @koike2023], while others find small negative correlations [@lin2023] or no correlation at all [@murphy2019].
Positive correlations have also been observed with the "observation" and "description" subscales of the Five Facet Mindfulness Questionnaire [FFMQ, @baer2006using; @brand2023; @koike2023], as well as with the "non-reactivity" and "acting with awareness" subscales [@koike2023].
Additionally, the IAS has shown a positive correlation with the interoceptive awareness subscale of the Eating Disorder Inventory [EDI-IA, @lin2023] and a negative correlation with the Interoceptive Confusion Questionnaire [ICQ, @brewer2016alexithymia], as reported by @brand2023 and @murphy2019.
Lastly, the correlation with the Interoceptive Attention Scale [IATS, @gabriele2022dissociations] appears rather small [@lin2023; @koike2023].

<!-- Convergent validity (correlates)  -->

While assessing the predictive validity of an interoception scale can be conceived as theoretically challenging, expected negative associations were observed between the IAS and alexithymia [@brand2023; @koike2023; @campos2021; @murphy2019; @lin2023], somatic symptoms [@brand2023; @koike2023; @lin2023], depressive symptoms [@brand2023; @koike2023; @lin2023], anxiety [@brand2023], neuroticism [@brand2023] and self-esteem [@murphy2019].
Taken together, these findings support the IAS as measuring an adaptive aspect of interoception, although its pattern of associations with other interoception (or interoception-related) questionnaires points towards some overlap across various theoretical dimensions, casting some doubt on the orthogonal models of interoception and the possibility of its faithful capture by questionnaires.

<!-- current study -->

Given the increasing relevance of interoception across psychology and neuroscience, and the growing use of the IAS as an interesting self-report measure of interoceptive accuracy, the current study aims at 1) clarifying the structure of the IAS with a mega-analytic (which involves a re-analysis at the raw data level by aggregating datasets) approach that leverages existing data and contrast the traditional CFA/SEM factor-based analyses with network-based ones (Exploratory Graph Analysis);
1) provide an overview of the dispositional correlates of the IAS, clarifying its general pattern of associations, which is key to better understand the nature, place and role of interoception questionnaires within a larger context.

# Study 1

Study 1 will re-analyse and assess the factor structure of the IAS by taking advantage of the large number of open-access datasets [@brand2023; @campos2021; @lin2023; @murphy2019; @arslanova2022; @todd2022; @gaggero2021; @brand2022; @von2023; @poerio2024interoceptive].
While combining these studies might provide a more robust and generalizable understanding of the IAS' factor structure, we also additionally provide an individual analysis (i.e., applying the same method on all samples separately) to add nuance to the general picture, as all studies differ in their sample size, demographic characteristics, language, and procedure.

## Methods

### Datasets

Our search focused on studies citing the original IAS validation paper [@murphy2019], identifying 136 papers (as of 01/05/2024).
To qualify for inclusion, papers needed to (1) provide accessible data in open-access, (2) employ the IAS as a measure, and (3) report individual IAS items scores.
We also included the data of five unpublished studies.
A total of 16 datasets was included (see **Table 1**).

The total number of participants was 33,272 participants (Mean = 48.09 $\pm$ 13.1, 71.3\% Female).

<!-- Note: this table is generated during data analysis -->

<!-- \begin{landscape} -->

<!-- \input{../../analysis/figures/table1.tex} -->

<!-- \end{landscape} -->

```{r}
#| label: tbl-table1
#| tbl-cap: "Description of the samples used in the study."
#| apa-note: "*Information taken from the sample description of relevant paper rather than recomputed."
#| apa-twocolumn: true

library(flextable)
set_flextable_defaults(font.size = 6)

table <- readRDS("../../analysis/figures/table1.rds")

table |>
    dplyr::rename("Age (Mean  ± SD)" = Age, "Female %" = Female_Percentage) |>
    flextable() |>
    set_table_properties(
        layout = "autofit",
        opts_pdf = list(arraystretch = 1)
    ) |>
    theme_apa()
```

### Data Analysis

Psychometrically good items should exhibit various qualities, such as validity, reliability and discrimination, to which one of the contributing factors is the amount of inter-individual variability captured by an item. Items to which all participants' answers are concentrated around one option - i.e., exhibiting a narrow distribution - will be flagged as potentially problematic.

<!-- UVA -->
After examining the distributions and correlations of all IAS items, we will test for "redundant" items (e.g., due to multicollinearity or local dependency) using Unique Variable Analysis [UVA, @christensen2023unique], a novel method derived from network psychometric designed to identify and merge items that share substantial variance (which can distort the structure estimation).
We will use a threshold of 0.30 that detects large to very large overlap to remain conservative and only suggest scale modifications if strongly justified.

Following the analysis of items, we will analyze the factor structure of the IAS using three different approaches, each with particular trade-offs and assumptions, to provide a multi-verse picture of likely solutions. Namely, we will apply traditional exploratory and confirmatory Factor Analysis (EFA/CFA), hierarchical clustering (HCA), and Exploratory Graph Analysis (EGA), to the whole sample, as well as to each dataset separately (details being available in the analysis document at https://github.com/RealityBending/InteroceptionIAS), and our decisions and conclusions will try to take into account both levels of the analysis.

<!-- EGA -->
By combining network analysis with psychometric methods, the recently-developed Exploratory Graph Analysis [EGA, @golino2017exploratory] framework allows to jointly estimate the number of dimensions (i.e., groups of items), the structure as well as its stability [@golino2017exploratory; @golino2020investigating]. 
Evidence has underlined its suitability as an alternative to traditional factor analysis, addressing some of its limitations such as the assumption of a "latent" source of variability, possible biasing in the estimation of the optimal factor numbers depending on sample size, and the poor performance of other methods in complex population structures, while remaining comparable and interpretable [@christensen2021equivalency; @jimenez2023dimensionality]. 
At a fundamental level, EGA conceptualizes variables as nodes in a network, with connections (edges) reflecting associations between them. 
Clustering these nodes reveals distinct communities of related items, in practice akin to traditional latent factors - but without explicitly assuming their presence [@christensen2021equivalency]. 
We used the EGAnet package [@christensen2021estimating] to fit a hierachical EGA with the Leiden community detection algorithm.

<!-- Factor Analysis -->
While EGA offers a robust alternative to traditional factor analysis, factor analysis remains a widely used method for dimensionality assessment. As our goal is to provide a general - yet nuanced - picture, with room to show potential discrepancies emerging from the methods used, we will also include it in the present study.
Unlike EGA, factor analysis assumes that  a latent source of variability — a common latent variable — underlies the observed set of manifest variables [@cosemans2022exploratory].
A critical step in factor analysis is determining the optimal number of factors, for which we will used the Method Agreement Procedure [@ludecke2021performance], which involves a consensus-based decision based on multiple factor estimation methods applied concurrently.

<!-- HCA -->
Finally, we will also apply Hierarchical Clustering Analysis [HCA, @murtagh2014ward], which differs from factor analysis in that it does not assume any latent source of variability, but instead iteratively groups items based on their similarity (e.g., correlation) into a hierarchy of clusters. The benefits of HCA include its interpretability and ability to capture complex relationships among items without relying on strict assumptions about data distribution or latent variables.

<!-- CFA -->
In a typical 2-step fashion, this first analysis run will be followed by a structure refinement with a further selection of items, and the final pool of items will be tested again. Additionally, various solutions (e.g., adding general factors) will be statistically compared using Confirmatory Factor Analysis (CFA).

## Results

<!-- Distributions of items -->

The distribution of the items across samples suggests the presence of a consistent modal value (@fig-distributions).
In other words, participants are most likely to answer 4/5 (i.e., agree) on most items, with the exception of "blood sugar" and "bruise", which exhibit a different distributional pattern with a lower mode (\~2/5). While it is not a problem *per se*, the contrasting distribution might be indicative of items with non-homogeneous psychometric "difficulty".
This is also the case for "affective touch" in samples 8a and 8b (the Chinese validation samples), which might indicate localisation issues.
Additionally, one can note the low occurrence of extreme values (1 and 5), meaning that the bulk of answers varies between 3 values (assuming the IAS is implemented as a 5-point Likert scale following its validation).
The samples using an analogue scale (samples 10a, 10b and 10c in the figure) display a more continuous and progressive spread of answers, seemingly improving the interindividual variability, although potentially displaying a secondary lower mode at \~2 (which might be suggesting the existence of potential clusters of participants).
The correlation matrix between all items shows an overall positive correlation pattern, with highly correlated pairs of items (e.g., Tickle-Itch, Urinate-Defecate, Pain-Wind, Hungry-Thursty) or triplets (Vomit-Sneeze-Cough, Temperature-Muscles-Pain).

```{r}
#| warning: false
#| label: fig-distributions
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"
#| fig-cap: "Top: Distribution of responses across datasets reveals a consistent modal value, typically around 4 or 5 (indicating agreement), except for 'blood sugar' and 'bruise' - in dashed lines - and 'affective touch' (dotted lines) in the Chinese validation sample, which have lower modes. Most responses cluster around the middle values, with few extreme scores (1 and 5). Samples using an analogue scale (10a, 10b, 10c) show a more continuous distribution and increased interindividual variability. Since most samples use Likert scales (discrete), density plots may not be the most accurate representation but were chosen to clearly highlight variability patterns in the data. Bottom: The correlation matrix between all items shows an overall positive correlation pattern, with correlated pairs (e.g., Wind, Burp) of items or triplets (e.g., Vomit, Sneeze and Cough)."

knitr::include_graphics("../../analysis/figures/Figure1.png")
```

<!-- UVA -->
UVA flagged two strongly redundant variables, "itch" and "tickle" - suggesting to remove the latter.
Several more pairs of items were flagged as moderately redundant ("wind" and "burp"; "urinate" and "defecate") and mildly redundant ("sneeze" and "cough"; "heart" and "breathing"; "hungry" and "thirsty").
These patterns consistently appeared in most samples when considered individually. We removed "tickle" from further analysis due to its high redundancy (and because it is absent from some datasets due to translation issues).

<!-- Initial run -->
The HCA highlighted pairs and triplets of items consistently grouped together across samples, such as "wind" and "burp", "sneeze" and "cough", "itch" and "bruise", "urinate" and "defecate", and "pain", "muscles", and "temperature". This pattern was largely replicated by the EGA, with the additional presence of a unique cluster comprising "Sex arousal", "Affective touch", "Temperature", "Pain", "Muscles", and "Taste". EFA suggested the optimal number of factors to be 3, yielding one dimension with expulsion-related items ("burp", "wind", "cough", "sneeze", and "vomit"), a second dimension with visceroceptive items ("heart", "breathing", "hungry", "thirsty", "urinate", and "defecate"), and a third dimension with skin-related items ("bruise" and "blood sugar").

<!-- We discarded: -->

<!-- Taste: Lone item + unstable -->
<!-- Affective_touch: Cross-loadings + unstable -->
<!-- Vomit: Less strongly associated -->
<!-- Itch: Less strongly associated, did not form consistent cluster -->
<!-- More controversial: Temp and Sex_arousal: similar cluster yet less reliable -->

Importantly, this initial structure analysis run highlighted some problematic items: "taste" typically displayed a lone or unstable pattern of associations, "affective touch" exhibited cross-loadings and instability, "vomit" was less strongly associated with other items, and "itch" did not form a consistent cluster. Finally, "temperature" and "sexual arousal" showed redundant patterns of associations but were less reliable. These 6 items were thus removed, and a second run of structure analysis was performed on the remaining 14 items.

```{r}
#| warning: false
#| label: fig-structure
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"
#| fig-cap: "Four structure analysis methods (HCA, EGA, EFA, CFA) were applied and converged on a consistent optimal solution of 14 items formed seven pairs: Hungry-Thirsty, Bruise-Blood sugar, Urinate-Defecate, Muscles-Pain, Breathing-Heart, Cough-Sneeze, Wind-Burp. While EFA suggested 3 factors, CFA confirmed the superiority of the 7-factor model over alternative structures."

knitr::include_graphics("../../analysis/figures/Figure2.png")
```

<!-- Second run -->
HCA and EGA yielded highly consistent results, emphasizing pairs of items, namely Hungry-Thirsty, Bruise-Blood sugar, Urinate-Defecate, Muscles-Pain, Breathing-Heart, Cough-Sneeze, Wind-Burp. HCA also significantly grouped the Urinate-Defecate and Muscles-Pain pairs, as well as expulsion items (Wind-Burp and Cough-Sneeze).
EFA suggested once again 3 factors as the optimal solution, with the first factor including expulsion-related items ("burp", "wind", "cough", "sneeze"), the second factor being related to the Urinate-Defecate pair, and the third factor comprising the remaining items.

<!-- CFA Model Comparison -->
<!-- Name	BIC	RMSEA	Chi2	CFI	BF -->
<!-- fit_g	933909.6	0.124	27,249.398	0.783	Inf -->
<!-- fit_ega	1101773.3	0.035	2,306.303	0.984	NA -->
<!-- fit_egag	1106163.9	0.054	6,842.382	0.952	0 -->
<!-- fit_hclust	1112649.0	0.078	13,296.328	0.907	0 -->
<!-- fit_hclustg	1113829.4	0.078	14,528.732	0.898	0 -->
<!-- fit_efag	1115871.0	0.083	16,591.054	0.883	0 -->
<!-- fit_efa	1115871.0	0.083	16,591.054	0.883	0 -->

<!-- CFA -->
We then fitted and compared using CFA various candidate structures emerging from the previous analyses, including a 1-factor model (the G-model), a 3-factor model (EFA), a 3+1 model (EFA + general factor), a 5-factor model (HCA), a 5+1 model (HCA + general factor), a 7-factor model (EGA), and a 7+1 model (EGA + general factor). The EGA model with 7 factors of item pairs provided the best fit with the lowest RMSEA (0.035), lowest $\chi^2$ (2306.30), highest CFI (0.984). It was followed by the EGA + general factor model (RMSEA = 0.054, $\chi^2$ = 6842.38, CFI = 0.952), and the HCA model with 5 factors (RMSEA = 0.078, $\chi^2$ = 13296.33, CFI = 0.907). The other models performed poorly, with RMSEA > 0.08 and CFI < 0.90. The unique-factor model yielded the lowest BIC (which takes into account the number of factors), followed by the EGA model. All the other models displayed significantly (BIC-based Bayes Factor < 1/100) lower evidence compared to the EGA model.

This result was relatively consistent across unique datasets with the EGA model providing the best fit. In all samples where CFA converged, it showed excellent performance with low RMSEA (≈ 0.025–0.046), high CFI (≥ 0.97), and substantially lower $\chi^2$ compared to alternative models. Adding a general factor systematically worsened fit, while hierarchical clustering and EFA-based models showed poorer performance (RMSEA > 0.06, CFI < 0.93). Although BIC values occasionally favored alternative specifications, Bayes factor comparisons overwhelmingly supported the EGA solution, confirming its robustness across datasets.

<!-- did not converge: df1b, df2, df3, df5, df10 (missing items), df12 (errors), df15 (errors), df16,  -->

## Discussion

<!-- Aim and previous factor structures -->
Study 1 aimed to systematically evaluate the structure of the IAS. Previous research reported conflicting models, from unidimensional [@brand2023] to bifactor [@campos2021; @lin2023] or two-factor solutions [@murphy2019; @koike2023]. Concerns about specific items (e.g., low discrimination, local dependencies, and items reflecting phenomena less accessible to interoception) prompted a mega-analytic approach using both factor-analytic and network-based methods.

<!-- Item pairs and optimal structure -->
Analyses revealed that seven item pairs (e.g., Hunger–Thirst, Urinate–Defecate, Cough–Sneeze) provided a robust and replicable structure, leading to a refined 14-item IAS. These pairs consistently emerged across datasets and yielded superior fit indices compared to alternative models. This suggests that the IAS does not measure a broad latent construct of interoceptive accuracy but reflects clusters of tightly coupled items tied to specific bodily sensations.

<!-- Implications for scale interpretation -->
The presence of redundant or unstable items (e.g., Tickle, Taste, Affective Touch) undermines interpretability and likely explains inconsistencies in prior factor-analytic findings [@campos2021; @lin2023]. Nevertheless, the structured item pairs correspond to intuitive bodily domains—visceral, expulsion, musculoskeletal—supporting ecological validity. These findings highlight the need for targeted scale refinement and motivate future development of context-sensitive, multidimensional interoception measures.

<!-- Summary -->
In summary, study 1 suggests that once problematic items are removed, the IAS is best conceptualized as a set of item clusters rather than a unidimensional scale. This structural insight provides the foundation for Study 2, where dispositional correlates are examined to further evaluate the IAS's construct validity.

# Study 2

Building on the structural findings from Study 1, the second study turns to the validity of the IAS by examining its dispositional correlates. Although the IAS has been widely used, evidence regarding its associations with personality traits, affective dispositions, and related psychological constructs remains scattered and inconsistent [@murphy2019; @brand2023; @lin2023; @todd2022]. Study 2 therefore had two primary goals. First, to provide a robust overview of the dispositional correlates of the IAS by leveraging the large number of available datasets. Second, to evaluate the practical benefits of using the refined 14-item version of the scale identified in Study 1, compared to the original 21-item unidimensional solution, in order to assess whether structural improvements translate into clearer and more reliable external associations.

## Methods

### Materials

We selected measures that appeared in multiple datasets from Study 1 and/or were relevant given the scope of the study, i.e., constructs related to physiology, mood, personality, psychopathology, and beliefs and misbeliefs. We merged scores of regular and abridged versions for conciseness where applicable.

<!-- Interoception Related -->
#### Interoception Related

The Multidimensional Assessment of Interoceptive Awareness Version-2 [MAIA-2, @mehling2018multidimensional] is a 37-item questionnaire assessing eight dimensions of interoception: Noticing, Not-Distracting, Not-Worrying, Attention Regulation, Emotional Awareness, Self-Regulation, Body Listening, and Trust. Responses are rated on a 6-point Likert scale (e.g., “I notice when I am uncomfortable in my body”).

The Body Perception Questionnaire Short Form [BPQ-SF, @cabrera2018assessing] contains 46 items measuring body awareness and autonomic reactivity on a 5-point Likert scale (e.g., “My heart often beats irregularly”). The very-short form (BPQ-VSF) contains 12 items from the body awareness subscale. For this study, scores from both versions were combined.

The Interoceptive Confusion Questionnaire [ICQ, @brewer2016alexithymia] consists of 20 items rated on a 5-point Likert scale (e.g., “I cannot tell when my muscles are sore or tight”) assessing difficulties in interpreting non-affective physiological states, such as pain or hunger.

The Toronto Alexithymia Scale [TAS-20, @bagby1994twenty] contains 20 items rated on a 5-point forced-choice scale (e.g., “I have feelings I can’t quite identify”), divided into three dimensions: difficulty identifying feelings, difficulty describing feelings, and externally oriented thinking.

The Bermond–Vorst Alexithymia Questionnaire [BVAQ, @vorst2001validity] consists of 40 items across five subscales: fantasising, identifying, analysing, verbalising, and emotionalising. Items are rated on a 5-point Likert scale. The BVAQ also provides two higher-order factors, namely affective and cognitive (e.g., “I like to tell others how I feel”).

<!-- Mood and Anxiety -->
#### Mood and Anxiety Related
The Beck Depression Inventory-II [BDI-II, @beck1996beck] is a 21-item measure of depressive symptom severity, including psychological (e.g., guilt) and physiological (e.g., loss of energy) components. Items are rated on a 4-point scale (e.g., “I feel sad much of the time”).

The Patient Health Questionnaire-4 [PHQ-4, @kroenke2009ultra] is a 4-item screening tool for depression and anxiety, rated on a 4-point Likert scale (e.g., “Little interest or pleasure in doing things”; “Not being able to stop or control worrying”). The PHQ-2 [PHQ-2, @kroenke2003patient] and General Anxiety Disorder-2 [GAD-2, @kroenke2007anxiety] scores were combined with the PHQ-4 scores.

The Patient Health Questionnaire-9 [PHQ-9, @kroenke2001phq] includes 9 items assessing depressive symptoms on a 4-point Likert scale (e.g., “Feeling tired or having little energy?”).

The Patient Health Questionnaire-15 [PHQ-15, @kroenke2002phq] assesses somatic symptom distress with 15 items rated on a 3-point scale (e.g., “Over the last week, how often have you been bothered by back pain?”).

The Short Mood and Feelings Questionnaire [SMFQ, @angold1995development] is a 13-item measure of recent depressive symptoms in children aged 6–17, rated on a 3-point scale (e.g., “I felt miserable or unhappy”).

The State–Trait Anxiety Inventory Trait Version [STAI-T, @spielberger1970manual] is a 20-item trait anxiety questionnaire rated on a 4-point Likert scale (e.g., “I worry too much about something that really doesn’t matter”). Scores from the shorter version of the STAIT-5 [@zsido2020development] were combined with the STAI-T.

The Generalized Anxiety Disorder Scale [GAD-7, @spitzer2006brief] is a 7-item measure of generalized anxiety symptoms rated on a 4-point Likert scale (e.g., “Not being able to stop or control worrying”).

<!-- Personality -->
#### Personality
The NEO Five-Factor Inventory, neuroticism subscale [NEO-FFI, @costa1992personality] is a 12-item measure of neuroticism, assessing the general tendency to experience negative emotions such as fear, sadness, or disgust. Items are rated on a 5-point Likert scale (e.g., “I often feel inferior to others”).

The Mini International Personality Item Pool [MINI-IPIP6, @sibley2011mini] includes 24 items measuring six broad traits: Agreeableness, Conscientiousness, Neuroticism, Openness, Extraversion, and Honesty-Humility. Items were scored on visual analogue scales in our samples (originally 7-point Likert scales; e.g., “Have a vivid imagination”).

The Big Five Inventory Short Form [BFI-S, @rammstedt2007measuring] is a 10-item measure of the Big Five traits, rated on a 7-point Likert scale (e.g., “I see myself as someone who … is considerate and kind to almost everyone”).

<!-- Psychopathology -->

#### Psychopathology
The Personality Inventory for DSM-5 Short Form [PID-5-SF, @thimm2016personality] consists of 25 items rated on a 4-point Likert scale, measuring five domains: disinhibition, antagonism, detachment, negative affect, and psychoticism (e.g., “Plenty of people are out to get me”).

The Schizotypal Personality Questionnaire – Brief Revised [SPQ-BR, @davidson2016schizotypal] contains 32 items rated on a 5-point Likert scale, assessing four primary dimensions: cognitive-perceptual, interpersonal, disorganized, and social anxiety. These are subdivided into nine secondary factors, including constricted affect, eccentricity, magical thinking, lack of close friends, odd speech, referential thinking, social anxiety, suspiciousness, and unusual perceptions (e.g., “Plenty of people are out to get me”).

The McLean Screening Instrument for Borderline Personality Disorder [MSI-BPD, @zanarini2003zanarini] is a 10-item dichotomous screening measure for borderline personality disorder (present/absent; e.g., “Have any of your closest relationships been troubled by a lot of arguments or repeated breakups?”).

The Autism Spectrum Quotient Short Form [ASQ-S, @hoekstra2011construction] is a 28-item measure of five autistic traits: social skills, adherence to routines, cognitive flexibility, imagination, and patterns/numbers. Items are rated on a 4-point Likert scale (e.g., “I find it difficult to work out people’s intentions”).

<!-- Beliefs and Misbeliefs -->
#### Beliefs and Misbeliefs
The Generic Conspiracist Beliefs Scale [GCBS, @brotherton2013measuring] is a 15-item measure assessing five dimensions of conspiracy beliefs: government malfeasance, extraterrestrial cover-up, malevolent global conspiracies, personal wellbeing, and control of information. Items are scored on a 5-point Likert scale (e.g., “Secret organizations communicate with extraterrestrials, but keep this fact from the public”).

The Primal World Beliefs Inventory [PI-99, @clifton2019primal] is a 99-item measure assessing 26 high-order beliefs about the world. Items are scored on a 5-point Likert scale (e.g., “Nearly everything in the world is beautiful”) The PI-18 [@clifton2021brief] is a short form of the orginal questionnaire assessing the 4 main world beliefs: the world is Good (vs. Bad) and its three dimensions: Safe (vs. Dangerous), Enticing (vs. Dull), and Alive (vs. Mechanistic). Scores from the PI-18 scores were combined with the PI-99.

The Lie Scale [LIE, @makowski2023structure] is a 16-item measure of lying tendencies across four components: Ability, Frequency, Negativity, and Contextuality. Items are scored on visual analogue scales (e.g., “I am a good liar”).

### Data Analysis

<!-- Preprocessing and standardization -->
All variables from the 16 open-access datasets will be extracted and standardized (z-scored). For continuous predictors, we will first ensure consistent ranges across datasets; when necessary, ranges will be rescaled to achieve comparability. In cases where scales of different lengths measure the same construct, scores will be harmonized: when possible, means will be computed from overlapping items, while in other cases total scores will be normalized by the number of items. We will collapse redundant subscales (e.g., PHQ-2 and GAD-2 within PHQ-4) into unified composite scores. Similarly, short forms of trait anxiety (e.g., STAI-5-trait) will be collapsed with their longer counterpart (STAI-T), with scores computed as the mean to ensure comparable ranges. Shorter formats of the PI (PI-18) will be combined with scores of its longer version (PI-99). Missing data will be handled via listwise deletion on a per-model basis prior to fitting.

<!-- Multilevel modelling approach -->
Associations will be estimated using multilevel regression models implemented in the glmmTMB package [@glmTMB]. Each predictor will be entered separately, with the seven IAS item pairs identified in Study 1 (“Hungry–Thirsty”, “Muscles–Pain”, “Wind–Burp”, “Urinate–Defecate”, “Breathing–Heart”, “Bruise–Blood”, “Cough–Sneeze”) and the original IAS scale serving as outcomes. Models will include random intercepts and slopes by sample to account for between-study heterogeneity; when models fail to converge, a reduced structure with random intercepts only will be used. This approach will treat datasets as clusters, thereby providing pooled estimates of the associations while accounting for both within- and between-sample variability, akin to a meta-analytic framework.

<!-- Extraction of parameters -->
From each fitted model, we will extract standardized regression coefficients (β), 95% confidence intervals, p-values, convergence status, and effective sample size. Participants identifying as “Other” will be excluded from analyses due to limited representation.

## Results

<!-- Demographic correlates -->
Age was positively associated across Ias correlates (mean r = .12), with all correlations significant. Gender showed near-zero effects (mean r = −.01), with 63% reaching significance.

<!-- Interoceptive measures -->
Several MAIA subscales showed consistent positive associations with the IAS: Noticing, Attention Regulation, Trusting, Emotional Awareness, Listening, and Self-Regulation (βs ≈ .18–.26). These effects were generally strongest for the full IAS, but often replicated with Hungry/Thirsty or Bruise/Blood pairings. By contrast, Autonomic Reactivity (BPQ; β = −.17) was reliably negative across most pairings, while Body Awareness (BPQ; β = .12) was weak and only reliable for Hungry/Thirsty and Bruise/Blood. Alexithymia measures were consistently negative, with robust effects across both the full IAS and pairings, except for the Affective dimension of the BVAQ which showed only a positive association with the Bruise/Blood sugar pairing (β = .04).

<!-- Mood and psychopathology -->
Depression (BDI; β = −.20) and anxiety (STAI-T; β = −.19) were consistently negative across the IAS and most pairings, with Hungry/Thirsty again emerging as a particularly sensitive correlate. Broader symptom indices (PHQ-15, PHQ-4, GAD-7) showed weaker but often reliable effects, although some were restricted to certain pairings (e.g., PHQ-15 with Hungry/Thirsty, Bruise/Blood). By contrast, PHQ-9 and MFQ averaged negatively but were largely nonsignificant, even with the full scale. Autistic traits were generally negative, especially Imagination (β = −.18), which showed broad consistency across pairings except Cough/Sneeze.

<!-- Personality and beliefs -->
Within the Big Five, Openness (β = .14), Agreeableness (β = .05), and Conscientiousness (β = .03) showed small but consistent positive associations, most reliably captured by the full IAS but also evident in pairings. Honesty–Humility was uniquely negative, with significant associations for both the full scale and Hungry/Thirsty. Worldview measures were generally null, though belief in a comprehensible world (β = .11) and an “alive” world (β = .09) showed small but consistent positive associations, mainly with the full IAS and Hungry/Thirsty.

<!-- Differences between the full IAS and item pairings -->
Overall, the full IAS yielded the largest number of significant associations (25), but the Hungry/Thirsty (16) and Bruise/Blood sugar (11) pairings frequently showed equally strong or stronger effects. These item-level pairings often captured variance that the full scale did not, particularly for interoceptive and mood measures. In contrast, other pairings (e.g., Cough/Sneeze, Wind/Burp, Urinate/Defecate) contributed little unique variance. Together, these results suggest that while the full IAS captures broad dispositional tendencies, select pairings may provide more fine-grained sensitivity to specific domains, particularly interoception and affect.

## Discussion

<!-- Aim and conceptual focus -->
Study 2 examined the dispositional correlates of the IAS and tested whether the refined 14-item version offers practical advantages over the original 21-item scale. The study provides insight into the conceptual boundaries of interoceptive self-reported accuracy across psychological domains.

<!-- Interoception and mental health + context dependency -->
The IAS converged robustly with other interoceptive self-report measures, particularly facets of the MAIA, and correlated more strongly with dynamic bodily monitoring (Autonomic Reactivity on the BPQ) than general body awareness [@brand2023; @campos2021; @todd2022]. Item pairings reflecting visceral sensations, such as Hunger–Thirst, were consistently linked to lower depressive and anxiety symptoms [@brand2023; @lin2023], supporting theoretical models connecting interoception to emotion regulation and monitoring of internal bodily states [@khalsa2018interoception; @suksasilp2022towards].

At the same time, these associations were context-dependent: not all item pairings or bodily domains related equally to mood or somatic symptoms [@murphy2019; @koike2023]. Emerging evidence suggests that interoceptive abilities (e.g., sensitivity, precision and metacognitive insight) may not uniformly predict mental health outcomes and may depend on measurement approach or sample characteristics [@@banellis2025interoceptive].

<!-- Personality, worldview, and boundaries of interoception -->
Associations with broader personality traits and high-order beliefs were modest. Small positive correlations with Openness and Agreeableness and weak associations with primal world beliefs (e.g., perceiving the world as alive or comprehensible) suggest that interoception may subtly interact with higher-level cognitive frameworks [@simkute2025interoception; @haustein2024interoception; @gabriele2022dissociations]. However, the effects were weak and inconsistent, reinforcing that interoception primarily aligns with bodily self-perception rather than broad personality or worldview constructs.

<!-- Contextual sensitivity and scale refinement -->
Certain item pairings (e.g., Hunger–Thirst, Bruise–Blood sugar) captured associations with mood and interoception as strongly as, or stronger than, the full 21-item scale. This highlights the contextual nature of interoceptive accuracy: interoceptive tendencies manifest variably across bodily domains [@desmedt2022measures; @khalsa2018], emphasizing the need for context-specific items.

The refined 14-item IAS performed comparably to the original scale, and item pairings sometimes provided greater sensitivity to specific interoceptive or affective domains, demonstrating that structural refinement can enhance efficiency and interpretability without sacrificing validity.

<!-- Summary -->
Study 2 shows that interoceptive self-report is adaptive, context-sensitive, and most predictive in relation to mood and dynamic bodily monitoring. The refined IAS preserves construct validity while enabling more targeted assessment of domain-specific interoceptive tendencies.

# General Discussion

<!-- AIms -->
The present study provides a comprehensive evaluation of the IAS, examining both its structure (study 1) and its dispositional correlates (study 2). These studies offer novel insights into the psychometric characteristics, conceptual interpretation, and practical applications of the IAS. In this discussion, these findings are contextualized within the broader landscape of interoception research.

<!-- Study 1 findings -->
Study 1 demonstrated that the IAS is best conceptualised as comprising tightly coupled item pairs, reflecting discrete bodily domains such as visceral sensations, musculoskeletal cues, and expulsion phenomena. The pairs emerged after several items (e.g., Tickle, Taste, Affective Touch) were removed due to their instability, redundancy, or poor associations with the other items. The resulting 14-item IAS retained the psychometric integrity of the original scale while improving interpretability and efficiency.

A short version of the IAS has been proposed before.  @lin2023 suggested a 12-item version by removing items with local dependency, high difficulty, and/or low discrimination. Their parallel analysis suggested unidimensionality, although parallel analysis is known to underestimate factor numbers in multidimensional data [@golino2017exploratory; @markos2024dimensionality].

In contrast, our study applied multiple methods including EGA, FA, and HCA yet the same pairs consistently emerged across datasets and demonstrated superior fit compared to alternative models, which challenges the assumption of unidimensionality of the IAS [@murphy2019; @brand2023]. Although the short version of the IAS by  @lin2023 favoured a one-factor solution, their 12-item IAS was derived by removing locally dependent items identified through Item Response Theory (IRT).

We found that these locally dependent items can instead be interpreted as meaningful pairs (e.g., Heart–Breath), reflecting distinct bodily domains. Because IRT assumes unidimensionality [@nguyen2014introduction], it may have treated these natural item clusters as statistical artefacts rather than as theoretically relevant structures within interoceptive accuracy.

<!-- Study 2 findings -->
Study 2 provided further evidence for a domain-specific perspective. Certain item pairings, particularly Hunger-Thirst, Bruise-Blood sugar, and Muscles-Pain, showed consistent negative correlations with depression and anxiety measures, sometimes approaching or even surpassing the strength of associations observed for the full IAS scale. Notably, Hunger–Thirst and Bruise–Blood sugar were also the only pairs, out of the seven examined, that displayed significant positive correlations with the BPQ Body Awareness subscale, whereas the full IAS scale showed no such association.

Additionally, specific item pairings revealed distinct patterns of association with other interoceptive measures and mental health indices that were not captured by the full IAS. For instance, while the full IAS showed no significant associations with Borderline Personality, the Hunger-Thirst pair exhibited negative correlations. These findings collectively suggest that item-level analyses may uncover more nuanced relationships than those identified by the overall scale. This pattern highlights that questionnaires like the IAS may capture multiple facets of interoception rather than a single underlying dimension. Consequently, adopting a domain-specific approach could enhance construct validity by better reflecting the heterogeneity of interoceptive experiences.

These findings are in line with inconsistencies in the literature concerning associations between the IAS and the BPQ Body Awareness. While some studies report significant negative associations [@lin2023], others have observed significant positive associations  [@brand2023; @campos2021; @gaggero2021; @todd2022], and still others none at all  [@murphy2019]. In our study the full IAS was also not associated with the BPQ Body Awareness.

This pattern is consistent with the idea that this subscale may capture maladaptive bodily experiences or autonomic dysfunction, rather than interoceptive accuracy per se. However, the Body Awareness subscale of the BPQ has also been criticized for its ambiguous instructions. It is often unclear whether respondents should interpret the items as assessing attentional focus on bodily sensations, perceived accuracy in detecting them, or simply the frequency of experiencing such signals. As a result, studies have reported that observed associations vary depending on how the BPQ Body Awareness is interpreted [@campos2021; @gabriele2022dissociations; @gaggero2021]. As such, it is unclear whether the lack of a relationship between the IAS and the BPQ Body Awareness in this study reflects a genuine divergence between the constructs measured by the two scales, or instead arises from ambiguities in how the BPQ Body Awareness is operationalized and interpreted across studies.

We highlight that some IAS pairings may capture interoceptive self-reported accuracy more effectively than the full IAS measure, raising questions about the interpretability of the full scale. In contrast , other pairings (e.g., Cough–Sneeze, Wind–Burp) were weakly or inconsistently related to other variables, raising concerns about their usefulness. Together, these findings indicate that the strength and relevance of interoceptive self-reported accuracy may vary systematically across physiological domains.

At the same time, the interpretation of such associations is complicated by the lack of contextualisation in the items themselves. For example, a racing heart is often associated with anxiety, yet the same sensation can also arise during exercise or other forms of physiological arousal. Without contextual anchors, it remains unclear which bodily states participants are drawing upon when responding, highlighting the interpretive ambiguity of self-reports. In line with prior work emphasizing the importance of contextual factors in the measurement of interoception [@vlemincx2023novel], our results suggest that the value of self-report lies less in broad aggregation across domains and more in the targeted assessment of physiologically and situationally meaningful contexts.

<!-- Correlates of IAS Across Psychological Domains -->
The dispositional correlates of the IAS help delineate the conceptual boundaries of interoceptive accuracy as a construct. Self-reported interoceptive accuracy, as assessed by the IAS, has been robustly associated with mood-related outcomes, particularly depression and anxiety, underscoring its relevance for affective regulation and bodily state monitoring [@benau2023self; @brand2023]. This aligns with a broader body of research linking affective conditions such as anxiety and depression with altered interoceptive processing [e.g.,@garfinkel2015knowing].

However, recent findings have also reported null or inconsistent associations between interoceptive abilities and mental health symptomatology, suggesting that such relationships may not generalize across all interoceptive domains or measurement modalities (@banellis2025interoceptive; @jenkinson2024interoception). For instance, a recent review by @jenkinson2024interoception found no consistent evidence linking cardiac interoceptive accuracy tasks with anxiety or depression across multiple meta-analyses.Collectively, these findings indicate that relationships between interoceptive accuracy and mental health may vary across bodily domains and measurement modalities. This suggests that the IAS, while informative, may benefit from a more differentiated approach that accounts for domain-specific and contextual influences on self-reported interoceptive accuracy.

In contrast to the relatively stronger associations observed with affective measures, correlations between IAS scores and personality traits were small but directionally consistent. Modest positive associations emerged with Openness, Agreeableness, and Conscientiousness, while Neuroticism and Extraversion showed near-zero relationships, and Honesty–Humility displayed weak negative correlations. This pattern suggests that self-reported interoceptive accuracy, as indexed by the IAS, shares limited but systematic overlap with broad dispositional tendencies -particularly those reflecting intellectual curiosity and creativity, patience and tolerance in interpersonal interactions, and diligence and organizational tendencies [@ashton2007empirical].

Recent evidence supports the view that relationships between interoceptive measures and personality are multifaceted rather than uniform. For instance, @haustein2024interoception reported significant associations between interoceptive awareness (MAIA) and all personality traits except Agreeableness, with Openness showing positive correlations with Body Listening and Emotional Awareness, and Neuroticism showing negative correlations with Self- and Attention Regulation. Similarly, @simkute2025interoception found that task-based indices of interoceptive sensitivity related negatively to Extraversion. Collectively, these findings indicate that interoceptive constructs show distinct and sometimes opposing patterns of association across measurement modalities, suggesting that the IAS relates to a specific, self-evaluative aspect of interoceptive functioning rather than reflecting a general interoceptive or personality-based trait.

A potential key reason for discrepancies in observed associations, including those seen with the IAS, is that self-report questionnaires in interoception often tap into distinct constructs rather than a common underlying dimension [@desmedt2022measures]. Measures labeled as interoceptive sensibility, awareness, or self-reported accuracy capture partially overlapping but empirically distinct facets, with low intercorrelations between questionnaires and items. Consequently, differences in associations with psychological outcomes or personality traits may reflect differences in what each questionnaire assesses rather than true variation in interoceptive functioning.

<!-- IAS limitations and potential improvements -->
These two studies revealed three main points. First,  in the datasets that included analogue-style ratings, these scales captured greater variability, further enhancing the ability to detect individual differences. Second, a shortened, 14-item version of the IAS retained strong psychometric integrity while removing unstable or redundant items. Third, specific item pairings allowed more precise differentiation among individuals and often outperformed global scores in predictive value. Nevertheless, even this refined IAS has important limitations.

While the 14-item version improves interpretability and efficiency, it continues to underrepresent key interoceptive modalities such as cardiac, respiratory, and nuanced somatosensory signals, limiting its ability to capture the full breadth of interoceptive processes. This shortcoming underscores the need for broader coverage in future scale development. At the same time, analogue-style items and item pairings offer clear advantages: they increase response variability, allow more precise differentiation among individuals, and provide domain-specific indicators that often outperform global scores.

Building on these insights, a next-generation interoception scale could balance breadth (multiple items per modality) with depth (domain-specific subscales), retain the benefits of analogue or wider rating ranges (e.g., 1–7 instead of 1–5), and optimize coverage across physiologically and psychologically relevant bodily domains. Such a scale would improve interpretability, sensitivity, and practical utility, ultimately enabling a more nuanced mapping of interoceptive abilities onto affective, cognitive, and health-related outcomes.

<!-- Limitations -->
One limitation worth noting is that we did not evaluate the IAS alongside other widely used questionnaires, such as the BPQ or MAIA. Although both are frequently employed in interoception research, they present important limitations. For example, the MAIA-2 contains subscales (e.g., Not-Worrying, Not-Distracting) with weak reliability and minimal relation to a general interoception factor [@ferentzi2021examining; @rogowska2023validation]. Several other subscales primarily reflect higher-level cognitive tendencies, such as mindfulness and attentional regulation, rather than interoceptive body awareness [@todd2022; @vig2023self]. Similarly, the BPQ’s Body Awareness and Autonomic Reactivity subscales show conceptual and psychometric concerns, and BPQ scores often overlap substantially with negative affect, raising questions about their specificity as measures of interoceptive sensibility [@campos2021; @gabriele2022dissociations; @vig2023self].

Because the IAS is a newly developed questionnaire and the only one designed to assess interoceptive self-reported accuracy - a dimension previously assessed solely through objective tasks such as the HDT and HCT - our focus was on evaluating its structure and correlates without introducing established measures with documented shortcomings. Nonetheless, future research should employ novel methods, such as Exploratory Graph Analysis, to clarify the convergent and divergent validity of these questionnaires across interoceptive domains.

<!-- Conclusion -->
Taken together, these studies advance understanding of the IAS and interoceptive self-report more broadly. The refined 14-item IAS provides a more efficient and interpretable measure, capturing domain-specific interoceptive signals that show differential associations with psychological outcomes. Associations were consistently stronger with mood measures and compared to personality and beliefs, indicating that self-reported interoceptive accuracy relates more closely to affective experience than to stable dispositional traits.

At the same time, current findings highlight limitations in modality coverage and item specificity, suggesting that future scale development should incorporate multimodal, context-sensitive, and continuous analogue ratings or wider discrete scales to better capture the richness of interoceptive processes and clarify their role in affective, cognitive, and health-related outcomes [@desmedt2022measures; @vlemincx2023novel]. Ultimately, these studies show that the IAS is most informative when treated as a set of domain-specific, context-sensitive indicators rather than as a broad global index, advancing both the measurement and conceptual understanding of interoception.

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::